---
title: "07_GSEA_expression_data"
author: "Julin"
date: "11/13/2021"
output: html_document
editor_options: 
  chunk_output_type: inline
---

Write tables of voom stabilized expression data, corrected (or not) for experiment and block effects

This is modified for the week 4 data.  Week 2 WGCNA used `voom_expression.e1and3.resid.exp_gt.root.plusR500.txt.gz`

Here I will take a similar approach, but factor out block.  (alternate would be to combine this data with week 2 data)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,error=TRUE)
library(edgeR)
library(tidyverse)
library(readxl)
```


helper function to get residuals
```{r}
getResid <- function(gene_expression, block1, block2=NA, block3=NA, verbose = FALSE) { #
  if(is.na(block2[1])) lm1 <- lm(gene_expression ~ block1)
  else if(is.na(block3[1])) lm1 <- lm(gene_expression ~ block1 + block2)
  else lm1 <- lm(gene_expression ~ block1 + block2 + block3)
  if (verbose) print(summary(lm1))
residuals(lm1)
}
```

# read dge objects 
```{r}
load("../../output/edgeR_dge_objects_exp1_week4.Rdata")
```

## voom transformation
```{r}
## dge
root.design <- with(dge.root$samples, model.matrix(~ autoclave))
leaf.design <- with(dge.leaf$samples, model.matrix(~ autoclave))
## 
dge.root <- dge.root[rowSums(cpm(dge.root)>1) >= 1,,keep.lib.sizes=FALSE]
dge.root <- calcNormFactors(dge.root)
#
dge.leaf <- dge.leaf[rowSums(cpm(dge.leaf)>1) >= 1,,keep.lib.sizes=FALSE]
dge.leaf <- calcNormFactors(dge.leaf)

#
voom.root <- voom(dge.root, design = root.design, block = as.factor(dge.root$samples$block))
voom.leaf <- voom(dge.leaf, design = leaf.design, block = as.factor(dge.leaf$samples$block))
```

## alternative: take residuals to remove block.
```{r}

leaf.resid <- t(apply(voom.leaf$E, 1, getResid, dge.leaf$samples$block)) %>%
  as.data.frame() %>%
  mutate(genes=voom.leaf$genes$genes) %>%
  select(genes, everything())

root.resid <- t(apply(voom.root$E, 1, getResid, dge.root$samples$block)) %>%
  as.data.frame() %>%
  mutate(genes=voom.root$genes$genes) %>%
  select(genes, everything())

write_tsv(root.resid, "../../output/voom_expression.e1_week4.resid.root.txt.gz")
write_tsv(leaf.resid, "../../output/voom_expression.e1_week4.resid.leaf.txt.gz")

```

