---
title: "custom_categories_v5_making"
author: "Kazu"
date: "8/6/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Summary

History
* copied from "06_f_cor_table.Rmd"
* Use modified N_Peng_2007 to eliminate duplicated gene from root
* Use updated plusN_Wang2003 (080621)

To Do
*  Expand DEGs with FC threshold for all custom categoreis
* new custom categoreis (v4) that removed duplicated genes in N_peng_2007
* new custom categories (v5) that has updated plusN_Wang2003

# prep
```{r}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE,error=TRUE)
library(tidyverse);library(GGally);library(readxl)
library(readr)
library(stringr)
# https://bioconductor.org/packages/devel/workflows/vignettes/maEndToEnd/inst/doc/MA-Workflow.html
# functions for expression pattern
#source("../../../tools/Expression_pattern_graph.R",chdir=TRUE)
library(scales) # for muted function
library(ggdendro) # for dendrogram
library(cowplot)
library(ggnewscale)
```



# copied custom gene categories created in "06_d_cor_permutation.Rmd" D

# Instead of root.exp1and3.all, use Br.v3.0anno.At.BLAST.highscore
```{r}
# annotation file for v3.0annotation
Br.v3.0.At.BLAST <- read_csv(file.path("annotation","output","Brapa_v3.0_annotated.csv")) 
# This annotation is redundant with name (Br grene). Eg 
Br.v3.0.At.BLAST %>% filter(name=="BraA01g040570.3C")
# reduce the redundancy (112418)
Br.v3.0anno.At.BLAST.highscore <- Br.v3.0.At.BLAST %>% group_by(name) %>% arrange(desc(score)) %>% dplyr::slice(1)
# reduce the redundancy for Br (051120) (do not use)
#Br.v3.0anno.At.BLAST.highscore.Br <- Br.v3.0.At.BLAST %>% group_by(AGI) %>% arrange(desc(score)) %>% slice(1) #%>% View()
```



## custom (with Nishida_2017, 070720)
* note. some are no genes in **** (eg. "minusMn_Nishida_2017_root_FC_down")
```{r message=FALSE, warning=TRUE, error=TRUE}
custom.categories <- list.files(path=file.path("annotation","input","custom_categories"),pattern="(FC.csv.gz|FC.WT.csv.gz|DEGs.all.anno.csv.gz|DEGs.all.anno.csv|DEGs.csv.gz|Fe_def_KIM_2019_pnas.1916892116.sd01.csv)",full.names = TRUE) # not working, why??????? 061221
#
custom.categories.filename <- list.files(path=file.path("annotation","input","custom_categories"),pattern="(FC.csv.gz|FC.WT.csv.gz|DEGs.all.anno.csv.gz|DEGs.all.anno.csv|DEGs.csv.gz|Fe_def_KIM_2019_pnas.1916892116.sd01.csv)",full.names = FALSE) # 39 data, 15 publications
# read all csv.gz files
custom.categoreis.map <- custom.categories %>% map(read_csv) # tidyverse (or purr) way of lapply
# using file name for each data.frame
names(custom.categoreis.map) <- custom.categories.filename
```

# combined table (root): 
```{r, message=FALSE, warning=FALSE, error=TRUE}
all.data <-
  # root.exp1and3.FDR0.1 vs Al_Ligaba-OSena_2017_root
  #root.exp1plus3.all %>%
  Br.v3.0anno.At.BLAST.highscore %>% left_join(custom.categoreis.map[["plusAl_Ligaba-OSena_2017_root.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI")  %>% dplyr::rename(!!c(`plusAl_\nLigaba-OSena_2017_\nroot`="logFC",`plusAl_Ligaba-OSena_2017_root_FDR`="FDR"))  %>%
  # root.exp1and3.FDR0.1 vs cold.root.Kilian_2013 (lmer)
  left_join(custom.categoreis.map[["cold_Kilian_2007.root.fits.summary.AGI.FC.csv.gz"]] %>% dplyr::select(AGI,mean.logFC.AGI), by="AGI") %>% 
  dplyr::rename(`cold\nKilian_2013_\nroot`="mean.logFC.AGI") %>% 
  # root.exp1and3.FDR0.1 vs Fe_deficiency.Kailasam2019 
  left_join(custom.categoreis.map[["minusFe_Kailasam2019.DEGs.all.anno.csv"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>% 
  dplyr::rename(!!c(`minusFe_\nKailasam2019_\nroot`="logFC",`minusFe_Kailasam2019_root_FDR`="FDR")) %>%
  # root.exp1and3.FDR0.1 vs Fe_deficiency.Kailasam2019 
  left_join(custom.categoreis.map[["minusFe_Kim2019.fits.summary.AGI.FC.WT.csv.gz"]] %>% dplyr::select(AGI,mean.FC.AGI,`FDR_trtFe-`), by="AGI") %>%
  dplyr::rename(!!c(`minusFe_\nKim2019_\nroot`="mean.FC.AGI",`minusFe_Kim2019_root_FDR`="FDR_trtFe-")) %>%
  # root.exp1and3.FDR0.1 vs heat.root.Kilian_2013.AGI.exp.value.FC.csv.gz (lmer)
  left_join(custom.categoreis.map[["Heat_Kilian_2007.root.fits.summary.AGI.FC.csv.gz"]] %>% dplyr::select(AGI,mean.logFC.AGI), by="AGI") %>% 
  dplyr::rename(`heat_\nKilian_2013_\nroot`="mean.logFC.AGI") %>%
  # root.exp1and3.FDR0.1 vs Mg_high_Niu_2016_root.DEGs.csv.gz
  left_join(custom.categoreis.map[["plusMg_Niu_2016_root.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>% dplyr::rename(!!c(`plusMg_\nNiu_2016_\nroot`="logFC",`plusMg_Niu_2016_root_FDR`="FDR")) %>%
  # root.exp1and3.FDR0.1 vs Mg_low_Niu_2016_root.DEGs.csv.gz
  left_join(custom.categoreis.map[["minusMg_Niu_2016_root.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>%
  dplyr::rename(!!c(`minusMg_\nNiu_2016_\nroot`="logFC",`minusMg_Niu_2016_root_FDR`="FDR")) %>% #View()
  # root.exp1and3.FDR0.1 vs minusPi_1d_Liu_2016_root.DEGs.csv.gz
  left_join(custom.categoreis.map[["minusPi_1d_Liu_2016_root.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>%
  dplyr::rename(!!c(`minusPi_1d_\nLiu_2016_\nroot`="logFC",`minusPi_1d_Liu_2016_root_FDR`="FDR")) %>% #View()
  # root.exp1and3.FDR0.1 vs minusPi_3d_Liu_2016_root.DEGs.csv.gz
  left_join(custom.categoreis.map[["minusPi_3d_Liu_2016_root.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>%
  dplyr::rename(!!c(`minusPi_3d_\nLiu_2016_\nroot`="logFC",`minusPi_3d_Liu_2016_root_FDR`="FDR")) %>% #View()
  # root.exp1and3.FDR0.1 vs Mn_deficiency.Rodriguez-Celma_2016.DEGs.all.anno.csv.gz
  left_join(custom.categoreis.map[["minusMn_Rodriguez-Celma_2016.DEGs.all.anno.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>% dplyr::rename(!!c(`minusMn_\nRodriguez-Celma_2016_\nroot`="logFC",`minusMn_Rodriguez-Celma_2016_root_FDR`="FDR")) %>% #View()
  # root.exp1and3.FDR0.1 vs N_Wang2013.root.fits.summary.AGI.FC.csv.gz
  left_join(custom.categoreis.map[["plusN_Wang2003.root.fits.summary.AGI.FC.csv.gz"]] %>%
              dplyr::select(AGI,mean.FC.AGI,FDR_trtKCl), by="AGI") %>%
  dplyr::rename(!!c(`plusN_\nWang2003_\nroot`="mean.FC.AGI",`plusN_Wang2003_root_FDR`="FDR_trtKCl")) %>% 
  # minusS_Aarabi2016.fits.summary.AGI.FC.csv.gz
left_join(custom.categoreis.map[["minusS_Aarabi2016.fits.summary.AGI.FC.csv.gz"]] %>% dplyr::select(AGI,mean.FC.AGI,`FDR_trt-S`), by="AGI") %>% dplyr::rename(!!c(`minusS_\nAarabi2016_\nroot`="mean.FC.AGI",`minusS_Aarabi2016_root_FDR`="FDR_trt-S"))  %>%
  # minusN Peng 2007 whole plant
left_join(custom.categoreis.map[["minusN_Peng_2007.AGI.FC.csv.gz"]], by="AGI") %>% dplyr::rename(!!c(`minusN_\nPeng_2007_\nwhole`="logFC")) %>%
  # plusIAA
left_join(custom.categoreis.map[["plusIAA_Nemhauser_2006_seedlings.AGI.FC.csv.gz"]], by="AGI") %>% dplyr::rename(!!c(`plusIAA_\nNemhauser_2006_\nseedlings`="logFC")) %>%
## Needs to add Nishida_2017 data (062620)
left_join(custom.categoreis.map[["minusB_Nishida_2017_root.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>%
  dplyr::rename(!!c(`minusB_\nNishida_2017_\nroot`="logFC",`minusB_\nNishida_2017_\nroot_FDR`="FDR")) %>% 
  #   minusCa_Nishida_2017_root.DEGs.csv.gz
  left_join(custom.categoreis.map[["minusCa_Nishida_2017_root.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>%
  dplyr::rename(!!c(`minusCa_\nNishida_2017_\nroot`="logFC",`minusCa_\nNishida_2017_\nroot_FDR`="FDR")) %>%
# minusCu_Nishida_2017_root.DEGs.csv.gz
  left_join(custom.categoreis.map[["minusCu_Nishida_2017_root.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>%
  dplyr::rename(!!c(`minusCu_\nNishida_2017_\nroot`="logFC",`minusCu_\nNishida_2017_\nroot_FDR`="FDR")) %>%
# minusFe_Nishida_2017_root.DEGs.csv.gz
  left_join(custom.categoreis.map[["minusFe_Nishida_2017_root.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>%
  dplyr::rename(!!c(`minusFe_\nNishida_2017_\nroot`="logFC",`minusFe_\nNishida_2017_\nroot_FDR`="FDR")) %>%
# minusK_Nishida_2017_root.DEGs.csv.gz
    left_join(custom.categoreis.map[["minusK_Nishida_2017_root.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>%
  dplyr::rename(!!c(`minusK_\nNishida_2017_\nroot`="logFC",`minusK_\nNishida_2017_\nroot_FDR`="FDR")) %>%
# minusMg_Nishida_2017_root.DEGs.csv.gz
  left_join(custom.categoreis.map[["minusMg_Nishida_2017_root.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>%
  dplyr::rename(!!c(`minusMg_\nNishida_2017_\nroot`="logFC",`minusMg_\nNishida_2017_\nroot_FDR`="FDR")) %>%
# minusMn_Nishida_2017_root.DEGs.csv.gz
  left_join(custom.categoreis.map[["minusMn_Nishida_2017_root.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>%
  dplyr::rename(!!c(`minusMn_\nNishida_2017_\nroot`="logFC",`minusMn_\nNishida_2017_\nroot_FDR`="FDR")) %>%
# minusN_Nishida_2017_root.DEGs.csv.gz
  left_join(custom.categoreis.map[["minusN_Nishida_2017_root.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>%
  dplyr::rename(!!c(`minusN_\nNishida_2017_\nroot`="logFC",`minusN_\nNishida_2017_\nroot_FDR`="FDR")) %>%
# minusP_Nishida_2017_root.DEGs.csv.gz
  left_join(custom.categoreis.map[["minusP_Nishida_2017_root.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>%
  dplyr::rename(!!c(`minusP_\nNishida_2017_\nroot`="logFC",`minusP_\nNishida_2017_\nroot_FDR`="FDR")) %>%
# minusS_Nishida_2017_root.DEGs.csv.gz
  left_join(custom.categoreis.map[["minusS_Nishida_2017_root.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>%
  dplyr::rename(!!c(`minusS_\nNishida_2017_\nroot`="logFC",`minusS_\nNishida_2017_\nroot_FDR`="FDR")) %>%
# minusZn_Nishida_2017_root.DEGs.csv.gz
left_join(custom.categoreis.map[["minusZn_Nishida_2017_root.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>%
  dplyr::rename(!!c(`minusZn_\nNishida_2017_\nroot`="logFC",`minusZn_\nNishida_2017_\nroot_FDR`="FDR")) %>%
  dplyr::rename(genes=name)
colnames(all.data)
```
# 
what does the data sheet look like?  
```{r}
all.data
summary(all.data)
table(is.na(all.data$name)) # all FALSE
# When Nishida_2017 data have been added run next line. v5 (080621)
write_csv(all.data, file=file.path("annotation","output","root.all.comparison.v5.csv.gz")) # use this for making "Supplemental_Table4_custom_categories.xlsx"
```

# shoot
```{r, message=FALSE, warning=FALSE, error=TRUE}
all.data <-
  # shoot.exp1and3.FDR0.1 vs Al_Ligaba-OSena_2017_shoot
  #leaf.exp1plus3.all %>%
  Br.v3.0anno.At.BLAST.highscore %>% left_join(custom.categoreis.map[["plusAl_Ligaba-OSena_2017_shoot.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI")  %>% dplyr::rename(!!c(`plusAl_\nLigaba-OSena_2017_\nshoot`="logFC",`plusAl_Ligaba-OSena_2017_shoot_FDR`="FDR"))  %>%
  # root.exp1and3.FDR0.1 vs cold.root.Kilian_2013 (lmer)
  left_join(custom.categoreis.map[["cold_Kilian_2007.shoot.fits.summary.AGI.FC.csv.gz"]] %>% dplyr::select(AGI,mean.logFC.AGI), by="AGI") %>% 
  dplyr::rename(`cold\nKilian_2013_\nshoot`="mean.logFC.AGI") %>% 
  # # shoot.exp1and3.FDR0.1 vs Fe_deficiency.Kailasam2019
  # left_join(custom.categoreis.map[["minusFe_Kailasam2019.DEGs.all.anno.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR) , by="AGI") %>% 
  # dplyr::rename(!!c(`minusFe_\nKailasam2019_\nshoot`="logFC",`minusFe_Kailasam2019_shoot_FDR`="FDR")) %>%
  # shoot.exp1and3.FDR0.1 vs Fe_deficiency.Kailasam2019
  left_join(custom.categoreis.map[["minusFe_Kim2019.fits.summary.AGI.FC.WT.csv.gz"]] %>% dplyr::select(AGI,mean.FC.AGI,`FDR_trtFe-`), by="AGI") %>%
  dplyr::rename(!!c(`minusFe_\nKim2019_\nshoot`="mean.FC.AGI",`minusFe_Kim2019_shoot_FDR`="FDR_trtFe-")) %>%
  # shoot.exp1and3.FDR0.1 vs heat.shoot.Kilian_2013.AGI.exp.value.FC.csv.gz (lmer)
  left_join(custom.categoreis.map[["Heat_Kilian_2007.shoot.fits.summary.AGI.FC.csv.gz"]] %>% dplyr::select(AGI,mean.logFC.AGI), by="AGI") %>% 
  dplyr::rename(`heat_\nKilian_2013_\nshoot`="mean.logFC.AGI") %>%
  # shoot.exp1and3.FDR0.1 vs Mg_high_Niu_2016_shoot.DEGs.csv.gz
  left_join(custom.categoreis.map[["plusMg_Niu_2016_shoot.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>% dplyr::rename(!!c(`plusMg_\nNiu_2016_\nshoot`="logFC",`plusMg_Niu_2016_shoot_FDR`="FDR")) %>%
  # shoot.exp1and3.FDR0.1 vs Mg_low_Niu_2016_shoot.DEGs.csv.gz
  left_join(custom.categoreis.map[["minusMg_Niu_2016_shoot.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>%
  dplyr::rename(!!c(`minusMg_\nNiu_2016_\nshoot`="logFC",`minusMg_Niu_2016_shoot_FDR`="FDR")) %>% #View()
  # shoot.exp1and3.FDR0.1 vs minusPi_1d_Liu_2016_shoot.DEGs.csv.gz
  left_join(custom.categoreis.map[["minusPi_1d_Liu_2016_shoot.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR) , by="AGI") %>%
  dplyr::rename(!!c(`minusPi_1d_\nLiu_2016_\nshoot`="logFC",`minusPi_1d_Liu_2016_shoot_FDR`="FDR")) %>% #View()
  # shoot.exp1and3.FDR0.1 vs minusPi_3d_Liu_2016_shoot.DEGs.csv.gz
  left_join(custom.categoreis.map[["minusPi_3d_Liu_2016_shoot.DEGs.csv.gz"]] %>% dplyr::select(AGI,logFC,FDR), by="AGI") %>%
  dplyr::rename(!!c(`minusPi_3d_\nLiu_2016_\nshoot`="logFC",`minusPi_3d_Liu_2016_shoot_FDR`="FDR")) %>% rename(genes=name) #%>% #View()
  # shoot.exp1and3.FDR0.1 vs N_Wang2013.shoot.fits.summary.AGI.FC.csv.gz
  # full_join(custom.categoreis.map[["plusN_Wang2003.shoot.fits.summary.AGI.FC.csv.gz"]] %>%
  #             dplyr::select(AGI,mean.FC.AGI,FDR_trtKCl) , by="AGI") %>%
  # dplyr::rename(!!c(`plusN_\nWang2003_\nshoot`="mean.FC.AGI",`plusN_Wang2003_shoot_FDR`="FDR_trtKCl")) %>% #View()
  # shoot.exp1and3.FDR0.1 vs S_Aarabi2016.fits.summary.AGI.FC.csv.gz (tissu from root)
  # left_join(custom.categoreis.map[["minusS_Aarabi2016.fits.summary.AGI.FC.csv.gz"]] %>% dplyr::select(AGI,mean.FC.AGI,`FDR_trt-S`) , by="AGI") %>% dplyr::rename(!!c(`minusS_\nAarabi2016_\nshoot`="mean.FC.AGI",`minusS_Aarabi2016_shoot_FDR`="FDR_trt-S")) 

colnames(all.data)
```

# 
what does the data sheet look like?  
```{r}
all.data
summary(all.data)
table(is.na(all.data$genes)) # TRUE 19741 ???? why? (042720) -> fix full_join to left_join (042720)
#
write_csv(all.data, file=file.path("annotation","output","shoot.all.comparison.v5.csv.gz")) # use this for making "Supplemental_Table4_custom_categories.xlsx"
```
