---
title: "custom_categories_consolidate1_forV5"
author: "Kazu"
date: "5/9/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

Summary

History
* copied from "06_f_cor_table.Rmd"
* add GSEA_cust_rootshoot_updown3.FDR0.001.list (051221)
*  Expand DEGs with FC threshold for all custom categoreis
* root data including minusN_Peng_2007 and plusIAA_Nembauser_2006
* adding Nishida_2017 to root categories (070720)
* making top20, top50, and top100 for (a) among root/leaf categories, (b) among minusX_up, (c) among minusX_down
* More stringent FDR for custom category ORA of root.ivory module (050921)
* "In only" approach for custom ORA (051821)
* Fair "In only" approach for custom ORA using consolidated categoreis (052821)
* Using consolidated categories in InX selection process and ORA and heatmap with the consolidated categories 
* Another version (consolidate2) is using the consolidated categoreis to generate InX, but use the original categories for ORA.
* plot heatmap of InX comparison for root.ivory module
* and more modules.
* Making v4 custom categories and use it for this script
* Modify for updated plusN_Wang2003 using v5 (080621-081021)
* Save as "custom_categories_consolidate1_forV5"
* heatmap coloaring by soil FC sign (positive vs negative) (Sep 2021)
* Cleaning up scripts (Jan 25, 2021)
* split scripts into two (Jan 27, 2021)

To Do
* Cleaning up scripts more
* having the same x-axis and y-axis for WGCNA heatmap

#### making root + shoot category with up and down and
# prep
```{r}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE,error=TRUE)
library(tidyverse);library(GGally);library(readxl)
library(readr)
library(stringr)
# https://bioconductor.org/packages/devel/workflows/vignettes/maEndToEnd/inst/doc/MA-Workflow.html
# functions for expression pattern
#source("../../../tools/Expression_pattern_graph.R",chdir=TRUE)
library(scales) # for muted function
library(ggdendro) # for dendrogram
library(cowplot)
library(ggnewscale)
```

# read custom gene categories created in "06_d_cor_permutation.Rmd" -> needs to add this to this scripts -> use updated version (v4; 06**21) made by "custom_categories_v4_making.Rmd" -> use v5 made by custom_categories_v5_making.Rmd
```{r}
root.all5 <- read_csv(file.path("annotation","output","root.all.comparison.v5.csv.gz")) 
shoot.all5 <- read_csv(file.path("annotation","output","shoot.all.comparison.v5.csv.gz"))

```

fix names (root)
```{r}
root.all.fix.names <- root.all5
colnames(root.all.fix.names) <- ifelse(str_detect(colnames(root.all.fix.names), "(shoot|root|seedlings|whole)$"),
       str_c(colnames(root.all.fix.names), "_FC"),
       colnames(root.all.fix.names))
colnames(root.all.fix.names) <- str_remove_all(colnames(root.all.fix.names), fixed("\n") )
root.all.fix.names
```

fix names (shoot)
```{r}
shoot.all.fix.names <- shoot.all5
colnames(shoot.all.fix.names) <- ifelse(str_detect(colnames(shoot.all.fix.names), "(shoot|root)$"),
       str_c(colnames(shoot.all.fix.names), "_FC"),
       colnames(shoot.all.fix.names))
colnames(shoot.all.fix.names) <- str_remove_all(colnames(shoot.all.fix.names), fixed("\n") )
shoot.all.fix.names
```

# category (root and shoot combined): bugfixed (052620)
# FDR_cutoff <- 0.001 (053021)
# bug fixed by removng duplicated gene. 
```{r}
shoot.all.fix.names <- shoot.all.fix.names %>% dplyr::select(-AGI, -At_symbol, -At_short_description)  
colnames(shoot.all.fix.names)
# combine two columns (At_symbol from root and shoot) (work later)
custom.data <- root.all.fix.names %>% full_join(shoot.all.fix.names,by="genes")
colnames(custom.data)
custom.FC.names <- str_subset(colnames(custom.data), "FC")
custom.FC.names
# write csv file for filtered.custom.data to share with team members
write_csv(custom.data,file.path("annotation","output","custom_categories.csv.gz"))
# filter custom.data by FC and FDR cutoff
custom.data <- read_csv(file.path("annotation","output","custom_categories.csv.gz"))
FC_cutoff <- 0.58 # log2
#FDR_cutoff <- 0.01
FDR_cutoff <- 0.001
filtered.custom.data <- custom.data
### under construction
# needs to remove column that has only 1 in FDR, which was treated as TRUE/FALSE when reading csv.

for(FC in custom.FC.names) {
  FDR <- str_c(str_remove_all(FC, "_FC"), "_FDR")
  print(FDR)

  FC_up <- str_c(FC,"_up") %>% as.name()
  FC_up <- enquo(FC_up)
  FC_down <- str_c(FC,"_down") %>% as.name()
  FC_down <- enquo(FC_down)
  
  FDR_up <- str_c(FDR,"_up") %>% as.name()
  FDR_up <- enquo(FDR_up)
  FDR_down <- str_c(FDR,"_down") %>% as.name()
  FDR_down <- enquo(FDR_down)

  FC <- as.name(FC)
  FC <- enquo(FC)
  print(FC)
  FDR<- as.name(FDR)
  FDR <- enquo(FDR)

  filtered.custom.data <- filtered.custom.data %>%
    mutate(!!FC_up := ifelse( !! FC > FC_cutoff ,  !!FC, NA))  %>%
    mutate(!!FC_down := ifelse( !! FC < (-1)*FC_cutoff ,  !!FC, NA))

  if(!str_detect(rlang::as_name(FC), "cold|heat|Peng|Nemhauser")) {  # no FDR for cold/heat/Peng/Nemhauser
    print(FC)
    print("not cold")
    filtered.custom.data <- filtered.custom.data %>%
    mutate(!!FDR_up := ifelse( !! FC >FC_cutoff, !!FDR, NA)) %>% # cut FDR
    mutate(!!FDR_down := ifelse( !! FC < (-1)*FC_cutoff, !!FDR, NA)) %>%  
    mutate(!!FDR_up := ifelse( !! FDR_up < FDR_cutoff, !!FDR_up, NA)) %>%
    mutate(!!FDR_down := ifelse( !! FDR_down < FDR_cutoff, !!FDR_down, NA)) %>%
      mutate(!!FC_up := ifelse( !! FDR_up < FDR_cutoff ,  !!FC, NA )) %>%
     mutate(!!FC_down := ifelse( !! FDR_down < FDR_cutoff ,  !!FC, NA)) 
  } #if
  
} #for

summary(!is.na(filtered.custom.data)) 
names(filtered.custom.data)
# remove 
filtered.custom.data %>% dplyr::select(minusZn_Nishida_2017_root_FDR_down) # all NA

# manually make csv file and save it as txt for gmx file
#write_csv(filtered.custom.data,file=file.path("annotation","output","filtered.custom.data.rootshoot.csv.gz")) # 052620. Note this is an intermediate data. "filtered.custom.data.Total.rootshoot.v3.gmx" created below was used for further analysis (050921)
# There is type problems in reading the csv

```

# formatting custom categories for GOseq (Sep 29, 2020) (run once)
# updated (062221)
```{r eval=FALSE}
#filtered.custom.data <- read_csv(file.path("annotation","output","filtered.custom.data.rootshoot.csv.gz"),guess_max = 3000)
# howmany genes in the list?
summary(!is.na(filtered.custom.data))
# how many categories?
TF <- names(filtered.custom.data) %>% str_detect(c("plus","minus"))
categories.up.down <- names(filtered.custom.data)[TF] %>% gsub("_FC","",.) %>% gsub("_FDR","",.) %>% unique() # 58 categories
categories.up.down.combined <- categories.up.down %>% gsub("_up","",.) %>% gsub("_down","",.) %>% unique() # 30 categories (responsive genes= "up" or "down")
```

# convert into list object ("value" is fold change): one time (that is why echo=FALSE)
```{r echo=FALSE}
 cust_rootshoot_updown2 <- filtered.custom.data %>%
  dplyr::select(genes, ends_with("up"), ends_with("down"), -starts_with("exp")) %>%
  pivot_longer(-genes) 
cust_rootshoot_updown5.consolidated.FDR0.001.list <- cust_rootshoot_updown2 %>% drop_na(value) %>% filter(str_detect(name,"FC")) %>% dplyr::select(-value) %>%split(.$genes) %>%   map(~dplyr::select(.,name)) %>% map(~as_vector(.))
save(cust_rootshoot_updown5.consolidated.FDR0.001.list,file=file.path("annotation","output","cust_rootshoot_updown5.consolidated.FDR0.001.list.Rdata"))
```

# load consolidated custom categories (start here? necessary?)
```{r}
load(file.path("annotation","output","cust_rootshoot_updown5.consolidated.FDR0.001.list.Rdata"))
```

# Find unique genes among all custom categories and remove them by calculating majority roles (root and shoot categories)
```{r}
filtered.custom.data # Needs 
# library(tidyverse) 
# tentatively stop reading the csv file and create custom.data from the beggining.
#x <- file.path("annotation","output","filtered.custom.data.rootshoot.csv.gz")
#filtered.custom.data <- read_csv(x,guess_max = 17000)
# 
#problems(x)
#spec_csv(x)
# convert into 0/1 data.frame
filtered.custom.data.TF <- filtered.custom.data %>% dplyr::select(matches("FC_up|FC_down")) %>% is.na() 
summary(!filtered.custom.data.TF)  
# invert TRUE/FALSE
filtered.custom.data.TF2 <- !filtered.custom.data.TF 
# conver into 1 or 0
filtered.custom.data.TF3 <- filtered.custom.data.TF2*1
summary(filtered.custom.data.TF3==1)
# 
filtered.custom.data.TF4 <- bind_cols(filtered.custom.data[,"genes"],as_tibble(filtered.custom.data.TF3))
```

# consolidate categories
```{r}
filtered.custom.data.TF5 <- filtered.custom.data.TF4 %>% pivot_longer(cols=-genes,names_to="category")  %>% mutate(category=str_replace_all(category,"Kailasam2019","Kailasam_2019")) %>% mutate(category=str_replace_all(category,"minusPi_1d","minusPi")) %>% mutate(category=str_replace_all(category,"minusPi_3d","minusPi")) %>% mutate(category=str_replace_all(category,"coldKilian","cold_Kilian")) %>% mutate(category=str_replace_all(category,"Rodriguez-Celma","RodriguezCelma")) %>% mutate(category=str_replace_all(category,"Kim2019","Kim_2019")) %>% mutate(category=str_replace_all(category,"Wang2003","Wang_2003")) %>% mutate(category=str_replace_all(category,"Aarabi2016","Aarabi_2016")) %>%  mutate(category=str_replace_all(category,"Ligaba-OSena","LigabaOSena")) %>% separate(category,c("category2","au","yr","tissue","FC","up_down"),remove=FALSE) #%>% View()

filtered.custom.data.TF5 %>% dplyr::count(genes,category2,au,yr,tissue,up_down) %>% View()
# consolidate categories (value is "0" (absent) or "1" present in each category)
filtered.custom.data.consolidated <- filtered.custom.data.TF5 %>% filter(value==1) %>% dplyr::count(genes,category2,tissue,FC,up_down) %>% unite(category.cons,category2,tissue,FC,up_down) #%>% View()
#
filtered.custom.data.consolidated %>% View()
# how many genes in each category?
filtered.custom.data.consolidated %>% dplyr::select(-n) %>% dplyr::count(category.cons) %>% View()
# save filtered.custom.data.consolidated (Jan 25, 2021)
save(filtered.custom.data.consolidated,file=file.path("annotation","output","filtered.custom.data.consolidated.Rdata"))
#
load(file.path("annotation","output","filtered.custom.data.consolidated.Rdata"))

# make list
filtered.custom.data.consolidated.list <- filtered.custom.data.consolidated  %>% dplyr::select(-n) %>% split(.$genes) %>%   map(~dplyr::select(.,category.cons)) %>% map(~as_vector(.))
save(filtered.custom.data.consolidated.list,file=file.path("annotation","output","cust_rootshoot_updown5.consolidated.FDR0.001.list.Rdata")) 
```

# convert filtered.custom.data.consolidated into filtered.custom.data.numeric format (does not work)
# the final goal is to make list object for GOseq ORA with InX.con.cat (=consolidated category) like In01.forORA.list2 
```{r}
filtered.custom.data.consolidated.numeric <- filtered.custom.data.consolidated %>% pivot_wider(names_from=category.cons, values_from = n) %>% 
  mutate(across(-genes, ~ ifelse(is.na(.), 0, 1))) #%>%View() # why
```

```{r}
filtered.custom.data.consolidated %>% dplyr::select(-n) %>% dplyr::count(genes) %>% View()
```

# below is Julin's scripts for making 

Alternate plotting
```{r}
apply(filtered.custom.data.TF3, 1, sum) %>% hist()
```

How many left after we remove...

redo the conversion so that we keep gene names
```{r}
#filtered.custom.data.numeric <- filtered.custom.data %>%
#  dplyr::select(genes, matches("FC_up|FC_down")) %>%
#  mutate_at(vars(-genes), ~ as.numeric(!is.na(.)))
```

for each gene total the number of sets that it is in
```{r}
filtered.custom.data.numeric  # original. no more exist
filtered.custom.data.consolidated.numeric
filtered.custom.data.consolidated.numeric <- filtered.custom.data.consolidated.numeric %>%   mutate(totalSetN=rowSums(filtered.custom.data.consolidated.numeric[,-1])) %>%
  dplyr::select(genes, totalSetN, everything())
head(filtered.custom.data.consolidated.numeric)
```

Now we want to know if we filter by X totalSetN, how many genes are left in each gene set.

```{r}
filtered.custom.data.consolidated.numericL <-
  filtered.custom.data.consolidated.numeric %>%
  pivot_longer(-c("genes", "totalSetN"), names_to="set", values_to = "presentInSet")
head(filtered.custom.data.consolidated.numericL)
```

# new filtered.custom.data.consolidated.numericL with "XorMore" (061321)
```{r}
filtered.custom.data.consolidated.numericL <- filtered.custom.data.consolidated.numericL %>%
  mutate(In01 = presentInSet*(totalSetN <= 1), #would be nice to automate this...
         In02orLess = presentInSet*(totalSetN <= 2),
         In03orLess = presentInSet*(totalSetN <= 3),
         In05orLess = presentInSet*(totalSetN <= 5),
         In10orLess = presentInSet*(totalSetN <= 10),
         In20orLess = presentInSet*(totalSetN <= 20),
         In05orMore = presentInSet*(totalSetN >= 5),
         In10orMore = presentInSet*(totalSetN >= 10),
         In20orMore = presentInSet*(totalSetN >= 20),
         Total = presentInSet)

filtered.custom.data.consolidated.numericL
save(filtered.custom.data.consolidated.numericL,file=file.path("RNA","output","filtered.custom.v5.data.consolidated.numericL.Rdata"))
```

# Something is wrong with making "filtered.custom.data.consolidated.numericLL" below (chunk 17)? (061821, Kazu): Start here? (Dec 21, 2021)
```{r}
load(file.path("RNA","output","filtered.custom.v5.data.consolidated.numericL.Rdata"))

filtered.custom.data.consolidated.numericLL <- filtered.custom.data.consolidated.numericL %>%
  pivot_longer(c(starts_with("In"),starts_with("Total", ignore.case = FALSE)), names_to = "group", values_to = "group_summary")
head(filtered.custom.data.consolidated.numericLL,20)
```

now summarize it
```{r}
filtered.custom.data.consolidated.summary <- filtered.custom.data.consolidated.numericLL %>%
  group_by(set, group) %>%
  summarize(genesInSet=sum(group_summary))
filtered.custom.data.consolidated.summary
```

plot it
```{r}
filtered.custom.data.consolidated.summary %>%
  ggplot(aes(x=set, y=genesInSet, color=group)) +
  geom_point()  +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

plot it percentage (needs to facet_grid with Xorless and Xormore)
```{r}
InSet.percent.plot <- filtered.custom.data.consolidated.summary %>%
  group_by(set) %>%
  mutate(percent_of_set=100*genesInSet/max(genesInSet)) %>%
    ggplot(aes(x=set, y=percent_of_set, color=group)) +
  geom_point()  +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
ggsave(InSet.percent.plot,path=file.path("annotation","output"),filename="InSet.percent.plot.consolidation.png",width=8,height=6)
```

