---
title: "N_deficiency (minusN) responsive_genes"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

Purpose:
* Analize ATH1 microarray raw data () by lmer and broom.mixed packages for tidy way. 

|our purpose|Study|plant|samples|transcriptome platform|available data|available raw data|
|----|----|--|------|--|--------|-|
|N starvation|[Bi et al. 2007](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1994689)|
Arabidopsis thaliana|Shoots of 3-week-old wild-type plants.  There were three nitrate levels (treatments) in the experiment with the 3 mM nitrate as sufficient, 1 mM nitrate as the mild N-limiting condition and 0.3 mM nitrate as the severe N-limiting condition|Microarray (ATH1 chip)|FoldChange of some genes are listed in tables but not all genes.  All DGE listed in supplementary excel sheet but not FoldChange.  May not be useful...|[NASCARRAYS-408](http://bar.utoronto.ca/NASCArrays/index.php?ExpID=408)|

To Do
* This is not ATH1. Stop working on it.
* Expression pattern graph
* Knit to finalize analysis.

# installation
```{r eval=FALSE, include=FALSE}
install.packages("tidyverse")
install.packages("broom")
install.packages("broom.mixed")
install.packages("lmerTest")
install.packages("rmarkdown")
# https://bioconductor.org/packages/devel/workflows/vignettes/maEndToEnd/inst/doc/MA-Workflow.html
install.packages("devtools")
library(devtools)
devtools::install_github("r-lib/remotes")
#library(remotes)
#packageVersion("remotes") # has to be 1.1.1.9000 or later
remotes::install_github("b-klaus/maEndToEnd", ref="master")
```

# prep
```{r setup, include=FALSE}
library(lmerTest);library(broom);library(broom.mixed);library(tidyverse);library(lme4);library(affy)
# https://bioconductor.org/packages/devel/workflows/vignettes/maEndToEnd/inst/doc/MA-Workflow.html
source("Expression_pattern_graph.R")
```
# Read cel files and normalize data (copied from "RNAseqVSmicroarray_publication_final.R")
```{r eval=TRUE}
cel.filename<-list.files(pattern=".cel$") # no directory info
cel.filename2<-list.files(pattern=".cel",path=file.path("","Volumes","data_work","Data8","Microarray","minusN_Bi_2007","Exp408"),full.names=TRUE) # "" is needed for file.path()!!!  with directory info
Data1<-ReadAffy(filenames =cel.filename2) 
head(Data1)
eset <- rma(Data1) # normalize chips (old way)
# using expresso (see affy vignett)
 eset.expresso<-expresso(Data1,normalize.method="qspline",
               bgcorrect.method="rma",pmcorrect.method="pmonly",
               summary.method="liwong")
 class(eset.expresso) # ExpressionSet object. See Biobase vignette
 head(eset.expresso)
 exprs(eset.expresso) %>% head()# expression data
 experimentData(eset.expresso)
```
# read probe ID and AGI
```{r eval=TRUE}
affyData <- read_csv(file.path("..","Annotation","input","affy_ATH1_array_elements-2010-12-20.csv")) #read TAIR annotation data from file (TAIR10)
affyData2<-affyData[,c("array_element_name","is_control","locus","description","chromosome","start","stop")]
names(affyData2) <- c("element","is.control","AGI","description","chromosome","start","stop")#get rid of underlines
head(affyData2)
# only unique element per AGI (is this OK? ASk Julin)
affyData2.elementPerAGI.table<- affyData2 %>% group_by(AGI) %>% summarize(n=n())
table(affyData2.elementPerAGI.table$n)
affyData2unique<-affyData2.elementPerAGI.table %>% filter(n<2) %>% left_join(affyData2,by="AGI")
dim(affyData2unique) #[1] 21789     8 use this instead of affyData2
dim(affyData2) #[1] 22810     7
```


# Make expression value tibble
```{r eval=TRUE}
featureNames(eset.expresso)  # only probe name
exp.values<-exprs(eset.expresso) %>% as_tibble(rownames="element") 
```

# experimental design
```{r eval=TRUE}
# extract experiment data info
cel.filename<-list.files(pattern=".CEL.gz",path=file.path("","Volumes","data_work","Data8","Microarray","Bi_2007_minusN","Exp408"),full.names=FALSE) # "" is needed for file.path()!!!  with directory info


# design (under construction)
design <- tibble(cel_file=cel.filename) %>% mutate(cel_file2=str_replace(cel_file,"Wild-type","WT")) %>% separate(cel_file2,into=c("NRID","Wang","num","condition","rep","extra"),extra="drop",sep="_") %>% mutate(condition=str_replace(condition,"K-nitrate","KNO3"),condition=str_replace(condition,"KCL","KCl")) %>% 
  separate(condition,into=c("gt1","gt2","conc","unit","trt","time","unit2","tissue")) %>% mutate(rep=str_replace(rep,"Rep",""),tissue=str_replace(tissue,"s$","")) %>% dplyr::select(cel_file,gt1,trt,tissue,rep)
#
design
# root or shoot
design.root <- design %>% filter(tissue=="root") 
design.shoot <- design %>% filter(tissue=="shoot") 
```

# combine
```{r eval=TRUE}
exp.values.combined <- exp.values %>% gather("cel_file","value",-element) %>% left_join(design,by="cel_file") %>% 
  left_join(affyData2, by="element") 
# check 10 genes
exp.values.combined %>% arrange(AGI) %>% head(10)
# factor
exp.values.combined %>% mutate(trt=factor(trt, levels=c("KNO3","KCl")),rep=as.character(rep)) %>% str() # N deficiency
# write in csv file
write_csv(exp.values.combined,path=file.path("intermediate","minusN_Bi_2007","exp.values.combined.csv.gz"))
```

# lmer function
```{r eval=TRUE}
dofit.lmer.ATH1 <- function(genedata) { # gives a lot of singularity warnings
  lmer(value ~ trt + (1|rep),
       data=genedata)
}
```

# root
# organize the data to make it easy to fit a model to each element (not gene)
```{r eval=TRUE}
exp.values.combined <- read_csv(file.path("intermediate","minusN_Bi_2007","exp.values.combined.csv.gz"))
exp.values.combined.root.nest <- exp.values.combined %>% filter(tissue=="root") %>% select(value,trt,rep,element) %>%
  mutate(trt=factor(trt,levels=c("KNO3","KCl"))) %>%
  group_by(element) %>%
  nest()
head(exp.values.combined.root.nest)
exp.values.combined.root.nest %>% summarize(num=n()) %>% arrange(num) %>% View() #OK
exp.values.combined.root.nest$data[[1]] %>% str() # factor...
```

# do a fit for each element
```{r eval=TRUE, warning=FALSE}
root.fits <- exp.values.combined.root.nest %>% #head(100) %>% 
  #magrittr::extract(1:10,) %>% # comment out this line to do full set; uncomment for testing on small set.
  mutate(fit.lmer.ATH1 = map(data, dofit.lmer.ATH1),
         lmersingular = map_lgl(fit.lmer.ATH1, isSingular))
head(root.fits)
```

# add model summaries (5-10 minutes) using broom package (read "introduction to broom" vegnet)
```{r eval=TRUE}
root.fits.summary <- root.fits %>%
  mutate(glancelmer = map(fit.lmer.ATH1, broom::glance),
         tidylmer = map(fit.lmer.ATH1, broom::tidy)
  ) # see tidy.lm
```

# unnesting by unnest() 
```{r}
root.fits.summary %>% unnest(tidylmer) %>% head(100) %>% View()
```

# make a 1 row summary of the tidy results (modified from copy from J's script "SingleGeneLME_DGE.Rmd")

```{r eval=TRUE}
root.fits.summary.table.all <- root.fits.summary %>% unnest(tidylmer, .drop = TRUE) %>%
  select(element, term, estimate, std.error, p.value) %>%
  na.omit() %>%
  mutate(p.value = ifelse(term=="(Intercept)", NA, p.value), # we don't care about intercept p-value and don't want it to mess up FDR
         FDR = p.adjust(p.value, "fdr")) %>%
  gather(key="key", value="value", estimate, std.error, p.value, FDR) %>%
  mutate(key=str_c(key, "_", term)) %>%
  select(-term) %>%
  spread(key=key, value=value) %>% 
  arrange(FDR_trtKCl) #%>% head(100) %>% View()
# how many DEGs?
root.fits.summary.table.all %>% filter(FDR_trtKCl < 0.05) %>% dim()
write_csv(root.fits.summary.table.all,path=file.path("custom_categories","minusN.Bi2007.root.fits.summary.table.all.csv.gz"))
```

# How to calculate logFC?
# read root.fits.summary.table.all 
```{r}
root.fits.summary.table.all<-read_csv(file.path("custom_categories","N.Wang2013.root.fits.summary.table.all.csv.gz"))
root.fits.summary.table.all
# 
# only this works
fits.summary.unnest<-root.fits.summary %>% 
  #head(10) %>% 
  mutate(predicted.values=map(fit.lmer.ATH1,predict)) %>% unnest(predicted.values) 
# adding samplename column
samplename<-design.root %>% unite(samplename,trt,rep,sep=".") %>% dplyr::select(samplename) %>% as_vector()
# calculate predicted expression values (used also for expression graph)
fits.element.exp.value<-fits.summary.unnest %>% ungroup() %>% inner_join(affyData2unique, by="element") %>% group_by(element) %>% mutate(samplename=samplename)  %>% 
  separate(samplename,c("trt","rep"),sep="\\.") %>% rename(value=predicted.values) %>% dplyr::select(element,value,AGI,description,trt,rep)
# save fits.element.exp.value
write_csv(fits.element.exp.value,file.path("custom_categories","N.Wang2013.root.fits.element.exp.value.csv.gz"))
# calculate mean value per element, treatment
fits.element.exp.value.mean <- fits.element.exp.value %>%
  group_by(element,trt) %>% 
  summarise(mean.element=mean(value)) %>% 
  spread(trt,mean.element,-1) %>%
  inner_join(affyData2unique, by="element") # only one element per one locus
#  check if "one" per one element
 fits.element.exp.value.mean %>% group_by(element) %>% summarize(num.element=n()) %>% arrange(num.element) %>% View()
 # calculate log fold changes per each locus (one locus = one element)
 fits.summary.AGI.FC <- fits.element.exp.value.mean %>% 
   group_by(AGI) %>% 
   mutate(mean.FC.AGI=log(KCl/KNO3)) %>%
  # add FDR_trtKCl
  inner_join(root.fits.summary.table.all 
            %>% dplyr::select(element,`FDR_trtKCl`), by="element") 
 fits.summary.AGI.FC %>% View()
 # check AGI number
 fits.summary.AGI.FC %>% group_by(AGI) %>% summarize(num.AGI=n()) %>% arrange(num.AGI) %>% View() # there is no  "no_match"
# write
write_csv(fits.summary.AGI.FC,path=file.path("custom_categories","N_Wang2013.root.fits.summary.AGI.FC.csv.gz"))
```

# shoot
# organize the data to make it easy to fit a model to each element (not gene)
```{r eval=TRUE}
exp.values.combined <- read_csv(file.path("intermediate","N_Wang_2013","exp.values.combined.csv.gz"))
exp.values.combined.shoot.nest <- exp.values.combined %>% filter(tissue=="shoot") %>% select(value,trt,rep,element) %>%
  mutate(trt=factor(trt,levels=c("KNO3","KCl"))) %>%
  group_by(element) %>%
  nest()
exp.values.combined.shoot.nest[1,2] %>% unnest(data)
exp.values.combined.shoot.nest %>% summarize(num=n()) %>% arrange(num) %>% View() #OK
exp.values.combined.shoot.nest$data[[1]] %>% str() # factor...
```

# do a fit for each element
```{r eval=TRUE, warning=FALSE}
shoot.fits <- exp.values.combined.shoot.nest %>% #head(100) %>% 
  #magrittr::extract(1:10,) %>% # comment out this line to do full set; uncomment for testing on small set.
  mutate(fit.lmer.ATH1 = map(data, dofit.lmer.ATH1),
         lmersingular = map_lgl(fit.lmer.ATH1, isSingular))
head(shoot.fits)
```

# add model summaries (5-10 minutes) using broom package (read "introduction to broom" vegnet)
```{r eval=TRUE}
shoot.fits.summary <- shoot.fits %>%
  mutate(glancelmer = map(fit.lmer.ATH1, broom::glance),
         tidylmer = map(fit.lmer.ATH1, broom::tidy)
  ) # see tidy.lm
```

# unnesting by unnest() 
```{r}
shoot.fits.summary %>% unnest(tidylmer) %>% head(100) %>% View()
```

# make a 1 row summary of the tidy results (modified from copy from J's script "SingleGeneLME_DGE.Rmd")

```{r eval=TRUE}
shoot.fits.summary.table.all <- shoot.fits.summary %>% unnest(tidylmer, .drop = TRUE) %>%
  select(element, term, estimate, std.error, p.value) %>%
  na.omit() %>%
  mutate(p.value = ifelse(term=="(Intercept)", NA, p.value), # we don't care about intercept p-value and don't want it to mess up FDR
         FDR = p.adjust(p.value, "fdr")) %>%
  gather(key="key", value="value", estimate, std.error, p.value, FDR) %>%
  mutate(key=str_c(key, "_", term)) %>%
  select(-term) %>%
  spread(key=key, value=value) %>% 
  arrange(FDR_trtKCl) #%>% head(100) %>% View()
# how many DEGs?
shoot.fits.summary.table.all %>% filter(FDR_trtKCl < 0.05) %>% dim()
write_csv(shoot.fits.summary.table.all,path=file.path("custom_categories","N.Wang2013.shoot.fits.summary.table.all.csv.gz"))
```

# How to calculate logFC?
# read shoot.fits.summary.table.all 
```{r}
shoot.fits.summary.table.all<-read_csv(file.path("custom_categories","N.Wang2013.shoot.fits.summary.table.all.csv.gz"))
shoot.fits.summary.table.all
# 
# only this works
fits.summary.unnest<-shoot.fits.summary %>% 
  #head(10) %>% 
  mutate(predicted.values=map(fit.lmer.ATH1,predict)) %>% unnest(predicted.values) 
# adding samplename column
samplename<-design.shoot %>% unite(samplename,trt,rep,sep=".") %>% dplyr::select(samplename) %>% as_vector()
# calculate predicted expression values (used also for expression graph)
fits.element.exp.value<-fits.summary.unnest %>% ungroup() %>% inner_join(affyData2unique, by="element") %>% group_by(element) %>% mutate(samplename=samplename)  %>% 
  separate(samplename,c("trt","rep"),sep="\\.") %>% rename(value=predicted.values) %>% dplyr::select(element,value,AGI,description,trt,rep)
write_csv(fits.element.exp.value,file.path("custom_categories","N.Wang2013.shoot.fits.element.exp.value.csv.gz"))
# calculate mean value per element, treatment
fits.element.exp.value.mean <- fits.element.exp.value %>%
  group_by(element,trt) %>% 
  summarise(mean.element=mean(value)) %>% 
  spread(trt,mean.element,-1) %>%
  inner_join(affyData2unique, by="element") # only one element per one locus
#  check if "one" per one element
 fits.element.exp.value.mean %>% group_by(element) %>% summarize(num.element=n()) %>% arrange(num.element) %>% View()
 # calculate log fold changes per each locus (one locus = one element)
 fits.summary.AGI.FC <- fits.element.exp.value.mean %>% 
   group_by(AGI) %>% 
   mutate(mean.FC.AGI=log(KCl/KNO3)) %>%
  # add FDR_trtKCl
  inner_join(shoot.fits.summary.table.all 
            %>% dplyr::select(element,`FDR_trtKCl`), by="element") 
 fits.summary.AGI.FC %>% View()
 # check AGI number
 fits.summary.AGI.FC %>% group_by(AGI) %>% summarize(num.AGI=n()) %>% arrange(num.AGI) %>% View() # there is no  "no_match"
# write
write_csv(fits.summary.AGI.FC,path=file.path("custom_categories","N_Wang2013.shoot.fits.summary.AGI.FC.csv.gz"))
```

# expression pattern (root)
```{r eval=TRUE}
fits.summary.AGI.FC <- read_csv(file.path("custom_categories","N_Wang2013.root.fits.summary.AGI.FC.csv.gz"))
fits.element.exp.value <- read_csv(file.path("custom_categories","N.Wang2013.root.fits.element.exp.value.csv.gz"))
# N_Wang2013 N deficienty
gene.of.interest.FDR.up <- fits.summary.AGI.FC %>% ungroup() %>% rename(FDR=`FDR_trtKCl`) %>% filter(FDR<0.05,mean.FC.AGI>0) %>% dplyr::select(AGI,FDR) %>% arrange(FDR) 
print("logFC>0")
p.up<-expression.pattern.AT.graph.ATH1(target.genes.FDR=gene.of.interest.FDR.up[1:5,],
                              data=fits.element.exp.value,
                              sample.description=design,
                              title="N_Wang2013 root -N up") # good
p.up
ggsave(p.up,filename=file.path("custom_categories","N_Wang2013_root.up.png"))
expression.down.plot <- "N_Wang2013_root.down.png"

# N_Wang2013 N deficienty down
gene.of.interest.FDR.down <- fits.summary.AGI.FC %>% ungroup() %>% rename(FDR=`FDR_trtKCl`) %>% filter(FDR<0.05,mean.FC.AGI<0) %>% dplyr::select(AGI,FDR) %>% arrange(FDR) 
print("logFC<0")
p.down<-expression.pattern.AT.graph.ATH1(target.genes.FDR=gene.of.interest.FDR.down[1:5,],
                              data=fits.element.exp.value,
                              sample.description=design,
                              title="N_Wang2013 root -N down") # good
p.down
ggsave(p.down,filename=file.path("custom_categories","N_Wang2013_root.down.png"))

```

# expression pattern (shoot)
```{r eval=TRUE}
fits.summary.AGI.FC <- read_csv(file.path("custom_categories","N_Wang2013.shoot.fits.summary.AGI.FC.csv.gz"))
fits.element.exp.value <- read_csv(file.path("custom_categories","N.Wang2013.shoot.fits.element.exp.value.csv.gz"))
# N_Wang2013 N deficienty
gene.of.interest.FDR.up <- fits.summary.AGI.FC %>% ungroup() %>% rename(FDR=`FDR_trtKCl`) %>% filter(FDR<0.05,mean.FC.AGI>0) %>% dplyr::select(AGI,FDR) %>% arrange(FDR) 
print("logFC>0")
p.up<-expression.pattern.AT.graph.ATH1(target.genes.FDR=gene.of.interest.FDR.up[1:5,],
                              data=fits.element.exp.value,
                              sample.description=design,
                              title="N_Wang2013 shoot -N up") # good
p.up
ggsave(p.up,filename=file.path("custom_categories","N_Wang2013_shoot.up.png"))
expression.down.plot <- "N_Wang2013_shoot.down.png"

# N_Wang2013 N deficienty down
gene.of.interest.FDR.down <- fits.summary.AGI.FC %>% ungroup() %>% rename(FDR=`FDR_trtKCl`) %>% filter(FDR<0.05,mean.FC.AGI<0) %>% dplyr::select(AGI,FDR) %>% arrange(FDR) 
print("logFC<0")
p.down<-expression.pattern.AT.graph.ATH1(target.genes.FDR=gene.of.interest.FDR.down[1:5,],
                              data=fits.element.exp.value,
                              sample.description=design,
                              title="N_Wang2013 shoot -N down") # good
p.down
ggsave(p.down,filename=file.path("custom_categories","N_Wang2013_shoot.down.png"))

```

# Session info
```{r}
sessionInfo()
```
