---
title: "Sulfer"
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: console
---

* To Do
  + cleanup scripts, especially correct directory names.
  + final check
  + Knit

Sulfer deficiency responsive genes.

|our purpose|Study|plant|samples|transcriptome platform|available data|available raw data|
|----|----|--|------|--|--------|-|
|Sulfer deficiency responsive genes|[Aarabi et al. 2016](https://advances.sciencemag.org/content/2/10/e1601087)|Arabidopsis thaliana|Col-0 (and some sulfur deficiency induced mutants SDI1 & SDI2) were raised on high S (1500uM SO4) or low S (15uM SO4).  Roots harvested after 10 days.|Microarray (ATH1). |Using (ANOVA), we selected 6000 genes as those that were differentially regulated among plant lines or conditions; many of the genes showed low P values negating problems caused by multiple testing (last stament sounds like BS).  Normalized expression data were converted through (PCA) to identify the directions of changes in gene expression at transcriptome levels shared with plant lines or conditions.  The direction indicated by PC1 showed high correlations with the –S responses in WT.   THESE PC1 GENES AND LOGRATIOS FOUND IN TABLE S6.  https://advances.sciencemag.org/highwire/filestream/191989/field_highwire_adjunct_files/1/1601087_Tables_S1_to_S11.zip|[GSE81347](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE81347)|

# installation
```{r eval=FALSE, include=FALSE}
install.packages("tidyverse")
install.packages("broom")
install.packages("broom.mixed")
install.packages("lmerTest")
install.packages("rmarkdown")
# https://bioconductor.org/packages/devel/workflows/vignettes/maEndToEnd/inst/doc/MA-Workflow.html
install.packages("devtools")
library(devtools)
devtools::install_github("r-lib/remotes")
#library(remotes)
#packageVersion("remotes") # has to be 1.1.1.9000 or later
remotes::install_github("b-klaus/maEndToEnd", ref="master")
```

# prep
```{r setup, include=FALSE}
#library(knitr)
knitr::opts_chunk$set(echo = TRUE,error=TRUE)
#library(d3heatmap)
#library(flexdashboard)
library(edgeR);library(tidyverse)
library(readr);library(readxl)
library(stringr)
# for microarray 
library(lmerTest);library(broom);library(broom.mixed);library(lme4);library(affy)
# https://bioconductor.org/packages/devel/workflows/vignettes/maEndToEnd/inst/doc/MA-Workflow.html
# for expression pattern
source("Expression_pattern_graph.R")
```

# Aarabi_2016. Microarray (ATH1)
Purpose:
* Analize ATH1 microarray raw data () by lmer and broom.mixed packages for tidy way. See "Single Gene LME" done by Julin.

# Read cel files and normalize data (copied from "RNAseqVSmicroarray_publication_final.R")
```{r eval=TRUE}
cel.filename<-list.files(pattern=".CEL.gz$","/Volumes/data_work/Data8/Microarray/GSE81347/") # no directory info
cel.filename2<-list.files(pattern=".CEL.gz$",path=file.path("","Volumes","data_work","Data8","Microarray","GSE81347"),full.names=TRUE) # "" is needed for file.path()!!!  with directory info
Data1<-ReadAffy(filenames =cel.filename2) 
head(Data1)
eset <- rma(Data1) # normalize chips (old way)
# using expresso (see affy vignett)
 eset.expresso<-expresso(Data1,normalize.method="qspline",
               bgcorrect.method="rma",pmcorrect.method="pmonly",
               summary.method="liwong")
 class(eset.expresso) # ExpressionSet object. See Biobase vignet
 head(eset.expresso)
 exprs(eset.expresso) %>% head() # expression data
 experimentData(eset.expresso) %>% head()
 # write text file (affy package vignett, 3.3) not used
# write.exprs(eset.expresso, file=file.path("..","output","eset.expresso.txt"))
```

# experimental design using GSE81347
```{r eval=TRUE}
cel.filename<-list.files(pattern=".CEL.gz$","/Volumes/data_work/Data8/Microarray/GSE81347") # no directory info
# GSM2150907_amaru912_-S_Col_1.CEL.gz: Col -S root Biological Rep1
# GSM2150908_amaru912_-S_Col_2.CEL.gz: Col -S root Biological Rep2
# GSM2150911_amaru912_+S_Col_1.CEL.gz: Col +S root Biological Rep1
# GSM2150912_amaru912_+S_Col_2.CEL.gz: Col +S root Biological Rep2
design<-tibble(cel_file=cel.filename) %>% separate(cel_file,c("title","title2","trt","gt","rep","extra"),sep="(\\_|\\.)",extra="drop",remove=FALSE) %>%
  dplyr::select(cel_file,trt,rep)
design
```

# read probe ID and AGI
```{r eval=TRUE}
affyData <- read_csv(file.path("..","Annotation","input","affy_ATH1_array_elements-2010-12-20.csv")) #read TAIR annotation data from file (TAIR10)
affyData2<-affyData[,c("array_element_name","is_control","locus","description","chromosome","start","stop")]
names(affyData2) <- c("element","is.control","AGI","description","chromosome","start","stop")#get rid of underlines
head(affyData2)
# only unique element per AGI (is this OK? ASk Julin)
affyData2.elementPerAGI.table<- affyData2 %>% group_by(AGI) %>% summarize(n=n())
table(affyData2.elementPerAGI.table$n)
affyData2unique<-affyData2.elementPerAGI.table %>% filter(n<2) %>% left_join(affyData2,by="AGI")
dim(affyData2unique) #[1] 21789     8 use this instead of affyData2
dim(affyData2) #[1] 22810     7
```

# Make expression value tibble
```{r eval=TRUE}
featureNames(eset.expresso) %>% head() # only probe name
exp.values<-exprs(eset.expresso) %>% as_tibble(rownames="element") 
```

# combine
```{r eval=TRUE}
exp.values.combined <- exp.values %>% gather("cel_file","value",-element) %>% left_join(design,by="cel_file") %>%
  left_join(affyData2unique, by="element") 
# check 10 genes
exp.values.combined %>% arrange(AGI) %>% head(10)
# factor
exp.values.combined %>% mutate(trt=factor(trt, levels=c("+S","-S")),rep=as.character(rep)) %>% str()
# write in csv file
write_csv(exp.values.combined,path=file.path("intermediate","Aarabi_2016","exp.values.combined.csv.gz"))
```

# lmer function
```{r eval=TRUE}
# gt*trt
dofit.lmer.ATH1 <- function(genedata) { # gives a lot of singularity warnings
  lmer(value ~ genotype*trt + (1|rep),
       data=genedata)
}
# for only trt (subset data for Col, one tissue type, and one treatment)
dofit.lmer2.ATH1 <- function(genedata) { # gives a lot of singularity warnings
  lmer(value ~ trt + (1|rep),
       data=genedata)
}
```

# organize the data to make it easy to fit a model to each element (not AGI)
```{r eval=TRUE}
exp.values.combined<-read_csv(file.path("intermediate","S_Aarabi_2016","exp.values.combined.csv.gz"))
exp.values.combined.nest <- exp.values.combined %>% 
  select(value,trt,rep,element) %>% 
  mutate(trt=factor(trt,levels=c("+S","-S"))) %>%
  group_by(element) %>%
  nest()
# check exp.values.combined.nest
exp.values.combined.nest %>% summarize(num=n()) %>% arrange(num) %>% head() #%>% View() #OK
exp.values.combined.nest$data[[1]] %>% str() # factor!

```

## do a fit for each gene.  
```{r eval=TRUE, warning=FALSE}
fits <- exp.values.combined.nest %>% #%>% head(100) %>% 
  #magrittr::extract(1:10,) %>% # comment out this line to do full set; uncomment for testing on small set.
  mutate(fit.lmer2.ATH1 = map(data, dofit.lmer2.ATH1),
         lmersingular = map_lgl(fit.lmer2.ATH1, isSingular))
head(fits)

```

## add model summaries (5-10 minutes) using broom package (read "introduction to broom" vegnet)
```{r eval=TRUE}
fits.summary <- fits %>%
  mutate(glancelmer = map(fit.lmer2.ATH1, broom::glance),
         tidylmer = map(fit.lmer2.ATH1, broom::tidy)
  ) # see tidy.lm
```

# unnesting by unnest() 
```{r}
fits.summary %>% unnest(tidylmer) %>% head(100) %>% View()
```

# make a 1 row summary of the tidy results (modified from copy from J's script "SingleGeneLME_DGE.Rmd")

```{r eval=TRUE}
fits.summary.table.all <- fits.summary %>% unnest(tidylmer) %>%
  select(element, term, estimate, std.error, p.value) %>%
  na.omit() %>%
  mutate(p.value = ifelse(term=="(Intercept)", NA, p.value), # we don't care about intercept p-value and don't want it to mess up FDR
         FDR = p.adjust(p.value, "fdr")) %>%
  gather(key="key", value="value", estimate, std.error, p.value, FDR) %>%
  mutate(key=str_c(key, "_", term)) %>%
  select(-term) %>%
  spread(key=key, value=value) %>% 
  arrange(`FDR_trt-S`) #%>% head(100) %>% View()
# number of DEGs
fits.summary.table.all %>% filter(`FDR_trt-S` < 0.05) %>% dim() # 3156 9
#
write_csv(fits.summary.table.all,path=file.path("intermediate","S_Aarabi_2016","S.Aarabi_2016.fits.summary.table.all.csv.gz"))
```
# DGE table
```{r}
fits.summary.table.all<-read_csv(file.path("intermediate","S_Aarabi_2016","S.Aarabi_2016.fits.summary.table.all.csv.gz"))
# number of DEGs
fits.summary.table.all %>% filter(`FDR_trt-S` < 0.05) %>% dim() # 
```

# How to calculate logFC?
* One option is using ?lme4::predict.merMod, which worked
* Another option is to add up estimate

```{r}
# 
fits.summary.table.all<-read_csv(file.path("intermediate","S_Aarabi_2016","S.Aarabi_2016.fits.summary.table.all.csv.gz"))
# only this works
fits.summary.unnest<-fits.summary %>% 
  #head(10) %>% 
  mutate(predicted.values=map(fit.lmer2.ATH1,predict)) %>% unnest(predicted.values) 
# adding samplename column
samplename<-design %>% unite(samplename,trt,rep,sep=".") %>% dplyr::select(samplename) %>% as_vector()
# calculate predicted expression values (used also for expression graph)
fits.element.exp.value<-fits.summary.unnest %>% ungroup() %>% inner_join(affyData2unique, by="element") %>% group_by(element) %>% mutate(samplename=samplename)  %>% 
  separate(samplename,c("trt","rep"),sep="\\.") %>% rename(value=predicted.values) %>% dplyr::select(element,value,AGI,description,trt,rep)
write_csv(fits.element.exp.value,file.path("custom_categories","S_Aarabi2016.fits.element.exp.value.csv.gz"))

# calculate mean value per element, treatment
fits.element.exp.value.mean <- fits.element.exp.value %>%
  group_by(element,trt) %>% 
  summarise(mean.element=mean(value)) %>% 
  spread(trt,mean.element,-1) %>%
  inner_join(affyData2unique, by="element") # only one element per one locus

#  check if "one" per one element
 fits.element.exp.value.mean %>% group_by(element) %>% summarize(num.element=n()) %>% arrange(num.element) %>% View()
 # calculate log fold changes per each locus (one locus = one element)
 fits.summary.AGI.FC <- fits.element.exp.value.mean %>% 
   group_by(AGI) %>% 
   mutate(mean.FC.AGI=log(`-S`/`+S`)) %>%
  # add FDR_trtS-
  inner_join(fits.summary.table.all 
            %>% dplyr::select(element,`FDR_trt-S`), by="element") 
 fits.summary.AGI.FC %>% View()
 # check AGI number
 fits.summary.AGI.FC %>% group_by(AGI) %>% summarize(num.AGI=n()) %>% arrange(num.AGI) %>% View() # there is no  "no_match"
# write
write_csv(fits.summary.AGI.FC,path=file.path("custom_categories","S_Aarabi2016.fits.summary.AGI.FC.csv.gz"))

```

# expression pattern 
```{r eval=TRUE}
fits.summary.AGI.FC <- read_csv(file.path("custom_categories","S_Aarabi2016.fits.summary.AGI.FC.csv.gz"))
fits.element.exp.value <- read_csv(file.path("custom_categories","S_Aarabi2016.fits.element.exp.value.csv.gz"))
# upregulated genes upon S deficienty
gene.of.interest.FDR.up <- fits.summary.AGI.FC %>% ungroup() %>% rename(FDR=`FDR_trt-S`) %>% filter(FDR<0.05,mean.FC.AGI>0) %>% dplyr::select(AGI,FDR) %>% arrange(FDR) 
print("logFC>0")
p.up<-expression.pattern.AT.graph.ATH1(target.genes.FDR=gene.of.interest.FDR.up[1:5,],
                              data=fits.element.exp.value,
                              sample.description=design,
                              title="S_Aarabi2016 root -S up") # good
p.up
ggsave(p.up,filename=file.path("custom_categories","S_Aarabi2016_root.up.png"))
# -S down
gene.of.interest.FDR.down <- fits.summary.AGI.FC %>% ungroup() %>% rename(FDR=`FDR_trt-S`) %>% filter(FDR<0.05,mean.FC.AGI<0) %>% dplyr::select(AGI,FDR) %>% arrange(FDR) 
print("logFC<0")
p.down<-expression.pattern.AT.graph.ATH1(target.genes.FDR=gene.of.interest.FDR.down[1:5,],
                              data=fits.element.exp.value,
                              sample.description=design,
                              title="S_Aarabi2016 root -S down") # good
p.down
ggsave(p.down,filename=file.path("custom_categories","S_Aarabi2016_root.down.png"))

```

# Session info
```{r}
sessionInfo()
```

