---
title: "Wrangle Leaf Data"
author: "Julin Maloof"
date: "7/25/2020"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We have leaf growth data for Mark.  I want to bring this in and get it associated with the appropriate RNAseq data.

```{r}
library(tidyverse)
library(readxl)
library(GGally)
```

## experiment 1

get leaf data
```{r}
leafexp1 <- read_excel("../input/exp1_leaf_ln_data.xlsx", sheet = 2) %>%
  filter(data_type=="RNAseq",
         time_point=="wk2")
leafexp1
```
quick lm
```{r}
leafdforlm <- leafexp1 %>% select(trt=`5E_soil_trts`, genotype, leaf_avg=avg_lf_03.24_no_germ) %>%
  mutate(trt=ifelse(trt=="live","live","dead_blank"))

lm1 <- lm(leaf_avg ~ genotype + trt, data = leafdforlm)

summary(lm1)

```


just keep the data we need
```{r}
leafexp1 <- leafexp1 %>% select(pot, leaf_avg=avg_lf_03.24_no_germ)
```


get RNAseq info
```{r}
rnaexp1 <- read_excel("../../RNA/input/wy001_RNAseq.xlsx") %>%
  filter(time_point=="wk2")
rnaexp1
```

we have tissues from both root and shoot, so this needs to be doubled
```{r}
rnaexp1 <- bind_rows({rnaexp1 %>% mutate(tissue="leaf")},
                     {rnaexp1 %>% mutate(tissue="root")})
```


merge them
```{r}
leafrnaexp1 <- rnaexp1 %>%
  left_join(leafexp1, by="pot") %>%
  mutate(exp="exp1")
leafrnaexp1
```
plot it
```{r}
leafrnaexp1 %>%
  mutate(group=str_c(soil, str_replace_na(autoclave))) %>%
  ggplot(aes(x=group, y=leaf_avg, color=genotype)) +
  geom_boxplot()
```

looks good!

## exp density


get leaf data
```{r}
leafexp_density <- read_excel("../input/density_exp_leaf_ln.xlsx", sheet = 2) %>%
  filter(density=="CR")
leafexp_density
```

just keep the data we need
```{r}
leafexp_density <- leafexp_density %>% select(pot=pot...2, leaf_avg=raw_avg_lf_11.07.17, genotype)
```

get RNAseq info
```{r}
rnaexp_density <- read_csv("../../RNA/input/wy003-metadata.csv") %>%
  filter(CROWDING=="CR") %>%
  rename_all(tolower)
rnaexp_density
```

merge leaf and RNA info
```{r}
leafrnaexp_density <- rnaexp_density %>%
  left_join(leafexp_density, by="pot") %>%
  mutate(exp="exp_density",
         soil=str_c(soil, "E"))
leafrnaexp_density
```
plot it
```{r}
leafrnaexp_density %>%
  mutate(group=str_c(soil, treatment)) %>%
  ggplot(aes(x=group, y=leaf_avg)) +
  geom_boxplot()
```

looks good!

## combine the experiments

wyo_root_FPsc_02_052

```{r}
leafrnaexp1 <- leafrnaexp1 %>% 
  mutate(sampleID=str_c("wyo", tissue, genotype, str_pad(block, 2, pad = "0"), str_pad(pot, 3, pad = "0"), sep = "_" ),
         pot=as.character(pot)) %>%
  select(sampleID, exp, pot, soil, tissue, genotype, autoclave, leaf_avg)
head(leafrnaexp1)
```


```{r}
leafrnaexp_density <- leafrnaexp_density %>% select(sampleID=`sample id`, exp, pot, soil, tissue, genotype, autoclave=treatment, leaf_avg)
head(leafrnaexp_density)
```

```{r}
leafrna <- bind_rows(leafrnaexp1, leafrnaexp_density)
leafrna
```
plot it

```{r}
leafrna %>%
  mutate(group=str_c(soil, autoclave)) %>%
  ggplot(aes(x=group,y=leaf_avg, color=genotype)) +
  geom_boxplot() +
  facet_wrap(~exp)
```

standardize for each group
```{r}
leafrna <- leafrna %>%
  mutate(group=str_c(exp, genotype, sep="_")) %>%
  group_by(group) %>%
  mutate(leaf_avg_std = scale(leaf_avg))
```

plot standardized
```{r}
leafrna %>%
  mutate(group=str_c(soil, autoclave)) %>%
  ggplot(aes(x=group,y=leaf_avg_std, color=genotype)) +
  geom_boxplot() +
  facet_wrap(~exp)
```

Do I get a similar result if I use residuals to remove gt and exp?

```{r}
lm1 <- lm(leaf_avg ~ genotype + exp, data = leafrna)
leafrna$leaf_avg_resid <- residuals(lm1)
```

```{r}
leafrna %>% ungroup() %>% select(starts_with("leaf")) %>% cor()
```

## now check that sample names match voom names

### rootnames
```{r}
rootnames <- readLines("../../RNA/output/voom_expression.e1and3.resid.root.txt.gz", n=1) %>%
  str_split("\t") %>%
  unlist() %>% magrittr::extract(-1)
rootnames
```


```{r}
leafnames <- readLines("../../RNA/output/voom_expression.e1and3.resid.leaf.txt.gz", n=1) %>%
  str_split("\t") %>%
  unlist() %>% magrittr::extract(-1)
leafnames
```

```{r}
voomnames <- c(rootnames, leafnames)
```

```{r}
voomnames %in% leafrna$sampleID
```

uggh, need to add the "S" number to density experiment

```{r}
voomnames <- tibble(fullname=voomnames) %>%
  mutate(basename=str_remove(fullname, "_S[0-9]*$"))
voomnames
```

```{r}
voomnames$basename %in% leafrna$sampleID
```


```{r}
leafrna <- leafrna %>% left_join(voomnames, by=c("sampleID" = "basename")) %>%
  select(sampleID, fullname, everything()) %>%
  mutate(fullname=ifelse(is.na(fullname), sampleID, fullname))
sum(leafrna$fullname %in% voomnames$fullname)
leafrna
```

```{r}
leafrna %>% select(fullname, exp, soil, tissue, genotype, autoclave, leaf_avg, leaf_avg_std)
write_csv(leafrna, "../output/leaf_lengths_combined.csv")
```


