---
title: "10_WGCNA_GSEA_core_enrichment_signedhybrid"
author: "Kazu"
date: "12/2/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
# History
* Extract core enrichment genes (030721)

# To Do
* modify this for signedhydrid WGCNA modules (041721)

```{r setup}
rm(list=ls())
knitr::opts_chunk$set(error = TRUE,warning=FALSE)
library(tidyverse);library(readr);library(readxl);library(openxlsx)
# The following setting is important for WGCNA, do not omit.
 options(stringsAsFactors = FALSE)
# install.packages(c("dynamicTreeCut", "cluster", "flashClust", "Hmisc", "reshape", "foreach", "doParallel") ) 
library(WGCNA);library(flashClust) # for WGCNA in Whitney
#allowWGCNAThreads(nThreads = NULL) # for Whitney
#enableWGCNAThreads(4) # for Whitney
library(scales) # for muted function
library(ggdendro) # for dendrogram
library(cowplot)
library(ggnewscale)

```

# annotation file for v3.0annotation
```{r}
# annotation file for v3.0annotation
Br.v3.0.At.BLAST <- read_csv(file.path("annotation","output","Brapa_v3.0_annotated.csv")) 
# This annotation is redundant with name (Br grene). Eg 
Br.v3.0.At.BLAST %>% filter(name=="BraA01g040570.3C")
# reduce the redundancy (112418)
Br.v3.0anno.At.BLAST.highscore <- Br.v3.0.At.BLAST %>% group_by(name) %>% arrange(desc(score)) %>% dplyr::slice(1)
# 052521 bHLH038??
Br.v3.0.At.BLAST %>% filter(str_detect(At_symbol,"b|BHLH")) %>% View()
Br.v3.0.At.BLAST %>% View()

```

# Julin's Machine Learning selected modules (signed hybrid)
```{r eval=FALSE}
# reading Julin's results of penalized regression (see slide) 
# 042721
ElasticNet0.5_multiCV_Pct_var_signedhybrid <- read_csv(file.path("RNA","output","ElasticNet0.5_multiCV_Pct_var.signed.hybrid.csv")) # Updated version for WGCNA singed hybrid modules (042321)
#
ElasticNet0.5_multiCV_Pct_var_signedhybrid <- ElasticNet0.5_multiCV_Pct_var_signedhybrid %>% unite(tissue.module,c("Module_Organ","Module_Name"),sep=".")  # all lower case

```

# reading GSEA results (all WGCNA modules)
# GSEA analysis with WGCNA modules
## read reports 
## function for oct30 reports copied from "10_WGCNA_including_R500.Rmd"
```{r eval=FALSE}
read.gsea_report_files_WGCNA <- function(gsea_report_path,GSEA.FDR.cutoff,tp=c("resid","removeblock"),data="exp_gt_block") {
# open xsl files and save as csvs
# gsea_report
gsea_report_files <- list.files(path=gsea_report_path,full.names=FALSE,include.dirs = FALSE,recursive=TRUE, pattern="(gsea_report_for)([[:print:]]+)(\\.xls)")
# ifelse(setCutoff=="Total", x <- "Total",if(setCutoff<10) {x <-  str_c("In0",setCutoff)} else {x <- str_c("In",setCutoff)})

# gsea_report_files <- gsea_report_files[str_detect(gsea_report_files,x)]
# print(str_c(x,"orless"))

gsea_report_files
# gsea_report (full name)
gsea_report_files2 <- list.files(path=gsea_report_path,full.names=TRUE,recursive=TRUE, pattern="(gsea_report_for)([[:print:]]+)(\\.xls)")
#gsea_report_files2 <- gsea_report_files2[str_detect(gsea_report_files2,x)] 
#gsea_report_files2
# read xls file (file name is "xls", actually tsv. That is why read_xls did not work)
GSEA_reports<-lapply(gsea_report_files2, function(x) read_tsv(file=file.path(x)))
# name
names(GSEA_reports) <- gsea_report_files
  gsub(".xls","",gsub("([[:print:]]+)(/)([[:print:]]+)","\\3",gsea_report_files))
# convert list into one data.frame
GSEA_reports.DF <- GSEA_reports %>% enframe(name="report") %>% unnest(value)
#GSEA_reports.DF %>% View()
# filter only significnt `FDR q-val` (GSEA.FDR.cutoff)
GSEA_reports.DF <- GSEA_reports.DF #%>% filter(`FDR q-val`< GSEA.FDR.cutoff)
# 
GSEA_reports.DF <- GSEA_reports.DF %>% separate(report,into=c("report2","report3"),sep="/") %>% mutate(report2=str_replace(report2,"remove.block","removeblock")) %>% separate(report2,into=c("exp","type","type2","tissue","categories"),sep="\\.",fill="left",extra="drop") 
# GSEA_reports.DF <- GSEA_reports.DF %>% filter(type2==data) # works?
#GSEA_reports.DF$LorD <- gsub("(gsea_report_for_)(0|1)([[:print:]]+)","\\2",GSEA_reports.DF$report3)
#GSEA_reports.DF <- GSEA_reports.DF %>% mutate(LorD=ifelse(LorD==0,"Live","Dead"))
GSEA_reports.DF %>% View()
ifelse(dim(GSEA_reports.DF)[1]==0,print("no significnat categories"),return(GSEA_reports.DF))
}
```


# Finding soil treatment responsive genes in WGCNA module (significant one) for GSEA_reports.DF.MLselected
```{r}
GSEA_reports.DF.MLselected <- read_csv(file.path("RNA","output","GSEA_reports.DFWGCNAsignedhybrid.MLselected.csv"))  # updated

GSEA_reports.DF.MLselected %>% View()
tsv.file <- GSEA_reports.DF.MLselected %>% dplyr::select(report3) %>% as_vector()

# function
read.coregenes <- function(default.path=gsea_report_path,
                           tsv.file.in.GSEA_reports=report3,
                           category.name = NAME
                           )

{
  gsea_report_files <- list.files(path=default.path,full.names=FALSE,include.dirs = FALSE,recursive=TRUE, pattern="(gsea_report_for)([[:print:]]+)(\\.tsv)")

temp.dir <- gsea_report_files[str_detect(gsea_report_files,pattern=tsv.file.in.GSEA_reports)] %>% str_split(pattern="/") %>% unlist() %>% .[1]
print(str_c("temp.dir is ",temp.dir))
# xls.file.name to read core gene list file
tsv.file.name <- str_c(category.name,".tsv") #%>% str_replace("UP","FC_UP") %>% str_replace("DOWN","FC_DOWN")
# 
core.data <- read_tsv(file.path(default.path,temp.dir,tsv.file.name)) %>% mutate(dirname=temp.dir) # actually the xls file is tsv files (found by Julin...) . X10 is blank, so read only nine columns (how to ?)
return(core.data)
}  
```
# using map() or pmap() to add "core" gene list tata to each row of GSEA_reports.DF.all (ended up to use old Loop way...)
## Is this script correct? Modules/tissue are not all used found in GSEA heatmap plot (Fig 3) (041321)
##
```{r}
gsea_report_path=file.path("~","Box","gsea_home","output","apr18_2021")
# loop (though this is not smart R way)
coregenes <-list()
temp <- GSEA_reports.DF.MLselected %>% filter(`FDR q-val`< 0.3) # filtered significant modules in GSEA
temp %>% View()
for(i in 1:nrow(temp)) {
  coregenes[[i]] <- read.coregenes(tsv.file.in.GSEA_reports=as_vector(temp[i,"report3"]),
                           category.name = as_vector(temp[i,grep("MSigDB",names(temp),value=TRUE)]))
} 
#names(coregenes) <- str_c(temp$tissue,"_tissue.",temp$NAME) # error fixed (temp$voom.module did not work) # this info misses LorD (041321)
###### needs to fix
names(coregenes) <- str_c(temp$tissue,"_tissue.",temp$NAME,".",temp$LorD)

#####################

# select genes with "yes" in enrichment
coregenes.yes <- list()
for(i in 1:nrow(temp)) {
  coregenes.yes[[i]] <- coregenes[[i]] %>% filter(`CORE ENRICHMENT`=="Yes") %>% dplyr::select(SYMBOL) %>% mutate(presence=as.numeric(1))
}
names(coregenes.yes) <- names(coregenes)
# combine (different length of column)
temp2 <- Br.v3.0anno.At.BLAST.highscore %>% dplyr::select(name,AGI,At_symbol, At_full_name,At_short_description) %>% left_join(coregenes.yes[[1]], by=c("name"="SYMBOL"))
for(i in 2:length(names(coregenes.yes))) {
temp2 <- temp2 %>% left_join(coregenes.yes[[i]], by=c("name"="SYMBOL"))
}
temp2 %>% View()
#
colnames(temp2)[str_detect(colnames(temp2),"presence")] <- names(coregenes.yes)
coregenes.yes.DF <- temp2
coregenes.yes.DF %>% View()
# replcae NA into 0 (only present/abset info)
coregenes.yes.DF[,-5:-1][is.na(coregenes.yes.DF[,-5:-1])] <- 0
# check
coregenes.yes.DF %>% View()
# write csv
write_csv(coregenes.yes.DF,file=file.path("RNA/output/GSEA_leaf_length_soil_core_enrichment_genes_signedhybrid.csv"))
# 
coregenes.yes.DF <- read_csv(file.path("RNA/output/GSEA_leaf_length_soil_core_enrichment_genes_signedhybrid.csv"))
# find common core genes
#coregenes.yes.DF %>% mutate(num=sum(colnames(coregenes.yes.DF)[6:12])) %>% dplyr::select(num) # does not work
str(coregenes.yes.DF)
# test
#coregenes.yes.DF %>% mutate(num=leaf_tissue.root.yellowgreen + root_tissue.leaf.lightsteelblue1) %>% View() # works
#coregenes.yes.DF %>% mutate(num=sum(leaf_tissue.root.yellowgreen,root_tissue.leaf.lightsteelblue1)) %>% View() # works
# use rowwise() and c_across()
coregenes.yes.DF <- coregenes.yes.DF %>% rowwise() %>% mutate(total=sum(c_across(leaf_tissue.root.darkred.Live:root_tissue.root.ivory.Dead)))# %>% View()
```

# how many genes are "Yes" in each module
```{r}
coregenes.yes.DF %>% dplyr::select(leaf_tissue.root.darkred.Live:root_tissue.root.ivory.Dead) %>% colSums()# work
```

# expression pattern
## expression data
```{r}
# plus R500 in residual
root.voom5 <- readr::read_tsv(file.path("RNA","output","voom_expression.e1and3.resid.exp_gt.root.plusR500.txt.gz"))
leaf.voom5 <- readr::read_tsv(file.path("RNA","output","voom_expression.e1and3.resid.exp_gt.leaf.plusR500.txt.gz"))
```

## WGCNA module (signed hybrid)
```{r}
#Br.v3.0anno.At.BLAST.highscore.modules2 <- read_csv(file.path("RNA","output","WGCNA.modules2.voom5.csv.gz"))
#head(Br.v3.0anno.At.BLAST.highscore.modules2)
```

## sample info
```{r}
sample.description.e1 <- read_csv(file.path("plant","output","Br.mbio.e1.sample.description.csv"))
sample.description.e3 <- read_csv(file.path("plant","output","Br.mbio.e3.sample.description.csv"))

sample.description.e1 %>% summarize(n_distinct(group))
sample.description.e1 %>% group_by(trt) %>% summarize(n())
##
root.voom5.e1 <- tibble(root.voom5 = colnames(root.voom5)[-1]) %>% inner_join(sample.description.e1[,c("sample","genotype","tissue","trt")],by=c("root.voom5"="sample"))
root.voom5.e3 <- tibble(root.voom5 = colnames(root.voom5)[-1]) %>% mutate(genotype="FPsc") %>% inner_join(sample.description.e3[,c("sample","tissue","trt")],by=c("root.voom5"="sample"))  
root.voom5.sample <- bind_rows(root.voom5.e1,root.voom5.e3) %>% mutate(trt=str_remove(trt,"5E_"))
root.voom5.sample %>% group_by(genotype) %>% summarise(num=n())

# leaf.voom5.sample
leaf.voom5.e1 <- tibble(leaf.voom5 = colnames(leaf.voom5)[-1]) %>% inner_join(sample.description.e1[,c("sample","genotype","tissue","trt")],by=c("leaf.voom5"="sample"))
leaf.voom5.e3 <- tibble(leaf.voom5 = colnames(leaf.voom5)[-1]) %>% mutate(genotype="FPsc") %>% inner_join(sample.description.e3[,c("sample","tissue","trt")],by=c("leaf.voom5"="sample"))  
leaf.voom5.sample <- bind_rows(leaf.voom5.e1,leaf.voom5.e3) %>% mutate(trt=str_remove(trt,"5E_"))
leaf.voom5.sample %>% group_by(genotype) %>% summarise(num=n())
```

# plotting expresion patterns
```{r}
# BraA01g010990.3C is ATIRT1
plotgene <- function(cpm=root.voom5, target="BraA01g010990.3C",samples=root.voom5.sample,numcol=4) {
  print(colnames(samples))
  print(unique(samples$tissue))
  data <- cpm %>% filter(genes %in% target) %>% 
    pivot_longer(cols=-genes,names_to="sampleID",values_to="log2value") %>%
    inner_join(samples,by=c("sampleID"=colnames(samples)[1]))
  print(data)
  plot <- data %>%
    ggplot(aes(y=log2value,x=trt,color=trt,shape=genotype)) + geom_jitter() +
    facet_wrap(genotype~genes, ncol=numcol, scales = "free_y") +
    theme(axis.text.x = element_text(angle = 90, hjust=1)) + labs(title=unique(samples$tissue))
  return(plot)
}

plotgene() # BraA01g010990.3C is ATIRT1

plotgene(cpm=leaf.voom5,samples=leaf.voom5.sample) # error due to not in leaf data
plotgene(cpm=leaf.voom5,samples=leaf.voom5.sample,target=leaf.voom5$genes[1]
) # no error. 
```

# root_tissue.leaf.maroon.Dead (71 genes)
```{r}
GSEA.root_tissue.leaf.maroon.Dead.coregenes.yes <- coregenes.yes.DF %>% filter(root_tissue.leaf.maroon.Dead==1) %>% dplyr::select(name) %>% as_vector()
p <- plotgene(target=GSEA.root_tissue.leaf.maroon.Dead.coregenes.yes)
ggsave(p,filename="GSEA.root_tissue.leaf.maroon.Dead.coregenes.yes.plot.png",path=file.path("RNA","output"),height=40,width=8,limitsize=FALSE)
```


# expression pattern of genes in leaf.red with specifc GO enriched in the module (revised for signed hybrid)
```{r}
# find target genes from module name and enriched GO term
# module ~ GO term
WGCNA.voom5.signedhybrid.all.modules.GOseq <- read_csv(file.path("RNA","output","WGCNA.voom5.signedhybrid.all.modules.GOseq.csv"))
# gene ~ GO term
load(file.path("annotation","output","Brgo.v3.0anno.Atgoslim.BP.list.Rdata"))
# gene ~ WGCNA module 
Br.v3.0anno.At.BLAST.highscore.modules.signedhybrid <- read_csv(file.path("RNA","output","WGCNA.signedhybrid.modules.voom5.csv.gz"))

#Br.v3.0anno.At.BLAST.highscore.modules2
# find genes in leaf.red module with enriched GO terms
Brgo.v3.0anno.Atgoslim.BP.list.DF <- Brgo.v3.0anno.Atgoslim.BP.list %>% enframe() %>% unnest(value)
# 
genes.category.module <- Brgo.v3.0anno.Atgoslim.BP.list.DF %>% left_join(Br.v3.0anno.At.BLAST.highscore.modules.signedhybrid[,c("name","root.voom5.module.signedhybrid","leaf.voom5.module.signedhybrid")],by="name")
# root.orangered4
target.GO <- WGCNA.voom5.signedhybrid.all.modules.GOseq %>% dplyr::select(name,category,over_represented_pvalue,term) %>% mutate(name=str_replace(name,".voom5.module.signedhybrid_",".")) %>%  filter(name=="root.orangered4")
target.GO
#
target.genes <- genes.category.module %>% filter(root.voom5.module.signedhybrid=="orangered4") %>% filter(value==target.GO$category[1]) %>% distinct(name)
# 
target.genes <- genes.category.module %>% left_join(Br.v3.0anno.At.BLAST.highscore,by="name") %>%   filter(root.voom5.module.signedhybrid=="orangered4") %>% filter(value==target.GO$category[1]) %>% distinct(name,.keep_all=TRUE)  # distinct(name) shows only "name" column.
target.genes %>% View()
# expression pattern ~ soil treatment 
p.target <- plotgene(target=target.genes$name,numcol=8)
p.target
ggsave(p.target,filename="root.orangered4.coregenes.yes.plot.png",path=file.path("RNA","output"),height=100,width=8,limitsize=FALSE) # soil treatment
```

# all modules with only target module 
* copied from 11_machine_learning.Rmd
leafdata (copied from 11_machine_learning.Rmd)
```{r}
leaf <- read_csv("plant/output/leaf_lengths_combined.csv") %>%
  mutate(sampleID=str_remove(sampleID, "_(leaf|root)"),
         sampleID=str_remove(sampleID, "^[a-h][1-8]_")) %>%
  filter(!duplicated(sampleID))

leaf %>% arrange(sampleID)
```

load root modules (copied from 11_machine_learning.Rmd) (revised 042021)
```{r}
root.voom5 <- readr::read_tsv(file.path("RNA","output","voom_expression.e1and3.resid.exp_gt.root.plusR500.txt.gz"))
load(file.path("RNA","output","voom_expression.e1and3.resid.exp_gt.root.plusR500.signedhybrid.txt.gz.WGCNA.softthresh10.signedhybrid.RData")) # revised (042021)
# adding module column
#root.voom5.mod <- root.voom5 %>% mutate(dat.voom.module=str_c("ME",dynamicColors,"_root"))
root.voom5.signedhybrid.mod <- root.voom5 %>% mutate(dat.voom.module=str_c("root.",dynamicColors))

# MEs
rootMEs <- MEs %>% as_tibble() %>%
  rename_all(str_c, "_root") %>%
  mutate(sampleID=colnames(root.voom5)[-1],
         sampleID=str_remove(sampleID, "_root"),
         sampleID=str_remove(sampleID, "_S[0-9]{1,3}$"),
         sampleID=str_remove(sampleID, "^[a-h][1-8]_")) %>%
  dplyr::select(sampleID, everything())
rootMEs
rm(MEs)
```

load leaf modules (copied from 11_machine_learning.Rmd) (revised 042021)
```{r}
leaf.voom5 <- readr::read_tsv(file.path("RNA","output","voom_expression.e1and3.resid.exp_gt.leaf.plusR500.txt.gz"))
load(file.path("RNA","output","voom_expression.e1and3.resid.exp_gt.leaf.plusR500.signedhybrid.txt.gz.WGCNA.softthresh10.signedhybrid.RData"))
# adding module column (tissue.module all lowercase that is consistent with )
#leaf.voom5.mod <- leaf.voom5 %>% mutate(dat.voom.module=str_c("ME",dynamicColors,"_leaf"))
leaf.voom5.signedhybrid.mod <- leaf.voom5 %>% mutate(dat.voom.module=str_c("leaf.",dynamicColors)) # modified (042021)
# 
leafMEs <- MEs %>% as_tibble() %>%
  rename_all(str_c, "_leaf") %>%
  mutate(sampleID=colnames(leaf.voom5)[-1],
         sampleID=str_remove(sampleID, "_leaf"),
         sampleID=str_remove(sampleID, "_S[0-9]{1,3}$"),
         sampleID=str_remove(sampleID, "^[a-h][1-8]_")) %>%
  dplyr::select(sampleID, everything())
leafMEs
rm(MEs)
```

bring it all together (copied from 11_machine_learning.Rmd)
```{r}
d <- leaf %>% dplyr::select(sampleID, leaf_avg_std) %>%
  inner_join(leafMEs, by="sampleID") %>% 
  inner_join(rootMEs, by="sampleID") %>% 
  # dplyr::select(-sampleID) %>% # Kazu need sampleID later
  as_tibble()
d
```

# plot
* see Julin's request (https://trello.com/c/Y4GWD7KJ/35-wgcna-and-mr-analysis)
* update for signed hybrid WGCNA modules (042721)
```{r}
# new ver (tissue.module lowercase)
ElasticNet0.5_multiCV_Pct_var_signedhybrid.module <- ElasticNet0.5_multiCV_Pct_var_signedhybrid %>% dplyr::select(tissue.module) %>% as_vector()
# new ver. use anti_join()
nonElasticNet0.5_multiCV_Pct_var_signedhybrid.module <- tibble(all = colnames(d)[-1:-2]) %>% separate(all, c("module","tissue")) %>% mutate(module=str_remove(module,"ME")) %>% unite(tissue.module.all,tissue,module,sep=".") %>% anti_join(ElasticNet0.5_multiCV_Pct_var_signedhybrid,by=c("tissue.module.all"="tissue.module")) %>% as_vector()
# target module (an example)
#mod <- ElasticNet0.5_multiCV_Pct_var.module[3]
mod <-"root.ivory"
# making expression pattern plot function (compare within/outside module)
#
d.root <- d %>% dplyr::select(sampleID,leaf_avg_std,ends_with("root"))
d.leaf <- d %>% dplyr::select(sampleID,leaf_avg_std,ends_with("leaf"))
# 
expression.module4 <- function(target.module=mod,dat.voom=root.voom5.signedhybrid.mod,tissue_MEs=d.root) { 
  print(str_c("The target.module is ",target.module,"."))
# add within target.module info
dat.voom <- dat.voom %>% mutate(TF=ifelse(dat.voom.module==target.module,"within the module","not in the module"),.keep="unused") %>% mutate(TF.module=str_c(target.module,"\n",TF))
# number of genes with/not in the module
dat.voom %>% group_by(TF) %>% summarize(nTF=n())
# longer format
dat.voom.long <- dat.voom %>% pivot_longer(-c(genes,TF,TF.module),names_to="sampleID",values_to="value") %>% mutate(sampleID=str_remove(sampleID,"_root")) %>% mutate(sampleID=str_remove(sampleID,"_leaf"))
# modify for new ME
temp.target.module <- tibble(tissue_MEs=colnames(tissue_MEs)[-2:-1]) %>% mutate(tissue_MEs = str_remove(tissue_MEs,"^ME")) %>% separate(tissue_MEs,c("module","tissue")) %>% unite(tissue.module,tissue,module,sep=".") %>% as_vector()
# change format of colnames of tissue_MEs
colnames(tissue_MEs)[-2:-1] <- temp.target.module
# get sample levels by leaf_avg_std
sample.levels <- tissue_MEs %>% dplyr::select(sampleID, leaf_avg_std,quo_name(target.module)) %>% arrange_at(.vars=2) %>% dplyr::select(sampleID) %>% as_vector()
# add expression data 
dat.voom.ME <- tissue_MEs %>% dplyr::select(sampleID,quo_name(target.module))  %>% inner_join(dat.voom.long,by="sampleID") %>%
  arrange_at(.vars=c(2,4)) %>% mutate(sampleID=factor(sampleID,levels=sample.levels))
# plot 
print(dat.voom.ME)
# but aes_() will be soft deprecated, so that need to use quasiquotation (!!, !!!, :=)  (https://www.tidyverse.org/blog/2018/07/ggplot2-tidy-evaluation/)
# https://adv-r.hadley.nz/quasiquotation.html
# plot expression value plus ME within the module
p.TF10.withinmod <- dat.voom.ME  %>%  filter(TF=="within the module") %>% ggplot(aes(x=sampleID,y=value,color=TF,group=genes)) + geom_line(alpha=0.1)  + geom_line(aes_string(x="sampleID",y=target.module,group="genes"),color="black",size=1) + theme(axis.text.x=element_text(angle=90),legend.position="none")  + facet_grid(.~TF.module) +  labs(ylab="normalized expression") #title=str_remove(target.module,"ME"),
# plot expression value not in the module
p.TF10.notinmod <- dat.voom.ME  %>%  filter(TF=="not in the module") %>% ggplot(aes(x=sampleID,y=value,color=TF,group=genes)) + geom_line(alpha=0.1)  +  theme(axis.text.x=element_text(angle=90),legend.position="none")  + facet_grid(.~TF.module) +  labs(ylab="normalized expression") #title=str_remove(target.module,"ME"),
# combine the two plots
p.TFall <- plot_grid(p.TF10.withinmod,p.TF10.notinmod)
return(p.TFall)
}
# test 
mod=ElasticNet0.5_multiCV_Pct_var_signedhybrid$tissue.module[1]
expression.module4(target.module=mod)
# 
plot.root.list4 <- tibble(tissue_MEs=colnames(d)[-2:-1]) %>% mutate(tissue_MEs = str_remove(tissue_MEs,"^ME")) %>% separate(tissue_MEs,c("module","tissue")) %>% unite(tissue.module,tissue,module,sep=".") %>% filter(str_detect(tissue.module,"^root")) %>% split(.$tissue.module) %>% map(~expression.module4(target.module=as.name(.$tissue.module),dat.voom=root.voom5.signedhybrid.mod,tissue_MEs=d.root)) # takes minutes

plot.leaf.list4 <- tibble(tissue_MEs=colnames(d)[-2:-1]) %>% mutate(tissue_MEs = str_remove(tissue_MEs,"^ME")) %>% separate(tissue_MEs,c("module","tissue")) %>% unite(tissue.module,tissue,module,sep=".") %>% filter(str_detect(tissue.module,"^leaf")) %>% split(.$tissue.module) %>% map(~expression.module4(target.module=as.name(.$tissue.module),dat.voom=leaf.voom5.signedhybrid.mod,tissue_MEs=d.leaf)) # takes minutes
# ElasticNet0.5_multiCV_Pct_var plots
plot.root.list4[[ElasticNet0.5_multiCV_Pct_var_signedhybrid.module[1]]]
plot.root.list4[[ElasticNet0.5_multiCV_Pct_var_signedhybrid.module[2]]]
plot.root.list4[[ElasticNet0.5_multiCV_Pct_var_signedhybrid.module[3]]]
# nonElasticNet0.5_multiCV_Pct_var plots
plot.root.list4[[nonElasticNet0.5_multiCV_Pct_var_signedhybrid.module[1]]]
plot.root.list4[[nonElasticNet0.5_multiCV_Pct_var_sinedhybrid.module[2]]]
plot.root.list4[[nonElasticNet0.5_multiCV_Pct_var_signedhybrid.module[3]]]
```

# how to write figs for plots?
```{r}
p.root.all <- plot_grid(plot.root.list4[[1]],plot.root.list4[[2]],plot.root.list4[[3]],plot.root.list4[[4]],
                   plot.root.list4[[5]],plot.root.list4[[6]],plot.root.list4[[7]],plot.root.list4[[8]],
                   plot.root.list4[[9]],plot.root.list4[[10]],plot.root.list4[[11]],plot.root.list4[[12]],
                   plot.root.list4[[13]],plot.root.list4[[14]],plot.root.list4[[15]],plot.root.list4[[16]],
                   plot.root.list4[[17]],plot.root.list4[[18]],plot.root.list4[[19]],plot.root.list4[[20]],
                   plot.root.list4[[21]],plot.root.list4[[22]],plot.root.list4[[23]],plot.root.list4[[24]],plot.root.list4[[25]])
ggsave(p.root.all,path=file.path("RNA","output"),filename="p.root.all.signedhybrid.png",height=40,width=20,dpi=72)
p.leaf.all <- plot_grid(plot.leaf.list4[[1]],plot.leaf.list4[[2]],plot.leaf.list4[[3]],plot.leaf.list4[[4]],
                   plot.leaf.list4[[5]],plot.leaf.list4[[6]],plot.leaf.list4[[7]],plot.leaf.list4[[8]],
                   plot.leaf.list4[[9]],plot.leaf.list4[[10]],plot.leaf.list4[[11]],plot.leaf.list4[[12]],
                   plot.leaf.list4[[13]],plot.leaf.list4[[14]],plot.leaf.list4[[15]],plot.leaf.list4[[16]],
                   plot.leaf.list4[[17]],plot.leaf.list4[[18]],plot.leaf.list4[[19]],plot.leaf.list4[[20]],
                   plot.leaf.list4[[21]],plot.leaf.list4[[22]],plot.leaf.list4[[23]],plot.leaf.list4[[24]],
                   plot.leaf.list4[[25]],plot.leaf.list4[[26]],plot.leaf.list4[[27]],plot.leaf.list4[[28]],
                   plot.leaf.list4[[29]],plot.leaf.list4[[30]],plot.leaf.list4[[31]])
ggsave(p.leaf.all,path=file.path("RNA","output"),filename="p.leaf.all.signedhybrid.png",height=40,width=20,dpi=72)

# only leaf length associated modules
p.root.leaf_length.associated <- plot_grid(plot.root.list4[[ElasticNet0.5_multiCV_Pct_var_signedhybrid.module[1]]],
          plot.root.list4[[ElasticNet0.5_multiCV_Pct_var_signedhybrid.module[2]]],
          plot.root.list4[[ElasticNet0.5_multiCV_Pct_var_signedhybrid.module[3]]],                                plot.root.list4[[ElasticNet0.5_multiCV_Pct_var_signedhybrid.module[8]]])
ggsave(p.root.leaf_length.associated,path=file.path("RNA","output"),filename="p.root.leaf_length.associated.signedhybrid.png",height=11,width=8,dpi=72)

p.leaf.leaf_length.associated <- plot_grid(plot.leaf.list4[[ElasticNet0.5_multiCV_Pct_var_signedhybrid.module[4]]],
          plot.leaf.list4[[ElasticNet0.5_multiCV_Pct_var_signedhybrid.module[5]]],                               plot.leaf.list4[[ElasticNet0.5_multiCV_Pct_var_signedhybrid.module[6]]],
          plot.leaf.list4[[ElasticNet0.5_multiCV_Pct_var_signedhybrid.module[7]]],                               plot.leaf.list4[[ElasticNet0.5_multiCV_Pct_var_signedhybrid.module[9]]],
          plot.leaf.list4[[ElasticNet0.5_multiCV_Pct_var_signedhybrid.module[10]]])
ggsave(p.leaf.leaf_length.associated,path=file.path("RNA","output"),filename="p.leaf.leaf_length.associated.signedhybrid.png",height=11,width=12,dpi=72)

```


# Expression pattern in leaf: leaf.red module genes with enriched GO term (use expression.module6())
# leaf length vs expression level 
```{r}
expression.module6 <- function(tissue=c("leaf","root"),target.module=mod,target=target.genes$name,dat.voom=leaf.voom5.mod,leaf_length=leaf,target.GOterm="") { 
  print(str_c("The target.module is ",target.module,"."))
# longer format
dat.voom.long <- dat.voom %>% filter(genes %in% target) %>% dplyr::select(-dat.voom.module) %>%   pivot_longer(-genes,names_to="sampleID",values_to="value") %>% mutate(sampleID=str_remove(sampleID,"_root")) %>% mutate(sampleID=str_remove(sampleID,"_leaf")) %>% inner_join(leaf_length[,c("sampleID","leaf_avg_std")],by="sampleID")
#
dat.voom.long %>% View()
# plot expression value 
p <- dat.voom.long %>% ggplot(aes(x=leaf_avg_std,y=value,group=genes)) + geom_line(alpha=0.1) + theme(axis.text.x=element_text(angle=90),legend.position="none")  +  labs(title=str_c(target.module,target.GOterm,sep=" "),y=str_c("expression in ",tissue))
return(p)
}
```

<!-- # Expression pattern in leaf: "leaf.red" module genes with enriched GO term -->

# revised version "root_tissue.leaf.maroon.Dead"
```{r}
# 042721
ElasticNet0.5_multiCV_Pct_var_signedhybrid <- read_csv(file.path("RNA","output","ElasticNet0.5_multiCV_Pct_var.signed.hybrid.csv")) # Updated version for WGCNA singed hybrid modules (042321)
#
ElasticNet0.5_multiCV_Pct_var_signedhybrid <- ElasticNet0.5_multiCV_Pct_var_signedhybrid %>% unite(tissue.module,c("Module_Organ","Module_Name"),sep=".")  # all lower case
# 
mod <- "leaf.maroon"
target.GO <- WGCNA.voom5.signedhybrid.all.modules.GOseq %>% dplyr::select(name,category,term) %>% mutate(name=str_replace(name,".voom5.module.signedhybrid_",".")) %>% filter(name==mod) %>% unite(category.term,c(category, term),sep=" ",remove=FALSE)
target.GO
# for root tissue expression pattern
genes.category.module <- Brgo.v3.0anno.Atgoslim.BP.list.DF %>% left_join(Br.v3.0anno.At.BLAST.highscore.modules.signedhybrid[,c("name","root.voom5.module.signedhybrid","leaf.voom5.module.signedhybrid")],by="name")
# GO 1
target.genes <- genes.category.module %>% dplyr::select(name,,term) %>% filter(leaf.voom5.module.signedhybrid==str_remove(mod,"leaf.")) %>% filter(value==target.GO$category[1]) %>% distinct(name)
#
p.root.leaf.maroon.GO1 <- expression.module6(tissue="root",
                                        target.module=mod,
                                        target=target.genes$name,
                                        dat.voom=root.voom5.signedhybrid.mod,
                                        leaf_length=leaf,
                                        target.GOterm=target.GO$category.term[1])
p.root.leaf.maroon.GO1
ggsave(p.root.leaf.maroon.GO1,filename="p.root.leaf.maroon.GO1.leaf_length.associated.png",path=file.path("RNA","output"))
```


# GO ORA of GSEA yes genes
```{r}
# test
GOseq.root_tissue.root.ivory.Dead <- coregenes.yes.DF %>% filter(root_tissue.root.ivory.Dead==1) %>% dplyr::select(name) %>% as_vector() %>% GOseq.Brgo.v3.0.Atgoslim.BP.list.ORA2(genelist=.)
# all 
GOseq.temp <- coregenes.yes %>% map(~GOseq.Brgo.v3.0.Atgoslim.BP.list.ORA2(genelist=.$SYMBOL))
GOseq.temp %>% enframe() %>% unnest(value) %>% write_csv(file.path("RNA","output","WGCNA.voom5.signedhybrid.GSEA.yes.modules.GO.GOseq.csv"))
```

# custom category ORA of GSEA yes genes
```{r}
# custom categories list
load(file.path("annotation","output","GSEA_cust_rootshoot_updown3.FDR0.001.list.Rdata")) # updated GSEA_cust_rootshoot_updown2.list# all 
GOseq.temp <- coregenes.yes %>% map(~GOseq.Brgo.v3.0.Atgoslim.BP.list.ORA2(genelist=.$SYMBOL,custom.category.list=GSEA_cust_rootshoot_updown3.FDR0.001.list,ontology="none"))
GOseq.temp %>% enframe() %>% unnest(value) %>% write_csv(file.path("RNA","output","WGCNA.voom5.signedhybrid.GSEA.yes.modules.customcat.GOseq.csv"))
```



