---
title: "10_WGCNA_GSEA_core_enrichment"
author: "Kazu"
date: "12/2/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
# History
* Extract core enrichment genes (030721)

# To Do

```{r setup}
rm(list=ls())
knitr::opts_chunk$set(error = TRUE,warning=FALSE)
library(tidyverse);library(readr);library(readxl);library(openxlsx)
# The following setting is important for WGCNA, do not omit.
 options(stringsAsFactors = FALSE)
# install.packages(c("dynamicTreeCut", "cluster", "flashClust", "Hmisc", "reshape", "foreach", "doParallel") ) 
library(WGCNA);library(flashClust) # for WGCNA in Whitney
#allowWGCNAThreads(nThreads = NULL) # for Whitney
#enableWGCNAThreads(4) # for Whitney
library(scales) # for muted function
library(ggdendro) # for dendrogram
library(cowplot)
library(ggnewscale)

```

# annotation file for v3.0annotation
```{r}
# annotation file for v3.0annotation
Br.v3.0.At.BLAST <- read_csv(file.path("annotation","output","Brapa_v3.0_annotated.csv")) 
# This annotation is redundant with name (Br grene). Eg 
Br.v3.0.At.BLAST %>% filter(name=="BraA01g040570.3C")
# reduce the redundancy (112418)
Br.v3.0anno.At.BLAST.highscore <- Br.v3.0.At.BLAST %>% group_by(name) %>% arrange(desc(score)) %>% dplyr::slice(1)
```

# Julin's Machine Learning selected modules
```{r eval=FALSE}
# reading Julin's results of penalized regression (see slide) (old version)
# ElasticNet0.5_LOO_prop_var <- read_csv(file.path("plant","output","ElasticNet0.5_LOO_prop_var.csv")) 
ElasticNet0.5_multiCV_Pct_var <- read_csv(file.path("plant","output","ElasticNet0.5_multiCV_Pct_var.csv")) # Updated version from 11_take2_Machine_Learning
# 
ElasticNet0.5_multiCV_Pct_var <- ElasticNet0.5_multiCV_Pct_var %>%
  #mutate(Module_Name=str_to_upper(Module_Name)) %>%  # tissue is lowercase, module is uppercase
  unite(tissue.module,c("Module_Organ","Module_Name"),sep=".") # all lower case
```

# reading GSEA results (all WGCNA modules)
# GSEA analysis with WGCNA modules
## read reports 
## function for oct30 reports copied from "10_WGCNA_including_R500.Rmd"
```{r eval=FALSE}
read.gsea_report_files_WGCNA <- function(gsea_report_path,GSEA.FDR.cutoff,tp=c("resid","removeblock"),data="exp_gt_block") {
# open xsl files and save as csvs
# gsea_report
gsea_report_files <- list.files(path=gsea_report_path,full.names=FALSE,include.dirs = FALSE,recursive=TRUE, pattern="(gsea_report_for)([[:print:]]+)(\\.xls)")
# ifelse(setCutoff=="Total", x <- "Total",if(setCutoff<10) {x <-  str_c("In0",setCutoff)} else {x <- str_c("In",setCutoff)})

# gsea_report_files <- gsea_report_files[str_detect(gsea_report_files,x)]
# print(str_c(x,"orless"))

gsea_report_files
# gsea_report (full name)
gsea_report_files2 <- list.files(path=gsea_report_path,full.names=TRUE,recursive=TRUE, pattern="(gsea_report_for)([[:print:]]+)(\\.xls)")
#gsea_report_files2 <- gsea_report_files2[str_detect(gsea_report_files2,x)] 
#gsea_report_files2
# read xls file (file name is "xls", actually tsv. That is why read_xls did not work)
GSEA_reports<-lapply(gsea_report_files2, function(x) read_tsv(file=file.path(x)))
# name
names(GSEA_reports) <- gsea_report_files
  gsub(".xls","",gsub("([[:print:]]+)(/)([[:print:]]+)","\\3",gsea_report_files))
# convert list into one data.frame
GSEA_reports.DF <- GSEA_reports %>% enframe(name="report") %>% unnest(value)
#GSEA_reports.DF %>% View()
# filter only significnt `FDR q-val` (GSEA.FDR.cutoff)
GSEA_reports.DF <- GSEA_reports.DF #%>% filter(`FDR q-val`< GSEA.FDR.cutoff)
# 
GSEA_reports.DF <- GSEA_reports.DF %>% separate(report,into=c("report2","report3"),sep="/") %>% mutate(report2=str_replace(report2,"remove.block","removeblock")) %>% separate(report2,into=c("exp","type","type2","tissue","categories"),sep="\\.",fill="left",extra="drop") 
# GSEA_reports.DF <- GSEA_reports.DF %>% filter(type2==data) # works?
#GSEA_reports.DF$LorD <- gsub("(gsea_report_for_)(0|1)([[:print:]]+)","\\2",GSEA_reports.DF$report3)
#GSEA_reports.DF <- GSEA_reports.DF %>% mutate(LorD=ifelse(LorD==0,"Live","Dead"))
GSEA_reports.DF %>% View()
ifelse(dim(GSEA_reports.DF)[1]==0,print("no significnat categories"),return(GSEA_reports.DF))
}
```


# Finding soil treatment responsive genes in WGCNA module (significant one) for GSEA_reports.DF.MLselected
```{r}
GSEA_reports.DF.MLselected <- read_csv(file.path("RNA","output","GSEA_reports.DF.MLselected.csv"))

GSEA_reports.DF.MLselected %>% View()
xls.file <- GSEA_reports.DF.MLselected %>% dplyr::select(report3) %>% as_vector()

# function
read.coregenes <- function(default.path=gsea_report_path,
                           xls.file.in.GSEA_reports=report3,
                           category.name = NAME
                           )

{
  gsea_report_files <- list.files(path=default.path,full.names=FALSE,include.dirs = FALSE,recursive=TRUE, pattern="(gsea_report_for)([[:print:]]+)(\\.xls)")

temp.dir <- gsea_report_files[str_detect(gsea_report_files,pattern=xls.file.in.GSEA_reports)] %>% str_split(pattern="/") %>% unlist() %>% .[1]
print(str_c("temp.dir is ",temp.dir))
# xls.file.name to read core gene list file
xls.file.name <- str_c(category.name,".xls") #%>% str_replace("UP","FC_UP") %>% str_replace("DOWN","FC_DOWN")
# 
core.data <- read_tsv(file.path(default.path,temp.dir,xls.file.name)) %>% mutate(dirname=temp.dir) # actually the xls file is tsv files (found by Julin...) . X10 is blank, so read only nine columns (how to ?)
return(core.data)
}  
```
# using map() or pmap() to add "core" gene list tata to each row of GSEA_reports.DF.all (ended up to use old Loop way...)
## Is this script correct? Modules/tissue are not all used found in GSEA heatmap plot (Fig 3) (041321)
##
```{r}
gsea_report_path=file.path("~","Box","gsea_home","output","oct30")
# loop (though this is not smart R way)
coregenes <-list()
temp <- GSEA_reports.DF.MLselected %>% filter(`FDR q-val`< 0.3)
temp %>% View()
for(i in 1:nrow(temp)) {
  coregenes[[i]] <- read.coregenes(xls.file.in.GSEA_reports=as_vector(temp[i,"report3"]),
                           category.name = as_vector(temp[i,grep("MSigDB",names(temp),value=TRUE)]))
} 
#names(coregenes) <- str_c(temp$tissue,"_tissue.",temp$NAME) # error fixed (temp$voom.module did not work) # this info misses LorD (041321)
###### needs to fix


#####################

# select genes with "yes" in enrichment
coregenes.yes <- list()
for(i in 1:nrow(temp)) {
  coregenes.yes[[i]] <- coregenes[[i]] %>% filter(`CORE ENRICHMENT`=="Yes") %>% dplyr::select(PROBE) %>% mutate(presence=as.numeric(1))
}
names(coregenes.yes) <- names(coregenes)
# combine (different length of column)
temp2 <- Br.v3.0anno.At.BLAST.highscore %>% dplyr::select(name,AGI,At_symbol, At_full_name,At_short_description) %>% left_join(coregenes.yes[[1]], by=c("name"="PROBE"))
for(i in 2:length(names(coregenes.yes))) {
temp2 <- temp2 %>% left_join(coregenes.yes[[i]], by=c("name"="PROBE"))
}
temp2 %>% View()
#
colnames(temp2)[str_detect(colnames(temp2),"presence")] <- names(coregenes.yes)
coregenes.yes.DF <- temp2
coregenes.yes.DF %>% View()
# replcae NA into 0 (only present/abset info)
coregenes.yes.DF[,-5:-1][is.na(coregenes.yes.DF[,-5:-1])] <- 0
# check
coregenes.yes.DF %>% View()
# write csv
write_csv(coregenes.yes.DF,file=file.path("RNA/output/GSEA_leaf_length_soil_core_enrichment_genes.csv"))
# 
coregenes.yes.DF <- read_csv(file.path("RNA/output/GSEA_leaf_length_soil_core_enrichment_genes.csv"))
# find common core genes
coregenes.yes.DF %>% mutate(num=sum(colnames(coregenes.yes.DF)[6:12])) %>% dplyr::select(num) # does not work
str(coregenes.yes.DF)
# test
coregenes.yes.DF %>% mutate(num=leaf_tissue.root.yellowgreen + root_tissue.leaf.lightsteelblue1) %>% View() # works
coregenes.yes.DF %>% mutate(num=sum(leaf_tissue.root.yellowgreen,root_tissue.leaf.lightsteelblue1)) %>% View() # works
# use rowwise() and c_across()
coregenes.yes.DF <- coregenes.yes.DF %>% rowwise() %>% mutate(total=sum(c_across(leaf_tissue.root.yellowgreen:root_tissue.root.yellowgreen)))# %>% View()
```

# how many genes are "Yes" in each module
```{r}
coregenes.yes.DF %>% summarise(across(leaf_tissue.root.yellowgreen:root_tissue.root.yellowgreen,sum)) # why does not work?
coregenes.yes.DF %>% summarise(across(leaf_tissue.root.yellowgreen:root_tissue.root.yellowgreen,mean)) # why does not work?
coregenes.yes.DF %>% dplyr::select(leaf_tissue.root.yellowgreen:root_tissue.root.yellowgreen) %>% colSums()# work
```



# expression pattern
## expression data
```{r}
# plus R500 in residual
root.voom5 <- readr::read_tsv(file.path("RNA","output","voom_expression.e1and3.resid.exp_gt.root.plusR500.txt.gz"))
leaf.voom5 <- readr::read_tsv(file.path("RNA","output","voom_expression.e1and3.resid.exp_gt.leaf.plusR500.txt.gz"))
```

## WGCNA module
```{r}
Br.v3.0anno.At.BLAST.highscore.modules2 <- read_csv(file.path("RNA","output","WGCNA.modules2.voom5.csv.gz"))
head(Br.v3.0anno.At.BLAST.highscore.modules2)
```

## sample info
```{r}
sample.description.e1 <- read_csv(file.path("plant","output","Br.mbio.e1.sample.description.csv"))
sample.description.e3 <- read_csv(file.path("plant","output","Br.mbio.e3.sample.description.csv"))

sample.description.e1 %>% summarize(n_distinct(group))
sample.description.e1 %>% group_by(trt) %>% summarize(n())
##
root.voom5.e1 <- tibble(root.voom5 = colnames(root.voom5)[-1]) %>% inner_join(sample.description.e1[,c("sample","genotype","tissue","trt")],by=c("root.voom5"="sample"))
root.voom5.e3 <- tibble(root.voom5 = colnames(root.voom5)[-1]) %>% mutate(genotype="FPsc") %>% inner_join(sample.description.e3[,c("sample","tissue","trt")],by=c("root.voom5"="sample"))  
root.voom5.sample <- bind_rows(root.voom5.e1,root.voom5.e3) %>% mutate(trt=str_remove(trt,"5E_"))
root.voom5.sample %>% group_by(genotype) %>% summarise(num=n())

# leaf.voom5.sample
leaf.voom5.e1 <- tibble(leaf.voom5 = colnames(leaf.voom5)[-1]) %>% inner_join(sample.description.e1[,c("sample","genotype","tissue","trt")],by=c("leaf.voom5"="sample"))
leaf.voom5.e3 <- tibble(leaf.voom5 = colnames(leaf.voom5)[-1]) %>% mutate(genotype="FPsc") %>% inner_join(sample.description.e3[,c("sample","tissue","trt")],by=c("leaf.voom5"="sample"))  
leaf.voom5.sample <- bind_rows(leaf.voom5.e1,leaf.voom5.e3) %>% mutate(trt=str_remove(trt,"5E_"))
leaf.voom5.sample %>% group_by(genotype) %>% summarise(num=n())
```

# plotting expresion patterns
```{r}
# BraA01g010990.3C is ATIRT1
plotgene <- function(cpm=root.voom5, target="BraA01g010990.3C",samples=root.voom5.sample) {
  print(colnames(samples))
  print(unique(samples$tissue))
  data <- cpm %>% filter(genes %in% target) %>% 
    pivot_longer(cols=-genes,names_to="sampleID",values_to="log2value") %>%
    inner_join(samples,by=c("sampleID"=colnames(samples)[1]))
  print(data)
  plot <- data %>%
    ggplot(aes(y=log2value,x=trt,color=trt,shape=genotype)) + geom_jitter() +
    facet_wrap(genotype~genes, ncol=4, scales = "free_y") +
    theme(axis.text.x = element_text(angle = 90, hjust=1)) + labs(title=unique(samples$tissue))
  return(plot)
}

plotgene() # BraA01g010990.3C is ATIRT1

plotgene(cpm=leaf.voom5,samples=leaf.voom5.sample) # error due to not in leaf data
plotgene(cpm=leaf.voom5,samples=leaf.voom5.sample,target=leaf.voom5$genes[1]
) # no error. 
```

# leaf_tissue.root.yellowgreen (40 genes)
```{r}
GSEA.leaf_tissue.root.yellowgreen.coregenes.yes <- coregenes.yes.DF %>% filter(leaf_tissue.root.yellowgreen==1) %>% dplyr::select(name) %>% as_vector()
p <- plotgene(target=GSEA.leaf_tissue.root.yellowgreen.coregenes.yes)
ggsave(p,filename="GSEA.leaf_tissue.root.yellowgreen.yes.plot.png",path=file.path("RNA","output"),height=50,width=8,limitsize=FALSE)
```

# root_tissue.leaf.lightsteelblue1 (54 genes)
```{r}
GSEA.root_tissue.leaf.lightsteelblue1.coregenes.yes <- coregenes.yes.DF %>% filter(root_tissue.leaf.lightsteelblue1==1) %>% dplyr::select(name) %>% as_vector()
p <- plotgene(target=GSEA.root_tissue.leaf.lightsteelblue1.coregenes.yes)
ggsave(p,filename="GSEA.root_tissue.leaf.lightsteelblue1.yes.plot.png",path=file.path("RNA","output"),height=50,width=8,limitsize=FALSE)
```
# root_tissue.leaf.midnightblue (140 ngenes)
```{r}
GSEA.root_tissue.leaf.midnightblue.coregenes.yes <- coregenes.yes.DF %>% filter(root_tissue.leaf.midnightblue==1) %>% dplyr::select(name) %>% as_vector()
p <- plotgene(target=GSEA.root_tissue.leaf.midnightblue.coregenes.yes)
ggsave(p,filename="GSEA.root_tissue.leaf.midnightblue.yes.plot.png",path=file.path("RNA","output"),height=50,width=8,limitsize=FALSE)
```

# root_tissue.leaf.thistle2 (31 genes)
```{r}
GSEA.root_tissue.leaf.thistle2.coregenes.yes <- coregenes.yes.DF %>% filter(root_tissue.leaf.thistle2==1) %>% dplyr::select(name) %>% as_vector()
p <- plotgene(target=GSEA.root_tissue.leaf.thistle2.coregenes.yes)
ggsave(p,filename="GSEA.root_tissue.leaf.thistle2.coregenes.yes.plot.png",path=file.path("RNA","output"),height=50,width=8,limitsize=FALSE)
```


# root_tissue.leaf.red (223 genes)
```{r}
GSEA.root_tissue.leaf.red.coregenes.yes <- coregenes.yes.DF %>% filter(root_tissue.leaf.red==1) %>% dplyr::select(name) %>% as_vector()
p <- plotgene(target=GSEA.root_tissue.leaf.red.coregenes.yes)
ggsave(p,filename="GSEA.root_tissue.leaf.red.coregenes.yes.plot.png",path=file.path("RNA","output"),height=50,width=8,limitsize=FALSE)
```

# root_tissue.root.plum1
```{r}
GSEA.root_tissue.root.plum1.coregenes.yes <- coregenes.yes.DF %>% filter(root_tissue.root.plum1==1) %>% dplyr::select(name) %>% as_vector()
p <- plotgene(target=GSEA.root_tissue.root.plum1.coregenes.yes)
ggsave(p,filename="GSEA.root_tissue.root.plum1.coregenes.yes.plot.png",path=file.path("RNA","output"),height=50,width=8,limitsize=FALSE)

```


# leaf_tissue.root.yellowgreen (76 genes)
```{r}
GSEA.leaf_tissue.root.yellowgreen.coregenes.yes <- coregenes.yes.DF %>% filter(leaf_tissue.root.yellowgreen==1) %>% dplyr::select(name) %>% as_vector()
p <- plotgene(target=GSEA.leaf_tissue.root.yellowgreen.coregenes.yes)
ggsave(p,filename="GSEA.leaf_tissue.root.yellowgreen.coregenes.yes.plot.png",path=file.path("RNA","output"),height=50,width=8,limitsize=FALSE)
```





# all modules with only target module 
* copied from 11_machine_learning.Rmd
leafdata (copied from 11_machine_learning.Rmd)
```{r}
leaf <- read_csv("plant/output/leaf_lengths_combined.csv") %>%
  mutate(sampleID=str_remove(sampleID, "_(leaf|root)"),
         sampleID=str_remove(sampleID, "^[a-h][1-8]_")) %>%
  filter(!duplicated(sampleID))

leaf %>% arrange(sampleID)
```

load root modules (copied from 11_machine_learning.Rmd)
```{r}
root.voom5 <- readr::read_tsv(file.path("RNA","output","voom_expression.e1and3.resid.exp_gt.root.plusR500.txt.gz"))
load(file.path("RNA","output","voom_expression.e1and3.resid.exp_gt.root.plusR500.txt.gz.WGCNA.softthresh10.RData"))
# adding module column
#root.voom5.mod <- root.voom5 %>% mutate(dat.voom.module=str_c("ME",dynamicColors,"_root"))
root.voom5.mod <- root.voom5 %>% mutate(dat.voom.module=str_c("root.",dynamicColors))

# MEs
rootMEs <- MEs %>% as_tibble() %>%
  rename_all(str_c, "_root") %>%
  mutate(sampleID=colnames(root.voom5)[-1],
         sampleID=str_remove(sampleID, "_root"),
         sampleID=str_remove(sampleID, "_S[0-9]{1,3}$"),
         sampleID=str_remove(sampleID, "^[a-h][1-8]_")) %>%
  dplyr::select(sampleID, everything())
rootMEs
rm(MEs)
```

load leaf modules (copied from 11_machine_learning.Rmd)
```{r}
leaf.voom5 <- readr::read_tsv(file.path("RNA","output","voom_expression.e1and3.resid.exp_gt.leaf.plusR500.txt.gz"))
load(file.path("RNA","output","voom_expression.e1and3.resid.exp_gt.leaf.plusR500.txt.gz.WGCNA.softthresh10.RData"))
# adding module column (tissue.module all lowercase that is consistent with )
#leaf.voom5.mod <- leaf.voom5 %>% mutate(dat.voom.module=str_c("ME",dynamicColors,"_leaf"))
leaf.voom5.mod <- leaf.voom5 %>% mutate(dat.voom.module=str_c("leaf.",dynamicColors))

# 
leafMEs <- MEs %>% as_tibble() %>%
  rename_all(str_c, "_leaf") %>%
  mutate(sampleID=colnames(leaf.voom5)[-1],
         sampleID=str_remove(sampleID, "_leaf"),
         sampleID=str_remove(sampleID, "_S[0-9]{1,3}$"),
         sampleID=str_remove(sampleID, "^[a-h][1-8]_")) %>%
  dplyr::select(sampleID, everything())
leafMEs
rm(MEs)
```

bring it all together (copied from 11_machine_learning.Rmd)
```{r}
d <- leaf %>% dplyr::select(sampleID, leaf_avg_std) %>%
  inner_join(leafMEs, by="sampleID") %>% 
  inner_join(rootMEs, by="sampleID") %>% 
  # dplyr::select(-sampleID) %>% # Kazu need sampleID later
  as_data_frame()
d
```

# plot
* see Julin's request (https://trello.com/c/Y4GWD7KJ/35-wgcna-and-mr-analysis)
```{r}
# new ver (tissue.module lowercase)
ElasticNet0.5_multiCV_Pct_var.module <- ElasticNet0.5_multiCV_Pct_var %>% dplyr::select(tissue.module) %>% as_vector()
# new ver. use anti_join()
nonElasticNet0.5_multiCV_Pct_var.module <- tibble(all = colnames(d)[-1:-2]) %>% separate(all, c("module","tissue")) %>% mutate(module=str_remove(module,"ME")) %>% unite(tissue.module.all,tissue,module,sep=".") %>% anti_join(ElasticNet0.5_multiCV_Pct_var,by=c("tissue.module.all"="tissue.module")) %>% as_vector()
# 
mod <- ElasticNet0.5_multiCV_Pct_var.module[3]
# making expression pattern plot function (compare within/outside module)
#
d.root <- d %>% dplyr::select(sampleID,leaf_avg_std,ends_with("root"))
d.leaf <- d %>% dplyr::select(sampleID,leaf_avg_std,ends_with("leaf"))
# 
expression.module4 <- function(target.module=mod,dat.voom=root.voom5.mod,tissue_MEs=d.root) { 
  print(str_c("The target.module is ",target.module,"."))
# add within target.module info
dat.voom <- dat.voom %>% mutate(TF=ifelse(dat.voom.module==target.module,"within the module","not in the module"),.keep="unused") %>% mutate(TF.module=str_c(target.module,"\n",TF))
# number of genes with/not in the module
dat.voom %>% group_by(TF) %>% summarize(nTF=n())
# longer format
dat.voom.long <- dat.voom %>% pivot_longer(-c(genes,TF,TF.module),names_to="sampleID",values_to="value") %>% mutate(sampleID=str_remove(sampleID,"_root")) %>% mutate(sampleID=str_remove(sampleID,"_leaf"))
# modify for new ME
temp.target.module <- tibble(tissue_MEs=colnames(tissue_MEs)[-2:-1]) %>% mutate(tissue_MEs = str_remove(tissue_MEs,"^ME")) %>% separate(tissue_MEs,c("module","tissue")) %>% unite(tissue.module,tissue,module,sep=".") %>% as_vector()
# change format of colnames of tissue_MEs
colnames(tissue_MEs)[-2:-1] <- temp.target.module
# get sample levels by leaf_avg_std
sample.levels <- tissue_MEs %>% dplyr::select(sampleID, leaf_avg_std,quo_name(target.module)) %>% arrange_at(.vars=2) %>% dplyr::select(sampleID) %>% as_vector()
# add expression data 
dat.voom.ME <- tissue_MEs %>% dplyr::select(sampleID,quo_name(target.module))  %>% inner_join(dat.voom.long,by="sampleID") %>%
  arrange_at(.vars=c(2,4)) %>% mutate(sampleID=factor(sampleID,levels=sample.levels))
# plot 
print(dat.voom.ME)
# but aes_() will be soft deprecated, so that need to use quasiquotation (!!, !!!, :=)  (https://www.tidyverse.org/blog/2018/07/ggplot2-tidy-evaluation/)
# https://adv-r.hadley.nz/quasiquotation.html
# plot expression value plus ME within the module
p.TF10.withinmod <- dat.voom.ME  %>%  filter(TF=="within the module") %>% ggplot(aes(x=sampleID,y=value,color=TF,group=genes)) + geom_line(alpha=0.1)  + geom_line(aes_string(x="sampleID",y=target.module,group="genes"),color="black",size=1) + theme(axis.text.x=element_text(angle=90),legend.position="none")  + facet_grid(.~TF.module) +  labs(ylab="normalized expression") #title=str_remove(target.module,"ME"),
# plot expression value not in the module
p.TF10.notinmod <- dat.voom.ME  %>%  filter(TF=="not in the module") %>% ggplot(aes(x=sampleID,y=value,color=TF,group=genes)) + geom_line(alpha=0.1)  +  theme(axis.text.x=element_text(angle=90),legend.position="none")  + facet_grid(.~TF.module) +  labs(ylab="normalized expression") #title=str_remove(target.module,"ME"),
# combine the two plots
p.TFall <- plot_grid(p.TF10.withinmod,p.TF10.notinmod)
return(p.TFall)
}
# test 
mod=ElasticNet0.5_multiCV_Pct_var.module[1]
expression.module4()
# 
plot.root.list4 <- tibble(tissue_MEs=colnames(d)[-2:-1]) %>% mutate(tissue_MEs = str_remove(tissue_MEs,"^ME")) %>% separate(tissue_MEs,c("module","tissue")) %>% unite(tissue.module,tissue,module,sep=".") %>% filter(str_detect(tissue.module,"^root")) %>% split(.$tissue.module) %>% map(~expression.module4(target.module=as.name(.$tissue.module),dat.voom=root.voom5.mod,tissue_MEs=d.root)) # takes minutes

plot.leaf.list4 <- tibble(tissue_MEs=colnames(d)[-2:-1]) %>% mutate(tissue_MEs = str_remove(tissue_MEs,"^ME")) %>% separate(tissue_MEs,c("module","tissue")) %>% unite(tissue.module,tissue,module,sep=".") %>% filter(str_detect(tissue.module,"^leaf")) %>% split(.$tissue.module) %>% map(~expression.module4(target.module=as.name(.$tissue.module),dat.voom=leaf.voom5.mod,tissue_MEs=d.leaf)) # takes minutes
# ElasticNet0.5_multiCV_Pct_var plots
plot.root.list4[[ElasticNet0.5_multiCV_Pct_var.module[1]]]
plot.root.list4[[ElasticNet0.5_multiCV_Pct_var.module[2]]]
plot.root.list4[[ElasticNet0.5_multiCV_Pct_var.module[3]]]
# nonElasticNet0.5_multiCV_Pct_var plots
plot.root.list4[[nonElasticNet0.5_multiCV_Pct_var.module[1]]]
plot.root.list4[[nonElasticNet0.5_multiCV_Pct_var.module[2]]]
plot.root.list4[[nonElasticNet0.5_multiCV_Pct_var.module[3]]]
```

# how to write figs for plots?
```{r}
p.root.all <- plot_grid(plot.root.list4[[1]],plot.root.list4[[2]],plot.root.list4[[3]],plot.root.list4[[4]],
                   plot.root.list4[[5]],plot.root.list4[[6]],plot.root.list4[[7]],plot.root.list4[[8]],
                   plot.root.list4[[9]],plot.root.list4[[10]],plot.root.list4[[11]],plot.root.list4[[12]],
                   plot.root.list4[[13]],plot.root.list4[[14]],plot.root.list4[[15]],plot.root.list4[[16]],
                   plot.root.list4[[17]],plot.root.list4[[18]],plot.root.list4[[19]],plot.root.list4[[20]],
                   plot.root.list4[[21]],plot.root.list4[[22]])
ggsave(p.root.all,path=file.path("RNA","output"),filename="p.root.all2.png",height=40,width=20,dpi=72)
p.leaf.all <- plot_grid(plot.leaf.list4[[1]],plot.leaf.list4[[2]],plot.leaf.list4[[3]],plot.leaf.list4[[4]],
                   plot.leaf.list4[[5]],plot.leaf.list4[[6]],plot.leaf.list4[[7]],plot.leaf.list4[[8]],
                   plot.leaf.list4[[9]],plot.leaf.list4[[10]],plot.leaf.list4[[11]],plot.leaf.list4[[12]],
                   plot.leaf.list4[[13]],plot.leaf.list4[[14]],plot.leaf.list4[[15]],plot.leaf.list4[[16]],
                   plot.leaf.list4[[17]],plot.leaf.list4[[18]],plot.leaf.list4[[19]],plot.leaf.list4[[20]],
                   plot.leaf.list4[[21]],plot.leaf.list4[[22]],plot.leaf.list4[[23]],plot.leaf.list4[[24]],
                   plot.leaf.list4[[25]],plot.leaf.list4[[26]],plot.leaf.list4[[27]],plot.leaf.list4[[28]],
                   plot.leaf.list4[[29]],plot.leaf.list4[[30]],plot.leaf.list4[[31]],plot.leaf.list4[[32]],
                  plot.leaf.list4[[33]],plot.leaf.list4[[34]])
ggsave(p.leaf.all,path=file.path("RNA","output"),filename="p.leaf.all2.png",height=40,width=20,dpi=72)

# only leaf length associated modules
p.root.leaf_length.associated <- plot_grid(plot.root.list4[[ElasticNet0.5_multiCV_Pct_var.module[1]]],
                                           plot.root.list4[[ElasticNet0.5_multiCV_Pct_var.module[2]]],
                                           plot.root.list4[[ElasticNet0.5_multiCV_Pct_var.module[3]]],
                                           plot.root.list4[[ElasticNet0.5_multiCV_Pct_var.module[5]]])
ggsave(p.root.leaf_length.associated,path=file.path("RNA","output"),filename="p.root.leaf_length.associated2.png",height=11,width=8,dpi=72)

p.leaf.leaf_length.associated <- plot_grid(plot.leaf.list4[[ElasticNet0.5_multiCV_Pct_var.module[4]]],
                                           plot.leaf.list4[[ElasticNet0.5_multiCV_Pct_var.module[6]]],
                                           plot.leaf.list4[[ElasticNet0.5_multiCV_Pct_var.module[7]]],
                                           plot.leaf.list4[[ElasticNet0.5_multiCV_Pct_var.module[8]]],
                                           plot.leaf.list4[[ElasticNet0.5_multiCV_Pct_var.module[9]]])
ggsave(p.leaf.leaf_length.associated,path=file.path("RNA","output"),filename="p.leaf.leaf_length.associated2.png",height=11,width=12,dpi=72)

```


