---
title: "11_machine_learning_take2"
author: "Julin Maloof"
date: "01/21/2021"
output: html_notebook
---

The goal of this file is to find which WGCNA modules are best able to predict leaf length.

Here I modify the original script to use a multi-CV approach that I developed from the metabolites.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(glmnet)
library(relaimpo)
library(tidyverse)
```

## get data

leafdata
```{r}
leaf <- read_csv("../../plant/output/leaf_lengths_combined.csv") %>%
  mutate(sampleID=str_remove(sampleID, "_(leaf|root)"),
         sampleID=str_remove(sampleID, "^[a-h][1-8]_")) %>%
  filter(!duplicated(sampleID))

leaf %>% arrange(sampleID)
```

load root modules
```{r}
root.voom5 <- readr::read_tsv(file.path("..","output","voom_expression.e1and3.resid.exp_gt.root.plusR500.txt.gz"))
load(file.path("..","output","voom_expression.e1and3.resid.exp_gt.root.plusR500.txt.gz.WGCNA.softthresh10.RData"))

rootMEs <- MEs %>% as_tibble() %>%
  rename_all(str_c, "_root") %>%
  mutate(sampleID=colnames(root.voom5)[-1],
         sampleID=str_remove(sampleID, "_root"),
         sampleID=str_remove(sampleID, "_S[0-9]{1,3}$"),
         sampleID=str_remove(sampleID, "^[a-h][1-8]_")) %>%
  select(sampleID, everything())
rootMEs
rm(MEs)
```

load leaf modules
```{r}
leaf.voom5 <- readr::read_tsv(file.path("..","output","voom_expression.e1and3.resid.exp_gt.leaf.plusR500.txt.gz"))
load(file.path("..","output","voom_expression.e1and3.resid.exp_gt.leaf.plusR500.txt.gz.WGCNA.softthresh10.RData"))

leafMEs <- MEs %>% as_tibble() %>%
  rename_all(str_c, "_leaf") %>%
  mutate(sampleID=colnames(leaf.voom5)[-1],
         sampleID=str_remove(sampleID, "_leaf"),
         sampleID=str_remove(sampleID, "_S[0-9]{1,3}$"),
         sampleID=str_remove(sampleID, "^[a-h][1-8]_")) %>%
  select(sampleID, everything())
leafMEs
rm(MEs)
```

bring it all together
```{r}
d <- leaf %>% select(sampleID, leaf_avg_std) %>%
  inner_join(leafMEs, by="sampleID") %>% 
  inner_join(rootMEs, by="sampleID") %>% 
  select(-sampleID) %>%
  as_data_frame()
d
```

```{r}
apply(d, 2, mean)
apply(d, 2, sd)
```


## elastic net regression

### Use a multi-cross validation to compare different alphas

With a small number of samples, there is a lot of run-to-run variation in cross validation attempts.  I will do many and average over them to choose an appropriate alpha and lambda.

## multi CV

set up the data
```{r}
x <- d %>% select(starts_with("ME")) %>% as.matrix()
y <- d %>% pull(leaf_avg_std)
```


Fit 101 4 fold CVs for each of 11 alphas
```{r}
set.seed(1245)

folds <- tibble(run=1:101) %>% 
  mutate(folds=map(run, ~ sample(rep(1:5,8))))

system.time (multiCV <- expand_grid(run=1:100, alpha=seq(0,1,.1)) %>%
               left_join(folds, by="run") %>%
               mutate(fit=map2(folds, alpha, ~ cv.glmnet(x=x, y=y, foldid = .x, alpha=.y
                                                         )))
             #, lambda=exp(seq(-5,0,length.out = 50)) )))
) #100 seconds

head(multiCV)
```

for each fit, pull out the mean cv error, lambda, min lambda, and 1se lambda 
```{r}
multiCV <- multiCV %>%
  mutate(cvm=map(fit, magrittr::extract("cvm")),
         lambda=map(fit, magrittr::extract("lambda")),
         lambda.min=map_dbl(fit, magrittr::extract("lambda.min" )),
         lambda.1se=map_dbl(fit, magrittr::extract("lambda.1se")),
         nzero=map(fit, magrittr::extract("nzero"))
  )

head(multiCV)
```


now calculate the mean and sem of cvm and min,1se labmdas.  These need to be done separately because of the way the grouping works
```{r}
summary_cvm <- multiCV %>% dplyr::select(-fit, -folds) %>% 
  unnest(c(cvm, lambda)) %>%
  group_by(alpha, lambda) %>%
  summarize(meancvm=mean(cvm), sem=sd(cvm)/sqrt(n()), high=meancvm+sem, low=meancvm-sem)

summary_cvm
```

```{r}
summary_lambda <- multiCV %>% dplyr::select(-fit, -folds, -cvm) %>% 
  group_by(alpha) %>%
  summarize(
    lambda.min.sd=sd(lambda.min), 
    lambda.min.mean=mean(lambda.min),
    #lambda.min.med=median(lambda.min), 
    lambda.min.high=lambda.min.mean+lambda.min.sd,
    #lambda.min.low=lambda.min.mean-lambda.min.sem,
    #lambda.1se.sem=sd(lambda.1se)/sqrt(n()), 
    lambda.1se.mean=mean(lambda.1se),
    #lambda.1se.med=median(lambda.1se), 
    #lambda.1se.high=lambda.1se+lambda.1se.sem,
    #lambda.1se.low=lambda.1se-lambda.1se.sem,
    nzero=nzero[1],
    lambda=lambda[1]
  )

summary_lambda
```


plot it
```{r}
summary_cvm %>%
  #filter(alpha!=0) %>% # worse than everything else and throwing the plots off
  ggplot(aes(x=log(lambda), y= meancvm,  ymin=low, ymax=high)) +
  geom_ribbon(alpha=.25) +
  geom_line(aes(color=as.character(alpha))) +
  facet_wrap(~ as.character(alpha)) +
   coord_cartesian(xlim=(c(-5,0))) +
  geom_vline(aes(xintercept=log(lambda.min.mean)), alpha=.5, data=summary_lambda) +
  geom_vline(aes(xintercept=log(lambda.min.high)), alpha=.5, data=summary_lambda, color="blue") 

```

Min MSE per alpha:

```{r}
summary_cvm %>% 
  group_by(as.factor(alpha)) %>%
  filter(rank(meancvm, ties.method="first")==1) %>%
  ggplot(aes(x=alpha,y=meancvm,ymin=low,ymax=high)) +
  geom_ribbon(color=NA, fill="gray80") +
  geom_line() +
  geom_point()
```

Not a huge amount of difference there.  Obviously alpha=0 is "best" but I want to do variable selection so will choose something higher.

Make a plot of number of nzero coefficients

```{r}
summary_lambda %>%
  unnest(c(lambda, nzero)) %>%
  group_by(alpha) %>%
  filter(abs(lambda.min.mean-lambda)==min(abs(lambda.min.mean-lambda))  ) %>%
  ungroup() %>%

ggplot(aes(x=as.character(alpha), y=nzero)) +
  geom_point() +
  ggtitle("Number of non-zero coefficents at minimum lambda") +
  ylim(0,24)
```

Let-s compromise and choose alpha = .5

Look at the actual fit

```{r}
bestlam_multiCV <- summary_lambda %>% filter(round(alpha,1)==.5) %>% pull(lambda.min.mean)
fit <- multiCV %>% filter(round(alpha,1)==.5, run==1) %>% pull(fit) %>% magrittr::extract2(1)
```


```{r}
multiCV_coef.tb <- coef(fit, s=bestlam_multiCV) %>% 
  as.matrix() %>% as.data.frame() %>% 
  rownames_to_column(var="ME") %>%
  dplyr::rename(beta_multiCV=`1`)
  
multiCV_coef.tb %>% filter(beta_multiCV!=0) %>% arrange(beta_multiCV)

```

pred and obs
```{r}
pred <- predict(fit, s=bestlam_multiCV, newx=x)
plot(y,pred)
cor.test(y, pred) #.73
mean((y-pred)^2) #0.64
```

check some train/test

```{r}
testtrain_multiCV <- tibble(run=1:10000) %>%
  mutate(train=map(run, ~ sample(1:40,8))) %>%
  mutate(fit=map(train, ~ glmnet(x=x[.,], y=y[.], lambda = bestlam_multiCV, 
                                 alpha = .5 ) )) %>%
  mutate(pred=map2(fit, train, ~ predict(.x, newx = x[-.y,]))) %>%
  mutate(cor=map2_dbl(pred, train, ~ cor(.x, y[-.y])  )) %>%
  mutate(MSE=map2_dbl(pred, train, ~ mean((y[-.y] - .x)^2)))

sum(testtrain_multiCV$cor<=0, na.rm = TRUE) + sum(is.na(testtrain_multiCV$cor))
mean(testtrain_multiCV$cor, na.rm=TRUE)
mean(testtrain_multiCV$cor^2, na.rm=TRUE)
mean(testtrain_multiCV$MSE)
hist(testtrain_multiCV$cor)
hist(testtrain_multiCV$cor^2)
t.test(testtrain_multiCV$cor)
```


```{r}
vars <- multiCV_coef.tb %>% 
  filter(beta_multiCV !=0, ME!="(Intercept)") %>%
  pull(ME) %>% c("leaf_avg_std", .)

relimpmultiCV <- d %>% dplyr::select(all_of(vars)) %>%
  calc.relimp() 

coef.tb <- relimpmultiCV@lmg %>% as.matrix() %>% as.data.frame() %>%
  rownames_to_column("ME") %>%
  rename(Prop_Var_Explained=V1) %>%
  full_join(multiCV_coef.tb) %>%
  arrange(desc(Prop_Var_Explained))

coef.tb <- coef.tb %>% arrange(desc(Prop_Var_Explained)) %>% 
  filter(!is.na(Prop_Var_Explained)) %>%
  mutate(Pct_Var_Explained=round(100*Prop_Var_Explained,1),
         Effect=ifelse(beta_multiCV>0, "Increase", "Decrease"),
         Module_Organ=str_extract(ME,"root|leaf"),
         Module_Name=str_replace(ME,"^ME(.*)_(leaf|root)$", "\\1")) %>%
  
  select(Module_Name, Module_Organ, Pct_Var_Explained, Effect) 

coef.tb %>% write_csv("../output/ElasticNet0.5_multiCV_Pct_var.csv")

coef.tb %>% write_csv("../../_Paper_Tables/ElasticNetWGCNA_Modules.csv")

relimpmultiCV

coef.tb
```
