---
title: "15_Hub_Genes"
author: "Julin Maloof"
date: "9/26/2024"
output: html_document
editor_options: 
  chunk_output_type: inline
---

Looking for hub genes in interesting modules.

Methods: can you standard network concepts, like degree centrality and betweeness centrality.  Alternatively, can use module membership (MM) from WGCNA and / or correlation with trait. 

```{r}
library(WGCNA)
library(tidyverse)
library(igraph)
```

NOTE: the adjacency and TOM matrices are too big for git.  If you need them, email jnmaloof@ucdavis or run the code at `/RNA/scripts/10_WGCNA_including_R500_signedhybrid_moduleColors.Rmd`

## Root

```{r}
load("../output/voom_expression.e1and3.resid.exp_gt.root.plusR500.signedhybrid.txt.gz.WGCNA.softthresh10.signedhybrid.RData")

load("../output/voom_expression.e1and3.resid.exp_gt.root.plusR500.signedhybrid.txt.gz.WGCNA.softthresh10.signedhybrid.adjacencyTOM.Rdata")
```


### Ivory
```{r}
root.ivory.adjacency <- adjacency[moduleColors=="ivory", moduleColors=="ivory"]
root.ivory.TOM <- TOM[moduleColors=="ivory", moduleColors=="ivory"]

root.ivory.adjacency[1:5, 1:5]

diag(root.ivory.adjacency) <- 0
diag(root.ivory.TOM) <- 0

hist(root.ivory.adjacency)
hist(log10(root.ivory.adjacency))
hist(root.ivory.TOM)
```

export to cytoscape
```{r}
# dir.create("../output/cytoscape")
# adjacency 

sum(root.ivory.adjacency>=.1) #300

exportNetworkToCytoscape(root.ivory.adjacency, 
                         edgeFile = "../output/cytoscape/root.ivory.a0.1.edge.txt",
                         nodeFile = "../output/cytoscape/root.ivory.a0.1.node.txt",
                         threshold = 0.1)
# TOM
sum(root.ivory.TOM>=.05) #420

exportNetworkToCytoscape(root.ivory.adjacency, 
                         edgeFile = "../output/cytoscape/root.ivory.T0.05.edge.txt",
                         nodeFile = "../output/cytoscape/root.ivory.T0.05.node.txt",
                         threshold = 0.05)

```


calculate connectivity
```{r}
root.ivory.connectivity <- root.ivory.adjacency %>% colSums() %>% sort(decreasing = TRUE) 
root.ivory.connectivity

as_tibble(root.ivory.connectivity)

tibble(geneID=names(root.ivory.connectivity), WGCNA_connectivity=root.ivory.connectivity) %>% write_tsv(
            file = "../output/cytoscape/root.ivory.connectivity.txt")
```
### Bisque 4

```{r}
root.bisque4.adjacency <- adjacency[moduleColors=="bisque4", moduleColors=="bisque4"]
root.bisque4.TOM <- TOM[moduleColors=="bisque4", moduleColors=="bisque4"]

diag(root.bisque4.adjacency) <- 0
diag(root.bisque4.TOM) <- 0

hist(root.bisque4.adjacency)
hist(root.bisque4.TOM)

dim(root.bisque4.adjacency)
```

export to cytoscape
```{r}
# dir.create("../output/cytoscape")
# adjacency 

sum(root.bisque4.adjacency>=.1) #368

exportNetworkToCytoscape(root.bisque4.adjacency, 
                         edgeFile = "../output/cytoscape/root.bisque4.a0.1.edge.txt",
                         nodeFile = "../output/cytoscape/root.bisque4.a0.1.node.txt",
                         threshold = 0.1)
# TOM
sum(root.bisque4.TOM>=.05) #256

exportNetworkToCytoscape(root.bisque4.adjacency, 
                         edgeFile = "../output/cytoscape/root.bisque4.T0.05.edge.txt",
                         nodeFile = "../output/cytoscape/root.bisque4.T0.05.node.txt",
                         threshold = 0.05)

```


## leaf
   
### leaf.darkseagreen4
   

```{r}
load("../output/voom_expression.e1and3.resid.exp_gt.leaf.plusR500.signedhybrid.txt.gz.WGCNA.softthresh10.signedhybrid.RData")

load("../output/voom_expression.e1and3.resid.exp_gt.leaf.plusR500.signedhybrid.txt.gz.WGCNA.softthresh10.signedhybrid.adjacencyTOM.Rdata")
```


### Darkseagreen4

```{r}
leaf.darkseagreen4.adjacency <- adjacency[moduleColors=="darkseagreen4", moduleColors=="darkseagreen4"]
leaf.darkseagreen4.TOM <- TOM[moduleColors=="darkseagreen4", moduleColors=="darkseagreen4"]

leaf.darkseagreen4.adjacency[1:5, 1:5]

diag(leaf.darkseagreen4.adjacency) <- 0
diag(leaf.darkseagreen4.TOM) <- 0

hist(leaf.darkseagreen4.adjacency)
hist(log10(leaf.darkseagreen4.adjacency))
hist(leaf.darkseagreen4.TOM)
```

export to cytoscape
```{r}
# dir.create("../output/cytoscape")
# adjacency 

dim(leaf.darkseagreen4.adjacency) #99

sum(leaf.darkseagreen4.adjacency>=.075) #284

exportNetworkToCytoscape(leaf.darkseagreen4.adjacency, 
                         edgeFile = "../output/cytoscape/leaf.darkseagreen4.a0.1.edge.txt",
                         nodeFile = "../output/cytoscape/leaf.darkseagreen4.a0.1.node.txt",
                         threshold = 0.075)
# TOM
sum(leaf.darkseagreen4.TOM>=.05) #162

exportNetworkToCytoscape(leaf.darkseagreen4.adjacency, 
                         edgeFile = "../output/cytoscape/leaf.darkseagreen4.T0.05.edge.txt",
                         nodeFile = "../output/cytoscape/leaf.darkseagreen4.T0.05.node.txt",
                         threshold = 0.05)

```
   