---
title: "PMN_Pathway Collage from an Omics Dataset"
author: "Kazunari Nozue"
date: "2/15/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# [Pathway Collage from an Omics Dataset](https://pmn.plantcyc.org/PToolsWebsiteHowto.shtml#celOverview)

```{r}
rm(list=ls())
library(tidyverse);library(readr);library(readxl);library(openxlsx)
# The following setting is important for WGCNA, do not omit.
options(stringsAsFactors = FALSE)
# install.packages(c("dynamicTreeCut", "cluster", "flashClust", "Hmisc", "reshape", "foreach", "doParallel") ) 
library(WGCNA);library(flashClust) # for WGCNA in Whitney
#allowWGCNAThreads(nThreads = NULL) # for Whitney
#enableWGCNAThreads(4) # for Whitney

```

# Automated + mannual compound name curration for PMN
* automated curation by PMN (imported to SmartTable)
* manual curation (added to the imported SmartTable) by using search engine in PMN and "add" function.
* summarized in "new_unique_associated_metabolites_02_09_21_PMNcuration.xlsx"

```{r}
mets.PMN <- read_xlsx("../output/new_unique_associated_metabolites_02_09_21_PMNcuration.xlsx")
mets.PMN <- mets.PMN %>% filter(!Compound_id=="0")

```

# read csv file
```{r}
met_amt_rotation_out <- read_csv("../output/Leaf_associated_metabolites_raw_NOBLANK.csv")

```

# combine metabolite conversion table, which is readable by PMN
```{r}
met_amt_rotation_out %>% inner_join(mets.PMN,by="metabolite") %>% dplyr::select(Compound_id,PC,loading) %>% #group_by(PC) %>% summarize(n())
  filter(PC=="root_PC3") %>% write_tsv(file.path("..","output","Leaf_associated_metabolites_raw_NOBLANK.PMN.root_PC3.tsv"))
```

# How to map the metabolite data onto PMN cellular view? (under construction)
* https://pmn.plantcyc.org/overviewsWeb/celOv.shtml?orgid=BRAPA_FPSC
* Operations > Overlay Experimental Data (Omics Viewer) > Upload Data from File
* Choose File, "Leaf_associated_metabolites_raw_NOBLANK.PMN.root_PC3.tsv"
* Select "Compound names and/or identifiers" under "Items in the first column of the file are:"
* Input "2" in "Data columns to use"
* "Relative" in "Select type of value"
* "0-centered scale" for Data values uses a 
* "No ratio of data columns"
* Show data: "As a Pathway Collage"

# Pathways to genes tables
# from compounds to genes
```{r}
#brapa_fpsccyc_fast <- read_csv(file.path("..","..","..","Annotation","input","brapa_fpsccyc.fasta"))
# compounds
brapa_fpsccyc_compounds.20180702 <- read_tsv(file.path("..","..","annotation","input","PMN","brapa_fpsccyc_compounds.20180702"))
write_csv(brapa_fpsccyc_compounds.20180702,file=file.path("..","..","annotation","input","PMN","brapa_fpsccyc_compounds.20180702.csv"))
brapa_fpsccyc_compounds.20180702
# pathways
brapa_fpsccyc_pathways.20180702 <- read_tsv(file.path("..","..","annotation","input","PMN","brapa_fpsccyc_pathways.20180702"))
brapa_fpsccyc_pathways.20180702
```

# combine them
```{r}
brapa_fpsccyc_compounds.20180702_to_genes <- brapa_fpsccyc_compounds.20180702 %>% inner_join(brapa_fpsccyc_pathways.20180702,by="EC") #%>% #View()
#write_csv(brapa_fpsccyc_compounds.20180702_to_genes,file=file.path("..","output","brapa_fpsccyc_compounds.20180702_to_genes.csv.gz"))
# 
brapa_fpsccyc_compounds.20180702_to_genes %>% group_by(Compound_id) %>% summarize(num=n())
```


# comparison PMN compund names and Julin's ML_leaf_associated_normalized_metabolites 
```{r}
met_amt_rotation_out.brapa_fpsccyc_compounds.20180702_to_genes <- met_amt_rotation_out %>% inner_join(mets.PMN,by="metabolite") %>% dplyr::select(Compound_id,PC,loading) %>% mutate(Compound_id=str_to_upper(Compound_id)) %>%  inner_join(brapa_fpsccyc_compounds.20180702_to_genes,by="Compound_id") 
met_amt_rotation_out.brapa_fpsccyc_compounds.20180702_to_genes %>% View()
# write csv
write_csv(met_amt_rotation_out.brapa_fpsccyc_compounds.20180702_to_genes,file.path("..","output","met_amt_rotation_out.brapa_fpsccyc_compounds.20180702_to_genes.csv"))
# how many metabolites in met_amt_rotation_out.brapa_fpsccyc_compounds.20180702_to_genes?
# the original metabolite
met_amt_rotation_out %>% group_by(metabolite) %>% summarise(n=n()) # 77
# met_amt_rotation_out (corresponding "Compound_id" in PMN)
met_amt_rotation_out %>% inner_join(mets.PMN,by="metabolite") %>% dplyr::select(Compound_id,PC,loading) %>% group_by(Compound_id) %>% summarise(n=n()) # 59
# to_genes (something is wrong)
met_amt_rotation_out.brapa_fpsccyc_compounds.20180702_to_genes %>% group_by(Compound_id) %>% summarise(n=n()) # 14 compound_id ... too small. why?????
```

# How to add omics data to a Pathway Collage
* File > Add or replace omics data
* Overlay our RNAse data. Which? (option 1) live/dead relative expression values of soil treatment associated genes ("Yes" in GSEA) among leaf length correlated modules.(option 2)

# Julin's Machine Learning selected modules
```{r eval=FALSE}
# reading Julin's results of penalized regression (see slide) (old version)
# ElasticNet0.5_LOO_prop_var <- read_csv(file.path("plant","output","ElasticNet0.5_LOO_prop_var.csv")) 
ElasticNet0.5_multiCV_Pct_var <- read_csv(file.path("plant","output","ElasticNet0.5_multiCV_Pct_var.csv")) # Updated version from 11_take2_Machine_Learning
# 
ElasticNet0.5_multiCV_Pct_var <- ElasticNet0.5_multiCV_Pct_var %>%
  #mutate(Module_Name=str_to_upper(Module_Name)) %>%  # tissue is lowercase, module is uppercase
  unite(tissue.module,c("Module_Organ","Module_Name"),sep=".") # all lower case
```


# reading GSEA results (all WGCNA modules)
# GSEA analysis with WGCNA modules
## read reports 
## function for oct30 reports copied from "10_WGCNA_including_R500.Rmd"
```{r}
read.gsea_report_files_WGCNA <- function(gsea_report_path,GSEA.FDR.cutoff,tp=c("resid","removeblock"),data="exp_gt_block") {
# open xsl files and save as csvs
# gsea_report
gsea_report_files <- list.files(path=gsea_report_path,full.names=FALSE,include.dirs = FALSE,recursive=TRUE, pattern="(gsea_report_for)([[:print:]]+)(\\.xls)")
# ifelse(setCutoff=="Total", x <- "Total",if(setCutoff<10) {x <-  str_c("In0",setCutoff)} else {x <- str_c("In",setCutoff)})

# gsea_report_files <- gsea_report_files[str_detect(gsea_report_files,x)]
# print(str_c(x,"orless"))

gsea_report_files
# gsea_report (full name)
gsea_report_files2 <- list.files(path=gsea_report_path,full.names=TRUE,recursive=TRUE, pattern="(gsea_report_for)([[:print:]]+)(\\.xls)")
#gsea_report_files2 <- gsea_report_files2[str_detect(gsea_report_files2,x)] 
#gsea_report_files2
# read xls file (file name is "xls", actually tsv. That is why read_xls did not work)
GSEA_reports<-lapply(gsea_report_files2, function(x) read_tsv(file=file.path(x)))
# name
names(GSEA_reports) <- gsea_report_files
  gsub(".xls","",gsub("([[:print:]]+)(/)([[:print:]]+)","\\3",gsea_report_files))
# convert list into one data.frame
GSEA_reports.DF <- GSEA_reports %>% enframe(name="report") %>% unnest(value)
#GSEA_reports.DF %>% View()
# filter only significnt `FDR q-val` (GSEA.FDR.cutoff)
GSEA_reports.DF <- GSEA_reports.DF %>% filter(`FDR q-val`< GSEA.FDR.cutoff)
# 
GSEA_reports.DF <- GSEA_reports.DF %>% separate(report,into=c("report2","report3"),sep="/") %>% mutate(report2=str_replace(report2,"remove.block","removeblock")) %>% separate(report2,into=c("exp","type","type2","tissue","categories"),sep="\\.",fill="left",extra="drop") 
#
GSEA_reports.DF <- GSEA_reports.DF %>% filter(type2==data) # works?
#GSEA_reports.DF$LorD <- gsub("(gsea_report_for_)(0|1)([[:print:]]+)","\\2",GSEA_reports.DF$report3)
#GSEA_reports.DF <- GSEA_reports.DF %>% mutate(LorD=ifelse(LorD==0,"Live","Dead"))
GSEA_reports.DF %>% View()
ifelse(dim(GSEA_reports.DF)[1]==0,print("no significnat categories"),return(GSEA_reports.DF))
}
```






# data: leaf.voom5.module or root,voom5.module (WGCNA modules calculated either from voom_expression.e1and3.resid.exp_gt.root.plusR500 or voom_expression.e1and3.resid.exp_gt.leaf.plusR500). Expression data is "voom_expression.e1and3.resid.exp_gt_block.root.plusR500.txt" or "voom_expression.e1and3.resid.exp_gt_block.leaf.plusR500.txt
```{r}
type="resid";cutoff <- 1;
gsea_report_path=file.path("~","Box","gsea_home","output","oct30")
GSEA_reports.DF <- read.gsea_report_files_WGCNA(gsea_report_path,GSEA.FDR.cutoff=cutoff) # 
# select root.voom5 and leaf.voom5 data for categories
GSEA_reports.DF <- GSEA_reports.DF %>% filter(str_detect(categories,"voom5")) #%>% View()

# rename "NAME" column (for module names)
GSEA_reports.DF <- GSEA_reports.DF %>% mutate(NAME= str_remove(NAME,"MODULE.")) #%>%
  #str_replace("ROOT.VOOM5","voom_expression.e1and3.resid.exp_gt.root.plusR500") %>%
  #str_replace("LEAF.VOOM5","voom_expression.e1and3.resid.exp_gt.leaf.plusR500"))  #%>% View()
# adding L or D info
GSEA_reports.DF$LorD <- gsub("(gsea_report_for_)(0|1)([[:print:]]+)","\\2",GSEA_reports.DF$report3)
GSEA_reports.DF <- GSEA_reports.DF %>% mutate(LorD=ifelse(LorD==0,"Live","Dead"))
GSEA_reports.DF %>% View()
```

# selecting GSEA results for Machine Learning selected modules
```{r}
GSEA_reports.DF.MLselected <- 
GSEA_reports.DF %>% mutate(NAME=NAME %>% str_remove_all(".VOOM5") %>% str_replace_all("LEAF.","leaf.") %>% str_replace_all("ROOT.","root.")) %>% 
inner_join(ElasticNet0.5_LOO_prop_var,by=c("NAME"="tissue.module")) #%>% View()
```

# modify this
```{r}


modTrait.leaf_avg.RP2.cor.GSEA.summary %>% View()
gsea_report_path=file.path("~","Box","gsea_home","output","aug15")
xls.file <- modTrait.leaf_avg.RP2.cor.GSEA.summary %>% dplyr::select(report3) %>% as_vector()

# function
read.coregenes <- function(default.path=gsea_report_path,
                           xls.file.in.GSEA_reports=report3,
                           category.name = NAME
)
  
{
  gsea_report_files <- list.files(path=default.path,full.names=FALSE,include.dirs = FALSE,recursive=TRUE, pattern="(gsea_report_for)([[:print:]]+)(\\.xls)")
  
  temp.dir <- gsea_report_files[str_detect(gsea_report_files,pattern=xls.file.in.GSEA_reports)] %>% str_split(pattern="/") %>% unlist() %>% .[1]
  print(str_c("temp.dir is ",temp.dir))
  # xls.file.name to read core gene list file
  xls.file.name <- str_c(category.name,".xls") #%>% str_replace("UP","FC_UP") %>% str_replace("DOWN","FC_DOWN")
  # 
  core.data <- read_tsv(file.path(default.path,temp.dir,xls.file.name)) %>% mutate(dirname=temp.dir) # actually the xls file is tsv files (found by Julin...) 
  return(core.data)
}  
```

# core enrichment genes (from 10_WGCNA_GSEA_core_enrichment.Rmd)
```{r}
coregenes.yes.DF <- read_csv(file.path("RNA/output/core_enrichment_genes.csv"))
```

# Overlaying RNAseq data
## reading expression data
```{r}
root.voom5 <- readr::read_tsv(file.path("RNA","output","voom_expression.e1and3.resid.exp_gt.root.plusR500.txt.gz"))
leaf.voom5 <- readr::read_tsv(file.path("RNA","output","voom_expression.e1and3.resid.exp_gt.leaf.plusR500.txt.gz"))
```
## ratio of expression value for soil treated or dead plant
###  sample description data frame {data-navmenu="Data import"}
```{r}
sample.description.e1 <- read_csv(file.path("RNA","input","Br.mbio.e1.sample.description.csv"))
sample.description.e3 <- read_csv(file.path("RNA","input","Br.mbio.e3.sample.description.csv"))

sample.description.e1 %>% summarize(n_distinct(group))
sample.description.e1 %>% group_by(trt) %>% summarize(n())
##
root.voom5.e1 <- tibble(root.voom5 = colnames(root.voom5)[-1]) %>% inner_join(sample.description.e1[,c("sample","genotype","tissue","trt")],by=c("root.voom5"="sample"))
root.voom5.e3 <- tibble(root.voom5 = colnames(root.voom5)[-1]) %>% mutate(genotype="FPsc") %>% inner_join(sample.description.e3[,c("sample","tissue","trt")],by=c("root.voom5"="sample"))  
root.voom5.sample <- bind_rows(root.voom5.e1,root.voom5.e3) %>% mutate(trt=str_remove(trt,"5E_"))
root.voom5.sample %>% group_by(genotype) %>% summarise(num=n())

# leaf.voom5.sample
leaf.voom5.e1 <- tibble(leaf.voom5 = colnames(leaf.voom5)[-1]) %>% inner_join(sample.description.e1[,c("sample","genotype","tissue","trt")],by=c("leaf.voom5"="sample"))
leaf.voom5.e3 <- tibble(leaf.voom5 = colnames(leaf.voom5)[-1]) %>% mutate(genotype="FPsc") %>% inner_join(sample.description.e3[,c("sample","tissue","trt")],by=c("leaf.voom5"="sample"))  
leaf.voom5.sample <- bind_rows(leaf.voom5.e1,leaf.voom5.e3) %>% mutate(trt=str_remove(trt,"5E_"))
leaf.voom5.sample %>% group_by(genotype) %>% summarise(num=n())
```

# mean value
```{r}
# root
root.voom5.soil.mean <- root.voom5 %>% pivot_longer(cols=-1,names_to="sampleID",values_to="value") %>% left_join(root.voom5.sample,by=c("sampleID"="root.voom5")) %>% group_by(genes,trt) %>% summarise(mean.value=mean(value)) 
# leaf
leaf.voom5.soil.mean <- leaf.voom5 %>% pivot_longer(cols=-1,names_to="sampleID",values_to="value") %>% left_join(leaf.voom5.sample,by=c("sampleID"="leaf.voom5")) %>% group_by(genes,trt) %>% summarise(mean.value=mean(value)) 
```

## select leaf length associated modules
```{r}
ElasticNet0.5_multiCV_Pct_var
WGCNA.modules2.voom5 <- read_csv(file.path("RNA","output","WGCNA.modules2.voom5.csv.gz"))

root.voom5.soil.mean 
```

## convert gene ID from V3.0 to PMN ()
```{r}

```




