---
title: "PMN_Pathway Collage from an Omics Dataset"
author: "Kazunari Nozue"
date: "2/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# [Pathway Collage from an Omics Dataset](https://pmn.plantcyc.org/PToolsWebsiteHowto.shtml#celOverview)

```{r}
library(tidyverse);library(readxl)
```

# Automated + mannual compound name curration for PMN
* automated curation by PMN (imported to SmartTable)
* manual curation (added to the imported SmartTable) by using search engine in PMN and "add" function.
* summarized in "new_unique_associated_metabolites_02_09_21_PMNcuration.xlsx"

```{r}
mets.PMN <- read_xlsx("../output/new_unique_associated_metabolites_02_09_21_PMNcuration.xlsx")

```

# read
```{r}
met_amt_rotation_out <- read_csv("../output/Leaf_associated_metabolites_raw_NOBLANK.csv")

```

# combine metabolite conversion table
```{r}
met_amt_rotation_out %>% inner_join(mets.PMN,by="metabolite") %>% dplyr::select(Compound_id,PC,loading) %>% #group_by(PC) %>% summarize(n())
  filter(PC=="root_PC3") %>% write_tsv(file.path("..","output","Leaf_associated_metabolites_raw_NOBLANK.PMN.root_PC3.tsv"))
```

# How to map the metabolite data onto PMN cellular view? (under construction)
* 
