---
title: "Metabolites...Back to basics"
author: "Julin Maloof"
date: "02/22/2021"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Take a different approach with metabolites:

1) ID metabolites responding to microbes
2) Ask if those are associated with leaf length

Here include genotype for the treatment models but use scaled and centered values for leaf associations.

NOTE: THIS IS THE FINAL ANALYSIS. TRIED AND DID NOT LIKE LME MODELS.  TOO MANY SINGULARITIES.  I THINK THERE IS NOT ENOUGH DATA TO ESTIMATE BLOCK EFFECCTS WELL.

NOTE: NEXT STEPS:  TRY MAKING TERNARY PLOT WITH SIMPLE MEANS.


```{r}
library(tidyverse)
library(broom)
```

get leaflength data
```{r}
leaflength <- read_csv("../../plant/output/leaf_lengths_metabolite.csv") %>%
  mutate(pot=str_pad(pot, width=3, pad="0"),
         sampleID=str_c("wyo", genotype, pot, sep="_")) %>%
  select(sampleID, genotype, trt, leaf_avg_std)
leaflength %>% arrange(sampleID)
```

get and wrangle metabolite data
```{r}
met_raw <-read_csv("../input/metabolites_set1.csv")
met <- met_raw %>% 
  mutate(pot=str_pad(pot, width = 3, pad = "0")) %>%
  mutate(sampleID=str_c("wyo", genotype, pot, sep="_")) %>%
  mutate(trt=ifelse(is.na(autoclave), "BLANK", autoclave)) %>%
  select(sampleID, genotype, tissue, trt, sample_mass = `sample_mass mg`, !submission_number:concatenate)  %>%
  
  #bring in leaf length
  left_join({leaflength %>% select(sampleID, leaf_avg_std)}) %>%
  select(sampleID, genotype, tissue, trt, leaf_avg_std, everything()) %>%
  
  #make long
  pivot_longer(!sampleID:sample_mass, names_to = "metabolite", values_to = "met_amt") %>%
  
  #filter away unnamed
  filter(str_detect(metabolite, pattern="^[0-9]+$", negate=TRUE)) %>%
  
  #adjust by sample mass
  mutate(met_per_mg=met_amt/sample_mass) %>%
  
  pivot_wider(id_cols = c(sampleID, genotype, trt, leaf_avg_std), 
              names_from = c(tissue, metabolite), 
              values_from = starts_with("met_"),
              names_sep = "_")

met 
```

split this into two data frames, one normalized by tissue amount and one not.
```{r}
met_per_mg <- met %>% select(sampleID, genotype, trt, leaf_avg_std, starts_with("met_per_mg")) %>%
  as.data.frame() %>% column_to_rownames("sampleID")
met_amt <- met %>% select(sampleID,  genotype, trt, leaf_avg_std, starts_with("met_amt")) %>%
  as.data.frame() %>% column_to_rownames("sampleID")
```


## find sig metabolites

Live vs blank dead

normalized
```{r}
met_per_mg_lm <- met_per_mg %>%
  mutate(trt=ifelse(str_detect(trt, "live"), "live", "blank_dead")) %>%
  pivot_longer(cols=starts_with("met"), names_to = "metabolite") %>%
  mutate(tissue=str_extract(metabolite, "leaf|root")) %>%

  #scale and center
  group_by(metabolite, genotype, tissue) %>%
  mutate(value_scale=scale(value)) %>%
  group_by(metabolite) %>%
  nest() %>%
  mutate(lm_trt=map(data, ~ lm(value~ trt*genotype   , data=.)),
         lm_leaf=map(data, ~ lm(leaf_avg_std ~ value_scale, data=.)))

```


```{r}
met_per_mg_lm_results_trt <- met_per_mg_lm %>% mutate(broomtidy.trt = map(lm_trt, broom::tidy)) %>%
  unnest(broomtidy.trt) %>%
  select(metabolite, term, estimate, p.value) %>%
  pivot_wider(id_cols = metabolite, names_from = term, values_from = c(estimate, p.value), names_prefix = "trt_") %>%
  select(-`p.value_trt_(Intercept)`)

met_per_mg_lm_results_leaf <- met_per_mg_lm %>% mutate(broomtidy.leaf = map(lm_leaf, broom::tidy)) %>%
  unnest(broomtidy.leaf) %>%
  select(metabolite, term, estimate, p.value) %>%
  pivot_wider(id_cols = metabolite, names_from = term, values_from = c(estimate, p.value), names_prefix = "leaf_") %>%
  select(-`p.value_leaf_(Intercept)`)

met_per_mg_lm_results <- inner_join(met_per_mg_lm_results_trt, met_per_mg_lm_results_leaf) %>%
  mutate(tissue=str_extract(metabolite,"(leaf|root)")) %>%
  group_by(tissue) %>%
  mutate(across(starts_with("p.value"), .fns = p.adjust, method = "fdr", .names = "FDR_{.col}")) %>%
  rename_with(~ str_replace(.x, "FDR_p.value", "FDR")) %>%
  ungroup()

met_per_mg_lm_results %>% 
  mutate(both=(FDR_trt_trtlive < 0.1) & (FDR_leaf_value_scale < 0.1)) %>%
  group_by(tissue) %>%
  summarize(across(starts_with("FDR"), ~ sum(.x < .1)), both=sum(both), count=n())

met_per_mg_lm_results %>% 
  mutate(both=(p.value_trt_trtlive < 0.1) & (p.value_leaf_value_scale < 0.1)) %>%
  group_by(tissue) %>%
  summarize(across(starts_with("p.value"), ~ sum(.x < .05)), both=sum(both), count=n())

met_per_mg_lm_results %>% filter(FDR_trt_trtlive < 0.1 & FDR_leaf_value_scale < 0.1) %>% 
  select(metabolite, FDR_trt_trtlive, FDR_leaf_value_scale)



met_per_mg_lm_results %>% 
  #filter(FDR.trt < 0.1 | FDR.leaf < 0.1) %>%
  mutate(sig.both = FDR_trt_trtlive < 0.1 & FDR_leaf_value_scale < 0.1) %>%
  arrange(desc(sig.both), min(.$FDR_trt_trtlive, .$FDR_leaf_value_scale)) %>% 
  write_csv("../output/met_per_mg_lm_full.csv")
```

raw
```{r}
met_amt_lm <- met_amt %>%
  mutate(trt=ifelse(str_detect(trt, "live"), "live", "blank_dead")) %>%
  pivot_longer(cols=starts_with("met"), names_to = "metabolite") %>%
  mutate(tissue=str_extract(metabolite, "leaf|root")) %>%

  #scale and center
  group_by(metabolite, genotype, tissue) %>%
  mutate(value_scale=scale(value)) %>%
  group_by(metabolite) %>%
  nest() %>%
  mutate(lm_trt=map(data, ~ lm(value ~ trt * genotype, data=.)),
         lm_leaf=map(data, ~ lm(leaf_avg_std ~ value_scale, data=.)))

```


```{r}
met_amt_lm_results_trt <- met_amt_lm %>% mutate(broomtidy.trt = map(lm_trt, broom::tidy)) %>%
  unnest(broomtidy.trt) %>%
  select(metabolite, term, estimate, p.value) %>%
  pivot_wider(id_cols = metabolite, names_from = term, values_from = c(estimate, p.value), names_prefix = "trt_") %>%
  select(-`p.value_trt_(Intercept)`)

met_amt_lm_results_leaf <- met_amt_lm %>% mutate(broomtidy.leaf = map(lm_leaf, broom::tidy)) %>%
  unnest(broomtidy.leaf) %>%
  select(metabolite, term, estimate, p.value) %>%
  pivot_wider(id_cols = metabolite, names_from = term, values_from = c(estimate, p.value), names_prefix = "leaf_") %>%
  select(-`p.value_leaf_(Intercept)`)

met_amt_lm_results <- inner_join(met_amt_lm_results_trt, met_amt_lm_results_leaf) %>%
  mutate(tissue=str_extract(metabolite,"(leaf|root)")) %>%
  group_by(tissue) %>%
  mutate(across(starts_with("p.value"), .fns = p.adjust, method = "fdr", .names = "FDR_{.col}")) %>%
  rename_with(~ str_replace(.x, "FDR_p.value", "FDR")) %>%
  ungroup()

met_amt_lm_results %>% 
  mutate(both=(FDR_trt_trtlive < 0.1) & (FDR_leaf_value_scale < 0.1)) %>%
  group_by(tissue) %>%
  summarize(across(starts_with("FDR"), ~ sum(.x < .1)), both=sum(both), count=n())

met_amt_lm_results %>% 
  mutate(both=(p.value_trt_trtlive < 0.1) & (p.value_leaf_value_scale < 0.1)) %>%
  group_by(tissue) %>%
  summarize(across(starts_with("p.value"), ~ sum(.x < .05)), both=sum(both), count=n())

met_amt_lm_results %>% filter(FDR_trt_trtlive < 0.1 & FDR_leaf_value_scale < 0.1) %>% 
  select(metabolite, FDR_trt_trtlive, FDR_leaf_value_scale)

met_amt_lm_results %>% 
  #filter(FDR.trt < 0.1 | FDR.leaf < 0.1) %>%
  mutate(sig.both = FDR_trt_trtlive < 0.1 & FDR_leaf_value_scale < 0.1) %>%
  arrange(desc(sig.both), min(.$FDR_trt_trtlive, .$FDR_leaf_value_scale)) %>% 
  write_csv("../output/met_amt_lm_full.csv")
```

# NO BLANK


## find sig metabolites
Live vs  dead

normalized
```{r}
met_per_mg_lm <- met_per_mg %>%
  filter(trt != "BLANK") %>% 
  mutate(trt=ifelse(str_detect(trt, "live"), "live", "blank_dead")) %>%
  pivot_longer(cols=starts_with("met"), names_to = "metabolite") %>%
  mutate(tissue=str_extract(metabolite, "leaf|root")) %>%

  #scale and center
  group_by(metabolite, genotype, tissue) %>%
  mutate(value_scale=scale(value)) %>%
  group_by(metabolite) %>%
  nest() %>%
  mutate(lm_trt=map(data, ~ lm(value ~ trt * genotype, data=.)),
         lm_leaf=map(data, ~ lm(leaf_avg_std ~ value_scale, data=.)))

```


```{r}
met_per_mg_lm_results_trt <- met_per_mg_lm %>% mutate(broomtidy.trt = map(lm_trt, broom::tidy)) %>%
  unnest(broomtidy.trt) %>%
  select(metabolite, term, estimate, p.value) %>%
  pivot_wider(id_cols = metabolite, names_from = term, values_from = c(estimate, p.value), names_prefix = "trt_") %>%
  select(-`p.value_trt_(Intercept)`)

met_per_mg_lm_results_leaf <- met_per_mg_lm %>% mutate(broomtidy.leaf = map(lm_leaf, broom::tidy)) %>%
  unnest(broomtidy.leaf) %>%
  select(metabolite, term, estimate, p.value) %>%
  pivot_wider(id_cols = metabolite, names_from = term, values_from = c(estimate, p.value), names_prefix = "leaf_") %>%
  select(-`p.value_leaf_(Intercept)`)

met_per_mg_lm_results <- inner_join(met_per_mg_lm_results_trt, met_per_mg_lm_results_leaf) %>%
  mutate(tissue=str_extract(metabolite,"(leaf|root)")) %>%
  group_by(tissue) %>%
  mutate(across(starts_with("p.value"), .fns = p.adjust, method = "fdr", .names = "FDR_{.col}")) %>%
  rename_with(~ str_replace(.x, "FDR_p.value", "FDR")) %>%
  ungroup()

met_per_mg_lm_results %>% 
  mutate(both=(FDR_trt_trtlive < 0.1) & (FDR_leaf_value_scale < 0.1)) %>%
  group_by(tissue) %>%
  summarize(across(starts_with("FDR"), ~ sum(.x < .1)), both=sum(both), count=n())

met_per_mg_lm_results %>% filter(FDR_trt_trtlive < 0.1 & FDR_leaf_value_scale < 0.1) %>% 
  select(metabolite, FDR_trt_trtlive, FDR_leaf_value_scale)

met_per_mg_lm_results %>% 
  #filter(FDR.trt < 0.1 | FDR.leaf < 0.1) %>%
  mutate(sig.both = FDR_trt_trtlive < 0.1 & FDR_leaf_value_scale < 0.1) %>%
  arrange(desc(sig.both), min(.$FDR_trt_trtlive, .$FDR_leaf_value_scale)) %>% 
  write_csv("../output/met_per_mg_lm_full_NOBLANK.csv")
```

raw
```{r}
met_amt_lm <- met_amt %>%
  filter(trt != "BLANK") %>% 
  pivot_longer(cols=starts_with("met"), names_to = "metabolite") %>%
  mutate(tissue=str_extract(metabolite, "leaf|root")) %>%

  #scale and center
  group_by(metabolite, genotype, tissue) %>%
  mutate(value_scale=scale(value)) %>%
  group_by(metabolite) %>%
  nest() %>%
  mutate(lm_trt=map(data, ~ lm(value ~ trt * genotype, data=.)),
         lm_leaf=map(data, ~ lm(leaf_avg_std ~ value_scale, data=.)))

```


```{r}
met_amt_lm_results_trt <- met_amt_lm %>% mutate(broomtidy.trt = map(lm_trt, broom::tidy)) %>%
  unnest(broomtidy.trt) %>%
  select(metabolite, term, estimate, p.value) %>%
  pivot_wider(id_cols = metabolite, names_from = term, values_from = c(estimate, p.value), names_prefix = "trt_") %>%
  select(-`p.value_trt_(Intercept)`)

met_amt_lm_results_leaf <- met_amt_lm %>% mutate(broomtidy.leaf = map(lm_leaf, broom::tidy)) %>%
  unnest(broomtidy.leaf) %>%
  select(metabolite, term, estimate, p.value) %>%
  pivot_wider(id_cols = metabolite, names_from = term, values_from = c(estimate, p.value), names_prefix = "leaf_") %>%
  select(-`p.value_leaf_(Intercept)`)

met_amt_lm_results <- inner_join(met_amt_lm_results_trt, met_amt_lm_results_leaf) %>%
  mutate(tissue=str_extract(metabolite,"(leaf|root)")) %>%
  group_by(tissue) %>%
  mutate(across(starts_with("p.value"), .fns = p.adjust, method = "fdr", .names = "FDR_{.col}")) %>%
  rename_with(~ str_replace(.x, "FDR_p.value", "FDR")) %>%
  ungroup()

met_amt_lm_results %>% 
  mutate(both=(FDR_trt_trtlive < 0.1) & (FDR_leaf_value_scale < 0.1)) %>%
  group_by(tissue) %>%
  summarize(across(starts_with("FDR"), ~ sum(.x < .1)), both=sum(both), count=n())

met_amt_lm_results %>% filter(FDR_trt_trtlive < 0.1 & FDR_leaf_value_scale < 0.1) %>% 
  select(metabolite, FDR_trt_trtlive, FDR_leaf_value_scale)

met_amt_lm_results %>% 
  #filter(FDR.trt < 0.1 | FDR.leaf < 0.1) %>%
  mutate(sig.both = FDR_trt_trtlive < 0.1 & FDR_leaf_value_scale < 0.1) %>%
  arrange(desc(sig.both), min(.$FDR_trt_trtlive, .$FDR_leaf_value_scale)) %>% 
  write_csv("../output/met_amt_lm_full_NOBLANK.csv")
```
