---
title: "Metabolites...Back to basics"
author: "Julin Maloof"
date: "02/22/2021"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Take a different approach with metabolites:

1) ID metabolites responding to microbes
2) Ask if those are associated with leaf length

```{r}
library(glmnet)
library(relaimpo)
library(tidyverse)
library(broom)
```

get leaflength data
```{r}
leaflength <- read_csv("../../plant/output/leaf_lengths_metabolite.csv") %>%
  mutate(pot=str_pad(pot, width=3, pad="0"),
         sampleID=str_c("wyo", genotype, pot, sep="_")) %>%
  select(sampleID, genotype, trt, leaf_avg_std)
leaflength %>% arrange(sampleID)
```

get and wrangle metabolite data
```{r}
met_raw <-read_csv("../input/metabolites_set1.csv")
met <- met_raw %>% 
  mutate(pot=str_pad(pot, width = 3, pad = "0")) %>%
  mutate(sampleID=str_c("wyo", genotype, pot, sep="_")) %>%
  mutate(trt=ifelse(is.na(autoclave), "BLANK", autoclave)) %>%
  select(sampleID, genotype, tissue, trt, sample_mass = `sample_mass mg`, !submission_number:concatenate)  %>%
  
  #bring in leaf length
  left_join({leaflength %>% select(sampleID, leaf_avg_std)}) %>%
  select(sampleID, genotype, tissue, trt, leaf_avg_std, everything()) %>%

  #make long
  pivot_longer(!sampleID:sample_mass, names_to = "metabolite", values_to = "met_amount") %>%
  
  #filter away unnamed
  filter(str_detect(metabolite, pattern="^[0-9]+$", negate=TRUE)) %>%
  
  #adjust by sample mass
  mutate(met_per_mg=met_amount/sample_mass) %>%
  
  #scale and center
  group_by(metabolite, genotype, tissue) %>%
  mutate(met_per_mg=scale(met_per_mg),
         met_amt=scale(met_amount)
  ) %>% 
  pivot_wider(id_cols = c(sampleID, genotype, trt, leaf_avg_std), 
              names_from = c(tissue, metabolite), 
              values_from = starts_with("met_"),
              names_sep = "_")

met 
```

split this into two data frames, one normalized by tissue amount and one not.
```{r}
met_per_mg <- met %>% select(sampleID, genotype, trt, leaf_avg_std, starts_with("met_per_mg")) %>%
  as.data.frame() %>% column_to_rownames("sampleID")
met_amt <- met %>% select(sampleID,  genotype, trt, leaf_avg_std, starts_with("met_amt")) %>%
  as.data.frame() %>% column_to_rownames("sampleID")
```


## find sig metabolites

Live vs blank dead

normalized
```{r}
met_per_mg_lm <- met_per_mg %>%
  mutate(trt=ifelse(str_detect(trt, "live"), "live", "blank_dead")) %>%
  pivot_longer(cols=starts_with("met"), names_to = "metabolite") %>%
  group_by(metabolite) %>%
  nest() %>%
  mutate(lm_trt=map(data, ~ lm(value ~ trt, data=.)),
         lm_leaf=map(data, ~ lm(leaf_avg_std ~ value, data=.)))

```


```{r}
met_per_mg_lm_results<- met_per_mg_lm %>% mutate(broomtidy.trt = map(lm_trt, broom::tidy),
                         broomtidy.leaf = map(lm_leaf, broom::tidy)) %>%
  unnest(broomtidy.trt, broomtidy.leaf) %>%
  select(metabolite, term.trt=term, p.value.trt=p.value,
         term.leaf=term1, p.value.leaf=p.value1) %>%
  mutate(FDR.trt=p.adjust(p.value.trt, method = "fdr"),
         FDR.leaf=p.adjust(p.value.leaf, method = "fdr")) %>%
  filter(! str_detect(term.trt, "Intercept"),
         ! str_detect(term.leaf, "Intercept"))

met_per_mg_lm_results %>% ungroup() %>%
  summarize(sig.trt=sum(FDR.trt<0.1), sig.leaf=sum(FDR.leaf<0.1), sig.both=sum(FDR.trt<0.1&FDR.leaf<0.1), count=n())

met_per_mg_lm_results %>% filter(FDR.trt < 0.1 & FDR.leaf < 0.1) %>% 
  select(metabolite, FDR.trt, FDR.leaf)
```

raw
```{r}
met_amt_lm <- met_amt %>%
  mutate(trt=ifelse(str_detect(trt, "live"), "live", "blank_dead")) %>%
  pivot_longer(cols=starts_with("met"), names_to = "metabolite") %>%
  group_by(metabolite) %>%
  nest() %>%
  mutate(lm_trt=map(data, ~ lm(value ~ trt, data=.)),
         lm_leaf=map(data, ~ lm(leaf_avg_std ~ value, data=.)))

```


```{r}
met_amt_lm_results<- met_amt_lm %>% mutate(broomtidy.trt = map(lm_trt, broom::tidy),
                         broomtidy.leaf = map(lm_leaf, broom::tidy)) %>%
  unnest(broomtidy.trt, broomtidy.leaf) %>%
  select(metabolite, term.trt=term, p.value.trt=p.value,
         term.leaf=term1, p.value.leaf=p.value1) %>%
  mutate(FDR.trt=p.adjust(p.value.trt, method = "fdr"),
         FDR.leaf=p.adjust(p.value.leaf, method = "fdr")) %>%
  filter(! str_detect(term.trt, "Intercept"),
         ! str_detect(term.leaf, "Intercept"))

met_amt_lm_results %>% ungroup() %>%
  summarize(sig.trt=sum(FDR.trt<0.1), sig.leaf=sum(FDR.leaf<0.1), sig.both=sum(FDR.trt<0.1&FDR.leaf<0.1), count=n())

met_amt_lm_results %>% filter(FDR.trt < 0.1 & FDR.leaf < 0.1) %>% select(metabolite, FDR.trt, FDR.leaf)
```


# NO BLANK



## find sig metabolites

Live vs  dead

normalized
```{r}
met_per_mg_lm_nb <- met_per_mg %>%
  filter(trt!="BLANK") %>%
  pivot_longer(cols=starts_with("met"), names_to = "metabolite") %>%
  group_by(metabolite) %>%
  nest() %>%
  mutate(lm_trt=map(data, ~ lm(value ~ trt, data=.)),
         lm_leaf=map(data, ~ lm(leaf_avg_std ~ value, data=.)))

```


```{r}
met_per_mg_lm_results_nb <- met_per_mg_lm_nb %>% mutate(broomtidy.trt = map(lm_trt, broom::tidy),
                         broomtidy.leaf = map(lm_leaf, broom::tidy)) %>%
  unnest(broomtidy.trt, broomtidy.leaf) %>%
  select(metabolite, term.trt=term, p.value.trt=p.value,
         term.leaf=term1, p.value.leaf=p.value1) %>%
  mutate(FDR.trt=p.adjust(p.value.trt, method = "fdr"),
         FDR.leaf=p.adjust(p.value.leaf, method = "fdr")) %>%
  filter(! str_detect(term.trt, "Intercept"),
         ! str_detect(term.leaf, "Intercept"))

met_per_mg_lm_results_nb %>% ungroup() %>%
  summarize(sig.trt=sum(FDR.trt<0.1), sig.leaf=sum(FDR.leaf<0.1), sig.both=sum(FDR.trt<0.1&FDR.leaf<0.1), count=n())

met_per_mg_lm_results_nb %>% filter(FDR.trt < 0.1 & FDR.leaf < 0.1) %>% select(metabolite, FDR.trt, FDR.leaf)
```

raw
```{r}
met_amt_lm_nb <- met_amt %>%
  filter(trt!="BLANK") %>%
  pivot_longer(cols=starts_with("met"), names_to = "metabolite") %>%
  group_by(metabolite) %>%
  nest() %>%
  mutate(lm_trt=map(data, ~ lm(value ~ trt, data=.)),
         lm_leaf=map(data, ~ lm(leaf_avg_std ~ value, data=.)))

```


```{r}
met_amt_lm_results_nb<- met_amt_lm_nb %>% mutate(broomtidy.trt = map(lm_trt, broom::tidy),
                         broomtidy.leaf = map(lm_leaf, broom::tidy)) %>%
  unnest(broomtidy.trt, broomtidy.leaf) %>%
  select(metabolite, term.trt=term, p.value.trt=p.value,
         term.leaf=term1, p.value.leaf=p.value1) %>%
  mutate(FDR.trt=p.adjust(p.value.trt, method = "fdr"),
         FDR.leaf=p.adjust(p.value.leaf, method = "fdr")) %>%
  filter(! str_detect(term.trt, "Intercept"),
         ! str_detect(term.leaf, "Intercept"))

met_amt_lm_results_nb %>% ungroup() %>%
  summarize(sig.trt=sum(FDR.trt<0.1), sig.leaf=sum(FDR.leaf<0.1), sig.both=sum(FDR.trt<0.1&FDR.leaf<0.1), count=n())

met_amt_lm_results_nb %>% filter(FDR.trt < 0.1 & FDR.leaf < 0.1) %>% select(FDR.trt, FDR.leaf)
```
