---
title: "Direct Penalized Regression--Metabolites"
author: "Julin Maloof"
date: "3/09/2021"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Goal: Dan suggested skipping the PCA step and just looking for metabolites associated with leaf length via penalized regression.

```{r}
library(glmnet)
library(relaimpo)
library(tidyverse)
library(broom)
```

get leaflength data
```{r}
leaflength <- read_csv("../../plant/output/leaf_lengths_metabolite.csv") %>%
  mutate(pot=str_pad(pot, width=3, pad="0"),
         sampleID=str_c("wyo", genotype, pot, sep="_")) %>%
  select(sampleID, genotype, trt, leaf_avg_std)
leaflength %>% arrange(sampleID)
```

get and wrangle metabolite data
```{r}
met_raw <-read_csv("../input/metabolites_set1.csv")
met <- met_raw %>% 
  mutate(pot=str_pad(pot, width = 3, pad = "0")) %>%
  mutate(sampleID=str_c("wyo", genotype, pot, sep="_")) %>%
  select(sampleID, genotype, tissue, sample_mass = `sample_mass mg`, !submission_number:concatenate) %>%
  pivot_longer(!sampleID:sample_mass, names_to = "metabolite", values_to = "met_amount") %>%
  
  #remove unnamed metabolites
  filter(str_detect(metabolite, "[A-Z]|[a-z]")) %>%
  
  #adjust by sample mass
  mutate(met_per_mg=met_amount/sample_mass) %>%
  
  #scale and center
  group_by(metabolite, genotype, tissue) %>%
  mutate(met_per_mg=scale(met_per_mg),
         met_amt=scale(met_amount)
  ) %>% 
  pivot_wider(id_cols = sampleID, 
              names_from = c(tissue, metabolite), 
              values_from = starts_with("met_"),
              names_sep = "_")

met 
```

split this into two data frames, one normalized by tissue amount and one not.
```{r}
met_per_mg <- met %>% select(sampleID, starts_with("met_per_mg")) %>%
  as.data.frame() %>% column_to_rownames("sampleID")
met_amt <- met %>% select(sampleID,  starts_with("met_amt")) %>%
  as.data.frame() %>% column_to_rownames("sampleID")
```

get leaf data order to match

```{r}
leaflength <- leaflength[match(met$sampleID, leaflength$sampleID),]
leaflength
```

## find metabolites sig for microbe

Live vs blank dead

normalized
```{r}
met_per_mg_lm <- met_per_mg %>%
  rownames_to_column("sampleID") %>%
  left_join(leaflength) %>%
  mutate(trt=ifelse(str_detect(trt, "live"), "live", "blank_dead")) %>%
  pivot_longer(cols=starts_with("met"), names_to = "metabolite") %>%
  group_by(metabolite) %>%
  nest() %>%
  mutate(lm_trt=map(data, ~ lm(value ~ trt, data=.)),
         lm_leaf=map(data, ~ lm(leaf_avg_std ~ value, data=.)))

```


```{r}
met_per_mg_lm_results<- met_per_mg_lm %>% mutate(broomtidy.trt = map(lm_trt, broom::tidy),
                         broomtidy.leaf = map(lm_leaf, broom::tidy)) %>%
  unnest(broomtidy.trt, broomtidy.leaf) %>%
  select(metabolite, term.trt=term, p.value.trt=p.value,
         term.leaf=term1, p.value.leaf=p.value1) %>%
  mutate(FDR.trt=p.adjust(p.value.trt, method = "fdr"),
         FDR.leaf=p.adjust(p.value.leaf, method = "fdr")) %>%
  filter(! str_detect(term.trt, "Intercept"),
         ! str_detect(term.leaf, "Intercept"))

met_per_mg_lm_results %>% ungroup() %>%
  summarize(sig.trt=sum(FDR.trt<0.1), sig.leaf=sum(FDR.leaf<0.1), sig.both=sum(FDR.trt<0.1&FDR.leaf<0.1), count=n())
```

raw
```{r}
met_amt_lm <- met_amt %>%
  rownames_to_column("sampleID") %>%
  left_join(leaflength) %>%
  mutate(trt=ifelse(str_detect(trt, "live"), "live", "blank_dead")) %>%
  pivot_longer(cols=starts_with("met"), names_to = "metabolite") %>%
  group_by(metabolite) %>%
  nest() %>%
  mutate(lm_trt=map(data, ~ lm(value ~ trt, data=.)),
         lm_leaf=map(data, ~ lm(leaf_avg_std ~ value, data=.)))

```


```{r}
met_amt_lm_results<- met_amt_lm %>% mutate(broomtidy.trt = map(lm_trt, broom::tidy),
                         broomtidy.leaf = map(lm_leaf, broom::tidy)) %>%
  unnest(broomtidy.trt, broomtidy.leaf) %>%
  select(metabolite, term.trt=term, p.value.trt=p.value,
         term.leaf=term1, p.value.leaf=p.value1) %>%
  mutate(FDR.trt=p.adjust(p.value.trt, method = "fdr"),
         FDR.leaf=p.adjust(p.value.leaf, method = "fdr")) %>%
  filter(! str_detect(term.trt, "Intercept"),
         ! str_detect(term.leaf, "Intercept"))

met_amt_lm_results %>% ungroup() %>%
  summarize(sig.trt=sum(FDR.trt<0.1), sig.leaf=sum(FDR.leaf<0.1), sig.both=sum(FDR.trt<0.1&FDR.leaf<0.1), count=n())
```


##  penalized regression for association with leaf

# normalized

## multi CV

Fit 101 CVs for each of 11 alphas
```{r}
set.seed(1245)

folds <- tibble(run=1:101) %>% 
  mutate(folds=map(run, ~ sample(rep(1:6,6))))

system.time (met_per_mg_multiCV <- expand_grid(run=1:100, alpha=round(seq(0,1,.1),1)) %>%
               left_join(folds, by="run") %>%
               mutate(fit=map2(folds, alpha, ~ cv.glmnet(x=as.matrix(met_per_mg),
                                                         y=leaflength$leaf_avg_std, 
                                                         foldid = .x, alpha=.y
                                                         )))
             #, lambda=exp(seq(-5,0,length.out = 50)) )))
) #100 seconds

head(met_per_mg_multiCV)
```

for each fit, pull out the mean cv error, lambda, min lambda, and 1se lambda 
```{r}
met_per_mg_multiCV <- met_per_mg_multiCV %>%
  mutate(cvm=map(fit, magrittr::extract("cvm")),
         lambda=map(fit, magrittr::extract("lambda")),
         lambda.min=map_dbl(fit, magrittr::extract("lambda.min" )),
         lambda.1se=map_dbl(fit, magrittr::extract("lambda.1se")),
         nzero=map(fit, magrittr::extract("nzero"))
  )

head(met_per_mg_multiCV)
```


now calculate the mean and sem of cvm and min,1se labmdas.  These need to be done separately because of the way the grouping works
```{r}
met_per_mg_summary_cvm <- met_per_mg_multiCV %>% dplyr::select(-fit, -folds) %>% 
  unnest(c(cvm, lambda)) %>%
  group_by(alpha, lambda) %>%
  summarize(meancvm=mean(cvm), sem=sd(cvm)/sqrt(n()), high=meancvm+sem, low=meancvm-sem)

met_per_mg_summary_cvm
```

```{r}
met_per_mg_summary_lambda <- met_per_mg_multiCV %>% dplyr::select(-fit, -folds, -cvm) %>% 
  group_by(alpha) %>%
  summarize(
    lambda.min.sd=sd(lambda.min), 
    lambda.min.mean=mean(lambda.min),
    #lambda.min.med=median(lambda.min), 
    lambda.min.high=lambda.min.mean+lambda.min.sd,
    #lambda.min.low=lambda.min.mean-lambda.min.sem,
    #lambda.1se.sem=sd(lambda.1se)/sqrt(n()), 
    lambda.1se.mean=mean(lambda.1se),
    #lambda.1se.med=median(lambda.1se), 
    #lambda.1se.high=lambda.1se+lambda.1se.sem,
    #lambda.1se.low=lambda.1se-lambda.1se.sem,
    nzero=nzero[1],
    lambda=lambda[1]
  )

met_per_mg_summary_lambda
```


plot it
```{r}
met_per_mg_summary_cvm %>%
  #filter(alpha!=0) %>% # worse than everything else and throwing the plots off
  ggplot(aes(x=log(lambda), y= meancvm,  ymin=low, ymax=high)) +
  geom_ribbon(alpha=.25) +
  geom_line(aes(color=as.character(alpha))) +
  facet_wrap(~ as.character(alpha)) +
   coord_cartesian(xlim=(c(-5,0))) +
  geom_vline(aes(xintercept=log(lambda.min.mean)), alpha=.5, data=met_per_mg_summary_lambda) +
  geom_vline(aes(xintercept=log(lambda.min.high)), alpha=.5, data=met_per_mg_summary_lambda, color="blue") 

```

Make a plot of MSE at minimum lambda for each alpha

```{r}
met_per_mg_summary_cvm %>% 
  group_by(alpha) %>%
  filter(rank(meancvm, ties.method = "first")==1) %>%
  ggplot(aes(x=alpha,y=meancvm,ymin=low,ymax=high)) +
  geom_ribbon(color=NA, fill="gray80") +
  geom_line() +
  geom_point()
```

Plot the number of nzero coefficients

```{r}
met_per_mg_summary_lambda %>%
  unnest(c(lambda, nzero)) %>%
  group_by(alpha) %>%
  filter(abs(lambda.min.mean-lambda)==min(abs(lambda.min.mean-lambda))  ) %>%
  ungroup() %>%

ggplot(aes(x=as.character(alpha), y=nzero)) +
  geom_point() +
  ggtitle("Number of non-zero coefficents at minimum lambda") +
  ylim(0,50)
```
OK let's do repeated test train starting from these CV lambdas

```{r}
multi_tt <- function(lambda, alpha, n=10000, sample_size=36, train_size=30, x, y=leaflength$leaf_avg_std) {
  print(lambda)
  print(alpha)
tt <-
  tibble(run=1:n) %>%
  mutate(train=map(run, ~ sample(1:sample_size, train_size))) %>%
  mutate(fit=map(train, ~ glmnet(x=x[.,], y=y[.], lambda = lambda, alpha = alpha ))) %>%
  
  mutate(pred=map2(fit, train, ~ predict(.x, newx = x[-.y,]))) %>%
  mutate(cor=map2_dbl(pred, train, ~ cor(.x, y[-.y])  )) %>%
  mutate(MSE=map2_dbl(pred, train, ~ mean((y[-.y] - .x)^2))) %>%
  summarize(
    num_na=sum(is.na(cor)), 
    num_lt_0=sum(cor<=0, na.rm=TRUE),
    avg_cor=mean(cor, na.rm=TRUE),
    avg_MSE=mean(MSE))
tt
}

per_mg_fit_test_train <- met_per_mg_summary_lambda %>% 
  select(alpha, lambda.min.mean)

per_mg_fit_test_train <- met_per_mg_multiCV %>%
  filter(run==1) %>%
  select(alpha, fit) %>%
  right_join(per_mg_fit_test_train)

per_mg_fit_test_train <- per_mg_fit_test_train %>%
  mutate(pred_full=map2(fit, lambda.min.mean, ~ predict(.x, s=.y, newx=as.matrix(met_per_mg))),
         full_R=map_dbl(pred_full, ~ cor(.x, leaflength$leaf_avg_std)),
         full_MSE=map_dbl(pred_full, ~ mean((leaflength$leaf_avg_std-.x)^2))) %>%
  
  mutate(tt=map2(lambda.min.mean, alpha, ~ multi_tt(lambda=.x, alpha=.y, x=as.matrix(met_per_mg))))



(per_mg_fit_test_train <- per_mg_fit_test_train %>% unnest(tt))
```

```{r}
per_mg_fit_test_train %>%
  ggplot(aes(x=alpha)) +
  geom_line(aes(y=avg_cor), color="red") +
  geom_point(aes(y=avg_cor), color="red") +
  geom_line(aes(y=avg_MSE), color="blue") +
  geom_point(aes(y=avg_MSE), color="blue")
```

## look at fit:

```{r}
alpha_per_mg <- .1

best_per_mg <- per_mg_fit_test_train %>% filter(alpha == alpha_per_mg) 
best_per_mg_fit <- best_per_mg$fit[[1]]
best_per_mg_lambda <- best_per_mg$lambda.min.mean

per_mg_coef.tb <- coef(best_per_mg_fit, s=best_per_mg_lambda) %>% 
  as.matrix() %>% as.data.frame() %>% 
  rownames_to_column(var="metabolite") %>%
  rename(beta=`1`)
  
per_mg_coef.tb %>% filter(beta!=0) %>% arrange(desc(abs(beta)))

```

pred and obs
```{r}
plot(leaflength$leaf_avg_std, best_per_mg$pred_full[[1]])
cor.test(leaflength$leaf_avg_std, best_per_mg$pred_full[[1]]) #.75
best_per_mg$full_MSE
```

# merge with lm results for trt association

```{r}
per_mg_coef.tb_and_lm <- per_mg_coef.tb %>%
  filter(beta !=0, str_detect(metabolite, "Intercept", negate = TRUE)) %>%
  left_join(met_per_mg_lm_results) %>%
  arrange(desc(abs(beta))) %>%
  select(-term.trt, -term.leaf)

write_csv(per_mg_coef.tb_and_lm, "../output/met_per_mg_elastic_leaf_lm_trt.csv")

per_mg_coef.tb_and_lm
```


# non-normazlized

## multi CV

Fit 101 CVs for each of 11 alphas
```{r}
set.seed(1245)

folds <- tibble(run=1:101) %>% 
  mutate(folds=map(run, ~ sample(rep(1:6,6))))

system.time (met_amt_multiCV <- expand_grid(run=1:100, alpha=round(seq(0,1,.1),1)) %>%
               left_join(folds, by="run") %>%
               mutate(fit=map2(folds, alpha, ~ cv.glmnet(x=as.matrix(met_amt), y=leaflength$leaf_avg_std, foldid = .x, alpha=.y
                                                         )))
             #, lambda=exp(seq(-5,0,length.out = 50)) )))
) #200 seconds

head(met_amt_multiCV)
```

for each fit, pull out the mean cv error, lambda, min lambda, and 1se lambda 
```{r}
met_amt_multiCV <- met_amt_multiCV %>%
  mutate(cvm=map(fit, magrittr::extract("cvm")),
         lambda=map(fit, magrittr::extract("lambda")),
         lambda.min=map_dbl(fit, magrittr::extract("lambda.min" )),
         lambda.1se=map_dbl(fit, magrittr::extract("lambda.1se")),
         nzero=map(fit, magrittr::extract("nzero"))
  )

head(met_amt_multiCV)
```


now calculate the mean and sem of cvm and min,1se labmdas.  These need to be done separately because of the way the grouping works
```{r}
met_amt_summary_cvm <- met_amt_multiCV %>% dplyr::select(-fit, -folds) %>% 
  unnest(c(cvm, lambda)) %>%
  group_by(alpha, lambda) %>%
  summarize(meancvm=mean(cvm), sem=sd(cvm)/sqrt(n()), high=meancvm+sem, low=meancvm-sem)

met_amt_summary_cvm
```

```{r}
met_amt_summary_lambda <- met_amt_multiCV %>% dplyr::select(-fit, -folds, -cvm) %>% 
  group_by(alpha) %>%
  summarize(
    lambda.min.sd=sd(lambda.min), 
    lambda.min.mean=mean(lambda.min),
    #lambda.min.med=median(lambda.min), 
    lambda.min.high=lambda.min.mean+lambda.min.sd,
    #lambda.min.low=lambda.min.mean-lambda.min.sem,
    #lambda.1se.sem=sd(lambda.1se)/sqrt(n()), 
    lambda.1se.mean=mean(lambda.1se),
    #lambda.1se.med=median(lambda.1se), 
    #lambda.1se.high=lambda.1se+lambda.1se.sem,
    #lambda.1se.low=lambda.1se-lambda.1se.sem,
    nzero=nzero[1],
    lambda=lambda[1]
  )

met_amt_summary_lambda
```


plot it
```{r}
met_amt_summary_cvm %>%
  #filter(alpha!=0) %>% # worse than everything else and throwing the plots off
  ggplot(aes(x=log(lambda), y= meancvm,  ymin=low, ymax=high)) +
  geom_ribbon(alpha=.25) +
  geom_line(aes(color=as.character(alpha))) +
  facet_wrap(~ as.character(alpha)) +
   coord_cartesian(xlim=(c(-5,0))) +
  geom_vline(aes(xintercept=log(lambda.min.mean)), alpha=.5, data=met_amt_summary_lambda) +
  geom_vline(aes(xintercept=log(lambda.min.high)), alpha=.5, data=met_amt_summary_lambda, color="blue") 

```


Make a plot of MSE at minimum lambda for each alpha

```{r}
met_amt_summary_cvm %>% 
  group_by(alpha) %>%
  filter(rank(meancvm, ties.method = "first")==1) %>%
  ggplot(aes(x=alpha,y=meancvm,ymin=low,ymax=high)) +
  geom_ribbon(color=NA, fill="gray80") +
  geom_line() +
  geom_point()
```

Plot the number of nzero coefficients

```{r}
met_amt_summary_lambda %>%
  unnest(c(lambda, nzero)) %>%
  group_by(alpha) %>%
  filter(abs(lambda.min.mean-lambda)==min(abs(lambda.min.mean-lambda))  ) %>%
  ungroup() %>%

ggplot(aes(x=as.character(alpha), y=nzero)) +
  geom_point() +
  ggtitle("Number of non-zero coefficents at minimum lambda") +
  ylim(0,50)
```
OK let's do repeated test train starting from these CV lambdas

```{r}
multi_tt <- function(lambda, alpha, n=10000, sample_size=36, train_size=30, x, y=leaflength$leaf_avg_std) {
  print(lambda)
  print(alpha)
tt <-
  tibble(run=1:n) %>%
  mutate(train=map(run, ~ sample(1:sample_size, train_size))) %>%
  mutate(fit=map(train, ~ glmnet(x=x[.,], y=y[.], lambda = lambda, alpha = alpha ))) %>%
  
  mutate(pred=map2(fit, train, ~ predict(.x, newx = x[-.y,]))) %>%
  mutate(cor=map2_dbl(pred, train, ~ cor(.x, y[-.y])  )) %>%
  mutate(MSE=map2_dbl(pred, train, ~ mean((y[-.y] - .x)^2))) %>%
  summarize(
    num_na=sum(is.na(cor)), 
    num_lt_0=sum(cor<=0, na.rm=TRUE),
    avg_cor=mean(cor, na.rm=TRUE),
    avg_MSE=mean(MSE))
tt
}

amt_fit_test_train <- met_amt_summary_lambda %>% 
  select(alpha, lambda.min.mean)

amt_fit_test_train <- met_amt_multiCV %>%
  filter(run==1) %>%
  select(alpha, fit) %>%
  right_join(amt_fit_test_train)

amt_fit_test_train <- amt_fit_test_train %>%
  mutate(pred_full=map2(fit, lambda.min.mean, ~ predict(.x, s=.y, newx=as.matrix(met_amt))),
         full_R=map_dbl(pred_full, ~ cor(.x, leaflength$leaf_avg_std)),
         full_MSE=map_dbl(pred_full, ~ mean((leaflength$leaf_avg_std-.x)^2))) %>%
  
  mutate(tt=map2(lambda.min.mean, alpha, ~ multi_tt(lambda=.x, alpha=.y, x=as.matrix(met_amt))))



(amt_fit_test_train <- amt_fit_test_train %>% unnest(tt))
```

```{r}
amt_fit_test_train %>%
  ggplot(aes(x=alpha)) +
  geom_line(aes(y=avg_cor), color="red") +
  geom_point(aes(y=avg_cor), color="red") +
  geom_line(aes(y=avg_MSE), color="blue") +
  geom_point(aes(y=avg_MSE), color="blue")
```
Not a ton of difference here. Ignoring 0.0, then 0.1 or 1 ar the best...kind of strange!
## look at fit:

```{r}
alpha_amt <- .1

best_amt <- amt_fit_test_train %>% filter(alpha == alpha_amt) 
best_amt_fit <- best_amt$fit[[1]]
best_amt_lambda <- best_amt$lambda.min.mean

amt_coef.tb <- coef(best_amt_fit, s=best_amt_lambda) %>% 
  as.matrix() %>% as.data.frame() %>% 
  rownames_to_column(var="metabolite") %>%
  rename(beta=`1`)
  
amt_coef.tb %>% filter(beta!=0) %>% arrange(desc(abs(beta)))

```

pred and obs
```{r}
plot(leaflength$leaf_avg_std, best_amt$pred_full[[1]])
cor.test(leaflength$leaf_avg_std, best_amt$pred_full[[1]]) #.85
best_amt$full_MSE
```

# merge with lm results for trt association

```{r}
amt_coef.tb_and_lm <- amt_coef.tb %>%
  filter(beta !=0, str_detect(metabolite, "Intercept", negate = TRUE)) %>%
  left_join(met_amt_lm_results) %>%
  arrange(desc(abs(beta))) %>%
  select(-term.trt, -term.leaf)

write_csv(amt_coef.tb_and_lm, "../output/met_amt_elastic_leaf_lm_trt.csv")

amt_coef.tb_and_lm
```